### 2023-10-29 paper/main.tex
start paper, stupidly

\documentclass[12pt]{article}
\usepackage[letterpaper]{geometry}
\usepackage{amsmath}

\frenchspacing
\begin{document}

\section*{\raggedright
Another formula for the area of a triangle: Useless, but explicitly in map-reduce format}
\noindent \textbf{David W. Hogg} and \textbf{Soledad Villar}

\paragraph{Abstract:}
It is known that any permutation-invariant function of data points $x_i$
can be written in the form $\rho(\sum_i\phi(x_i))$,
where $\rho$ and $\phi$ are (probably nonlinear) functions, and the function $\phi$ might output high-dimensional output (but can't depend on $i$).
This form can be considered map--reduce, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically.
The area of a triangle is a permutation-invariant function of the locations of the three corners.
We find (embarrassingly, by symbolic regression) the formula for the area of the triangle that is explicitly in this map--reduce form.
It is ugly.

\section{Introduction}
The area $A$ of a triangle can be written in many ways:
\begin{align}
A &= \frac{1}{2}\,\text{(base)}\,\text{(height)} \\
  &= \frac{1}{2}\,|(\vec{x}_3 - \vec{x}_1) \times (\vec{x}_2 - \vec{x}_1)| \\
  &= \frac{1}{2}\,a\,b\,\sin(C) \\
  &= \sqrt{s\,(s-a)\,(s-b)\,(s-c)} \label{eq:Heron} ~,
\end{align}
where the first formula is the usual introduction to the subject,
$(\vec{x}_1, \vec{x}_2, \vec{x}_3$ are three vectors pointing from an origin to the corners of the triangle,
$\times$ represents the cross-product operator (or its generalization to higher dimensions),
$a, b, c$ are side lengths of the triangle,
$C$ is the angle between the $a$ and $b$ sides,
and $s$ is the semi-perimeter $(a + b + c)/2$.

While the area of a triangle does not depend on how its sides or corners are labeled,
only the last of these formulae---Heron's formula \eqref{eq:Heron}---is explicitly permutation-invariant.
The theory of permutation-invariant functions says (CITE) that any per\-mu\-tation-invariant function (under conditions??) of $N$ objects $x_i$ (with $1\leq i\leq N$) can be approximated (to arbitrary precision??) with an expression of the form
\begin{equation}
    \rho(\sum_{i=1}^N\phi(x_i)) \label{eq:DeepSets}
\end{equation}
where $\rho$ and $\phi$ are (possibly very nonlinear) functions,
and the $x_i$ are the data points that can be permuted without changing the result.
This result---which we associate with the subject \emph{Deep Sets} (CITE)---is not at all obvious:
For example, even though Heron's formula \eqref{eq:Heron} is permutation-invariant, it is not actually in the form \eqref{eq:DeepSets}, if the triangle is seen as being defined by its corners (or even its sides).

A few comments about this Deep-Sets form \eqref{eq:DeepSets} we'd make are the following:
\begin{itemize}
    \item The function $\phi$ is universal to all the points $x_i$; it's form won't depend on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
    \item The point properties $x_i$ and the output of the function $\phi$ might both be very large.
    That is, each point $x_i$ might be a vector or even an ordered list of vectors and tensors and images and so on, and the output of the function $\phi$ might also be very large and complicated.
    The output of the function $\phi$ is some kind of informative, latent encoding of the point $x_i$.
    \item If a function can be written in Deep-Sets form \eqref{eq:DeepSets}, then that function can be executed in a map--reduce framework, in which the function $\phi$ is executed independently and asynchronously on each data point (maybe even in the data center) and the sums are performed hierarchically or in any order.
    That means that for large problems executed on large data sets in large data centers, problems in this form \eqref{eq:DeepSets} can be performed in wall-clock time that scales as $\log N$.
\end{itemize}

\section{A note on our method}

\section{Result and discussion}

\begin{align}
A^2 = &
-\frac{ 54}{368}\,(\sum_{i=1}^3 x_i^2\,y_i^2) 
-\frac{258}{368}\,(\sum_{i=1}^3 x_i\,y_i)^2 \nonumber \\ &
+\frac{285}{368}\,(\sum_{i=1}^3 x_i^2)\,(\sum_{i=1}^3 y_i^2) 
+\frac{ 36}{368}\,(\sum_{i=1}^3 y_i)\,(\sum_{i=1}^3 x_i^2\,y_i) \nonumber \\ &
+\frac{ 36}{368}\,(\sum_{i=1}^3 x_i)\,(\sum_{i=1}^3 x_i\,y_i^2) 
-\frac{101}{368}\,(\sum_{i=1}^3 y_i)^2\,(\sum_{i=1}^3 x_i^2) \nonumber \\ &
+\frac{148}{368}\,(\sum_{i=1}^3 x_i)\,(\sum_{i=1}^3 y_i)\,(\sum_{i=1}^3 x_i\,y_i) 
-\frac{101}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i^2) \nonumber \\ &
+\frac{  9}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i)^2 
\end{align}

\end{document}

### 2023-10-29 paper/main.tex
lots of pointless content

@@
 
+\newcommand{\documentname}{\textsl{Note}}
+\newcommand{\sectionname}{Section}
+\newcommand{\secref}[1]{\sectionname~\ref{#1}}
 \frenchspacing
+
 \begin{document}
@@
 \section*{\raggedright
-Another formula for the area of a triangle: Useless, but explicitly in map-reduce format}
+Another formula for the area of a triangle: Useless, but explicitly in Deep-Sets and map--reduce form}
 \noindent \textbf{David W. Hogg} and \textbf{Soledad Villar}
@@
 can be written in the form $\rho(\sum_i\phi(x_i))$,
-where $\rho$ and $\phi$ are (probably nonlinear) functions, and the function $\phi$ might output high-dimensional output (but can't depend on $i$).
-This form can be considered map--reduce, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically.
+where $\rho$ and $\phi$ are (probably nonlinear) functions, and the function $\phi$ might output a high-dimensional latent encoding of each point $x_i$.
+This form---known in the machine-learning literature as Deep Sets---is a map--reduce form, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically or in any order.
 The area of a triangle is a permutation-invariant function of the locations of the three corners.
-We find (embarrassingly, by symbolic regression) the formula for the area of the triangle that is explicitly in this map--reduce form.
-It is ugly.
+We find, possibly for the first time, the formula for the area of the triangle that is explicitly in the Deep-Sets form.
+No insights of any kind were gained by this work.
 
@@
 \begin{align}
-A &= \frac{1}{2}\,\text{(base)}\,\text{(height)} \\
-  &= \frac{1}{2}\,|(\vec{x}_3 - \vec{x}_1) \times (\vec{x}_2 - \vec{x}_1)| \\
-  &= \frac{1}{2}\,a\,b\,\sin(C) \\
+A &= \frac{1}{2}\,\text{(base)}\,\text{(height)} \label{eq:school} \\
+  &= \frac{1}{2}\,|(\vec{x}_3 - \vec{x}_1) \times (\vec{x}_2 - \vec{x}_1)| \label{eq:cross} \\
+  &= \frac{1}{2}\,a\,b\,\sin(C) \label{eq:sine} \\
   &= \sqrt{s\,(s-a)\,(s-b)\,(s-c)} \label{eq:Heron} ~,
@@
 and $s$ is the semi-perimeter $(a + b + c)/2$.
+Below, we present a new formula for the area of a triangle, which we believe may be novel.
 
@@
 only the last of these formulae---Heron's formula \eqref{eq:Heron}---is explicitly permutation-invariant.
+
 The theory of permutation-invariant functions says (CITE) that any per\-mu\-tation-invariant function (under conditions??) of $N$ objects $x_i$ (with $1\leq i\leq N$) can be approximated (to arbitrary precision??) with an expression of the form
@@
 
-\section{A note on our method}
+Although the contribution of this paper is minimal, or even zero, it was inspired by real problems in physics.
+For one very specific example, all known formulae for the gravitational potential energy of a system of $N$ point masses involve sums over pairs of points; they have not ever been written in the Deep Sets form \eqref{eq:DeepSets}.
+In cosmology, all known estimators for $k$-point functions (correlation functions, bispectra, and so on) involve sums over $k$-tuples; once again they have not been written in the Deep Sets form.
+Importantly, these kinds of problems cannot currently be performed (not even approximately) in any map--reduce framework; indeed they are very time-consuming because they involve sums over all elements of $N$-choose-$k$, which scales (badly) as $N^k$.
+Because 3-point functions involve the shapes and areas of triangles, consideration of the 3-point function led to the question answered in this \documentname:
+\emph{What is the Deep Sets (or map--reduce) expression for the area of a triangle?}
+We deliver in \secref{sec:result}.
 
-\section{Result and discussion}
+\section{A note on our method}\label{sec:method}
+What made us think $A^2$ rather than $A$?
 
+What do units tell you?
+
+How to construct all the invariant polynomials.
+
+Which polynomial terms are invariant to other kinds of transformations?
+
+Symbolic regression, determination of the magic number, and checking.
+
+\section{Result and discussion}\label{sec:result}
+Our result---and only contribution---is that the area $A$ of a triangle is given by
 \begin{align}
@@
 -\frac{101}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i^2) \nonumber \\ &
-+\frac{  9}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i)^2 
++\frac{  9}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i)^2 ~, \label{eq:result}
 \end{align}
+where the two-dimensional position of each of the three corners $i$ is $(x_i, y_i)$.
+This answer is in Deep-Sets form:
+Every expression in parentheses in \eqref{eq:result} is a pure sum over all three points.
+There are no sums over differences, and there is no ordering of the points.
+Being very explicit, this can be written in the Deep Sets form \eqref{eq:DeepSets} if we define $\rho$ and $\phi$ as follows:
+
+Note the symmetries....! See the $36/368$ terms, or the $-101/368$ terms
+
+But note that it isn't obviously invariant to arbitrary rotations.
+And it isn't obviously invariant to translations.
+Formulae \eqref{eq:cross}, \eqref{eq:sine}, and \eqref{eq:Heron} are obviously rotation-invariant and translation-invariant.
+
+Why not simplify the prefactors?
+
+What's the generalization to $d$ dimensions?
+
+Where does the factor of 23 come from?
+
+What happens if we try to do $A$ as a quadratic instead of $A^2$ as a quartic?
+
+Why not get this with computer algebra?
 

### 2023-11-09 paper/main.tex
notes to self
small changes

@@
+% Note to self: Please don't work on this, ever.
+
+% to-do items
+% -----------
+% - Show that there is NOT an exact solution at polynomial degree 2 order.
+% - Make the code output something that could be executed by Mathematica.
+% - Prove (maybe using Mathematica) that the formula reduces to Heron's formula.
+% - Prove that the formula is exactly translation-invariant.
+% - Prove that the formula is exactly rotation-invariant.
+% - Related: Needs a clear tikz figure showing the corners and notation.
+% - State exactly the "rules of the game" in the introdution.
+% - What are the 3-dimensional and d-dimensional generalizations of this formula?
+% - Discuss alternatives like Novara's; Thank Novara or invite him to join.
+
 \documentclass[12pt]{article}
@@
 \section*{\raggedright
-Another formula for the area of a triangle: Useless, but explicitly in Deep-Sets and map--reduce form}
+Another formula for the area of a triangle: \\
+Useless, but explicitly in Deep-Sets and map--reduce form}
 \noindent \textbf{David W. Hogg} and \textbf{Soledad Villar}
@@
 \paragraph{Abstract:}
-It is known that any permutation-invariant function of data points $x_i$
-can be written in the form $\rho(\sum_i\phi(x_i))$,
-where $\rho$ and $\phi$ are (probably nonlinear) functions, and the function $\phi$ might output a high-dimensional latent encoding of each point $x_i$.
+It is known that any permutation-invariant function of data points $r_i$
+can be written in the form $\rho(\sum_i\phi(r_i))$,
+where $\rho$ and $\phi$ are (probably nonlinear) functions, and the function $\phi$ might output a high-dimensional latent encoding of each point $r_i$.
 This form---known in the machine-learning literature as Deep Sets---is a map--reduce form, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically or in any order.
 The area of a triangle is a permutation-invariant function of the locations of the three corners.
-We find, possibly for the first time, the formula for the area of the triangle that is explicitly in the Deep-Sets form.
+We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in the Deep-Sets form.
+It is the square-root of a quartic permutation-invariant polynomial.
 No insights of any kind were gained by this work.
@@
 A &= \frac{1}{2}\,\text{(base)}\,\text{(height)} \label{eq:school} \\
-  &= \frac{1}{2}\,|(\vec{x}_3 - \vec{x}_1) \times (\vec{x}_2 - \vec{x}_1)| \label{eq:cross} \\
+  &= \frac{1}{2}\,|(\vec{r}_3 - \vec{r}_1) \times (\vec{r}_2 - \vec{r}_1)| \label{eq:cross} \\
   &= \frac{1}{2}\,a\,b\,\sin(C) \label{eq:sine} \\
@@
 where the first formula is the usual introduction to the subject,
-$(\vec{x}_1, \vec{x}_2, \vec{x}_3$ are three vectors pointing from an origin to the corners of the triangle,
+$\vec{r}_1, \vec{r}_2, \vec{r}_3$ are three vectors pointing from an origin to the corners of the triangle,
 $\times$ represents the cross-product operator (or its generalization to higher dimensions),
@@
 
-The theory of permutation-invariant functions says (CITE) that any per\-mu\-tation-invariant function (under conditions??) of $N$ objects $x_i$ (with $1\leq i\leq N$) can be approximated (to arbitrary precision??) with an expression of the form
+The theory of permutation-invariant functions says (CITE) that any per\-mu\-tation-invariant function (under conditions??) of $N$ objects $r_i$ (with $1\leq i\leq N$) can be approximated (to arbitrary precision??) with an expression of the form
 \begin{equation}
-    \rho(\sum_{i=1}^N\phi(x_i)) \label{eq:DeepSets}
+    \rho(\sum_{i=1}^N\phi(r_i)) \label{eq:DeepSets}
 \end{equation}
 where $\rho$ and $\phi$ are (possibly very nonlinear) functions,
-and the $x_i$ are the data points that can be permuted without changing the result.
+and the $r_i$ are the data points that can be permuted without changing the result.
 This result---which we associate with the subject \emph{Deep Sets} (CITE)---is not at all obvious:
@@
 \begin{itemize}
-    \item The function $\phi$ is universal to all the points $x_i$; it's form won't depend on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
-    \item The point properties $x_i$ and the output of the function $\phi$ might both be very large.
-    That is, each point $x_i$ might be a vector or even an ordered list of vectors and tensors and images and so on, and the output of the function $\phi$ might also be very large and complicated.
-    The output of the function $\phi$ is some kind of informative, latent encoding of the point $x_i$.
+    \item The function $\phi$ is universal to all the points $r_i$; it's form won't depend on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
+    \item The point properties $r_i$ and the output of the function $\phi$ might both be very large.
+    That is, each point $r_i$ might be a vector or even an ordered list of vectors and tensors and images and so on, and the output of the function $\phi$ might also be very large and complicated.
+    The output of the function $\phi$ is some kind of informative, latent encoding of the point $r_i$.
     \item If a function can be written in Deep-Sets form \eqref{eq:DeepSets}, then that function can be executed in a map--reduce framework, in which the function $\phi$ is executed independently and asynchronously on each data point (maybe even in the data center) and the sums are performed hierarchically or in any order.

### 2024-03-19 paper/main.tex
method section

--- 2023-10-29 paper/main.tex
@@
 \section*{\raggedright
-Another formula for the area of a triangle: Useless, but explicitly in Deep-Sets and map--reduce form}
+Another formula for the area of a triangle: Useless, but explicitly in Deep Sets and map--reduce form}
 \noindent \textbf{David W. Hogg} and \textbf{Soledad Villar}
@@
 The area of a triangle is a permutation-invariant function of the locations of the three corners.
-We find, possibly for the first time, the formula for the area of the triangle that is explicitly in the Deep-Sets form.
+We find, possibly for the first time, the formula for the area of the triangle that is explicitly in the Deep Sets form.
 No insights of any kind were gained by this work.
@@
 
-A few comments about this Deep-Sets form \eqref{eq:DeepSets} we'd make are the following:
+A few comments about this Deep Sets form \eqref{eq:DeepSets} we'd make are the following:
 \begin{itemize}
@@
     The output of the function $\phi$ is some kind of informative, latent encoding of the point $x_i$.
-    \item If a function can be written in Deep-Sets form \eqref{eq:DeepSets}, then that function can be executed in a map--reduce framework, in which the function $\phi$ is executed independently and asynchronously on each data point (maybe even in the data center) and the sums are performed hierarchically or in any order.
+    \item If a function can be written in Deep Sets form \eqref{eq:DeepSets}, then that function can be executed in a map--reduce framework, in which the function $\phi$ is executed independently and asynchronously on each data point (maybe even in the data center) and the sums are performed hierarchically or in any order.
     That means that for large problems executed on large data sets in large data centers, problems in this form \eqref{eq:DeepSets} can be performed in wall-clock time that scales as $\log N$.
@@
 \section{A note on our method}\label{sec:method}
-What made us think $A^2$ rather than $A$?
+Fundamentally, our method takes the form of a symbolic regression, followed by a confirmation with computational algebra.
 
-What do units tell you?
+For the symbolic regression, we considered a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
+We fit a linear combination of the monomials to a data set of generated triples of points, drawn from a two-dimensional Gaussian whose mean was also drawn from a Gaussian in a two-dimensional space.
+For each triple we attach a label, which is the squared area, and we require the linear combination of fourth-order, Deep Sets monomials to reproduce the labels to machine precision.
 
-How to construct all the invariant polynomials.
+We choose to work with squared area $A^2$ and not area $A$ because, geometrically, areas are signed quantities.
+Indeed, the formulae \eqref{eq:school} through \eqref{eq:Heron} either (implicitly) involve square roots of polynomials, or else absolute value signs, such that the corresponding pure polynomial expression would be for the squared area.
+We restricted to monomials that contain four powers of positional coordinates because any expression for the squared area must be units covariant (equivariant to changes of length units).
 
-Which polynomial terms are invariant to other kinds of transformations?
+Perhaps non-obviously, for a monomial to be in Deep Sets form, it must consist only of products of sums over the three points.
+It cannot contain any differences, or any sums over pairs of points (or edges).
+Thus $(\sum_{i=1}^2 x_i^2)\,(\sum_{i=1}^3 x_i\,y_i)$ is a permitted monomial, but nothing of the form $(\sum_{i=1}^3\sum_{j=1}^3 x_i^2\,y_j^2)$ is not permitted.
+These monomials are both correct from a units perspective, but the second is not in Deep Sets form, because it contains products of position coordinates from different points.
 
-Symbolic regression, determination of the magic number, and checking.
+HOGG: How to construct all the invariant polynomials.
+
+In detail, the formula for the area (or area squared) of a triangle must be invariant to permutations, translations, rotations, and reflections.
+The Deep Sets form enforces only permutation-invariance.
+Some of the monomials are invariant to reflections, but no individual monomial is invariant to translations or rotations.
+It is left as an exercise to the reader to show that our solution \eqref{eq:result} below is invariant to translations, rotations, and reflections.
+
+The symbolic regression produces a linear combination of monomials with real-valued coefficients.
+Trial-and-error delivered the integer 368, which makes all of the coefficients obviously rational.
+We have no theory of this number; we leave it as a second exercise to the reader to explain this large, non-prime integer.
+
+Finally, the expression found is output by the code in LaTeX form, to create the typesetting commands for \eqref{eq:result} and also output by the code in Mathematica form, to permit computational algebra operations.
+A Mathematica \texttt{simplify} operation [HOGG NO NOT REALLY] is used to show that the Deep Sets expression \eqref{eq:result} is identical to the classic formulae for a triangle's squared area $A^2$.
 
@@
 where the two-dimensional position of each of the three corners $i$ is $(x_i, y_i)$.
-This answer is in Deep-Sets form:
+This answer is in Deep Sets form:
 Every expression in parentheses in \eqref{eq:result} is a pure sum over all three points.
@@
 Being very explicit, this can be written in the Deep Sets form \eqref{eq:DeepSets} if we define $\rho$ and $\phi$ as follows:
+
+HOGG: Explicit vector $\phi()$ and explicit expression $\rho()$.
 

### 2024-03-20 paper/main.tex

--- 2023-11-09 paper/main.tex
@@
 \section*{\raggedright
-Another formula for the area of a triangle: \\
-Useless, but explicitly in Deep-Sets and map--reduce form}
+Another formula for the area of a triangle: Useless, but explicitly in Deep Sets and map--reduce form}
+
 \noindent \textbf{David W. Hogg} and \textbf{Soledad Villar}
@@
 The area of a triangle is a permutation-invariant function of the locations of the three corners.
-We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in the Deep-Sets form.
+We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in the Deep Sets form.
 It is the square-root of a quartic permutation-invariant polynomial.
@@
 
-A few comments about this Deep-Sets form \eqref{eq:DeepSets} we'd make are the following:
+A few comments about this Deep Sets form \eqref{eq:DeepSets} we'd make are the following:
 \begin{itemize}
-    \item The function $\phi$ is universal to all the points $r_i$; it's form won't depend on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
+    \item The function $\phi$ is universal to all the points $r_i$; its form won't depend on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
     \item The point properties $r_i$ and the output of the function $\phi$ might both be very large.

### 2024-03-20 paper/main.tex
Updates from Overleaf

--- 2023-11-09 paper/main.tex
@@
 \begin{itemize}
-    \item The function $\phi$ is universal to all the points $r_i$; it's form won't depend on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
+    \item The function $\phi$ is universal to all the points $r_i$; its form won't depend on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
     \item The point properties $r_i$ and the output of the function $\phi$ might both be very large.
@@
     \item If a function can be written in Deep-Sets form \eqref{eq:DeepSets}, then that function can be executed in a map--reduce framework, in which the function $\phi$ is executed independently and asynchronously on each data point (maybe even in the data center) and the sums are performed hierarchically or in any order.
-    That means that for large problems executed on large data sets in large data centers, problems in this form \eqref{eq:DeepSets} can be performed in wall-clock time that scales as $\log N$.
+    That means that for large problems executed on large data sets in large data centers, problems in this form \eqref{eq:DeepSets} can often be performed in wall-clock time that scales as $\log N$.
 \end{itemize}
@@
 \end{align}
-where the two-dimensional position of each of the three corners $i$ is $(x_i, y_i)$.
+where the two-dimensional position of each of the three corners $i$ is $\vec{r}_i = (x_i, y_i)$.
 This answer is in Deep-Sets form:

### 2024-03-21 paper/main.tex
affils

--- 2024-03-20 paper/main.tex
@@
 \section*{\raggedright
-Another formula for the area of a triangle: Useless, but explicitly in Deep Sets and map--reduce form}
+A new formula for the area of a triangle:
+Useless, but explicitly in Deep Sets and map--reduce form}
 
-\noindent \textbf{David W. Hogg} and \textbf{Soledad Villar}
+\noindent \textbf{David W.
+Hogg}\footnote{Email at \texttt{<david.hogg@nyu.edu>}.
+DWH is at the Center for Cosmology and Particle Physics, Department of Physics, New York University; the Max-Planck-Insitut f\"ur Astronomie; and the Center for Computational Astrophysics, Flatiron Institute.
+The Flatiron Institute is a division of the Simons Foundation.} and
+\textbf{Soledad Villar}\footnote{SV is at the Department of Applied Mathematics and Statistics, Johns Hopkins University; and the Mathematical Institute for Data Science, Johns Hopkins University.}
 

### 2024-03-21 paper/main.tex
Front matter and to-do lists.

@@
 % - Show that there is NOT an exact solution at polynomial degree 2 order.
-% - Make the code output something that could be executed by Mathematica.
-% - Prove (maybe using Mathematica) that the formula reduces to Heron's formula.
-% - Prove that the formula is exactly translation-invariant.
-% - Prove that the formula is exactly rotation-invariant.
-% - Related: Needs a clear tikz figure showing the corners and notation.
-% - State exactly the "rules of the game" in the introdution.
-% - What are the 3-dimensional and d-dimensional generalizations of this formula?
-% - Discuss alternatives like Novara's; Thank Novara or invite him to join.
+% - Make the code output something that could be executed by Mathematica. Then prove using Mathematica that the formula reduces to Heron's formula.
+% - Make the code output explicit expressions for phi and rho, and include them in section 3.
+% - Needs a clear tikz figure showing the corners and notation.
+% - What are the 3-dimensional and d-dimensional generalizations of this formula? Are they hard to find? YES I BET THEY ARE.
+% - Discuss alternatives like Novara's; And alternatives like Kaze's.
 
@@
 \noindent \textbf{David W.
-Hogg}\footnote{Email at \texttt{<david.hogg@nyu.edu>}.
+Hogg}\footnote{The authors thank Alexander Novara (NYU), Ben Wandelt (Sorbonne), and Kaze Wong (Flatiron) for valuable discussions.}\footnote{Email at \texttt{<david.hogg@nyu.edu>}.
 DWH is at the Center for Cosmology and Particle Physics, Department of Physics, New York University; the Max-Planck-Insitut f\"ur Astronomie; and the Center for Computational Astrophysics, Flatiron Institute.
@@
 This form---known in the machine-learning literature as Deep Sets---is a map--reduce form, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically or in any order.
-The area of a triangle is a permutation-invariant function of the locations of the three corners.
-We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in the Deep Sets form.
-It is the square-root of a quartic permutation-invariant polynomial.
+The area of a triangle is a permutation-invariant function of the locations $r_i$ of the three corners $1\leq i\leq 3$.
+We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in Deep Sets form.
+It is the square-root of a permutation-invariant quartic polynomial.
 No insights of any kind were gained by this work.
@@
 \section{Introduction}
-The area $A$ of a triangle can be written in many ways:
+The area $A$ of a triangle can be written in many ways.
+Here are a few:
 \begin{align}
@@
 \end{align}
-where the first formula is the usual introduction to the subject,
+where the first formula \eqref{eq:school} is the usual introduction to the subject,
 $\vec{r}_1, \vec{r}_2, \vec{r}_3$ are three vectors pointing from an origin to the corners of the triangle,
@@
 \begin{itemize}
-    \item The function $\phi$ is universal to all the points $r_i$; its form won't depend on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
+    \item The function $\phi$ is universal to all the points $r_i$; its form won't depend on label $i$ or on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
     \item The point properties $r_i$ and the output of the function $\phi$ might both be very large.
@@
 \section{A note on our method}\label{sec:method}
-Fundamentally, our method takes the form of a symbolic regression, followed by a confirmation with computational algebra.
+How did we find the formula \eqref{eq:result} presented below in \secref{sec:result}?
+Fundamentally, our method took the form of a symbolic regression, followed by a confirmation with computational algebra.
 
@@
 \end{align}
-where the two-dimensional position of each of the three corners $i$ is $(x_i, y_i)$.
+where the two-dimensional position of each of the three corners $i$ is $r_i \equiv (x_i, y_i)$.
 This answer is in Deep Sets form:

### 2024-03-21 paper/main.tex
typesetting and notes to self

@@
 \usepackage[letterpaper]{geometry}
+\usepackage[hidelinks]{hyperref}
 \usepackage{amsmath}
 
+% typesetting issues
 \newcommand{\documentname}{\textsl{Note}}
@@
 \newcommand{\secref}[1]{\sectionname~\ref{#1}}
-\frenchspacing
+\addtolength{\topmargin}{-0.20in}
+\addtolength{\textheight}{1.40in}
+\sloppy\sloppypar\raggedbottom\frenchspacing
+\pagestyle{myheadings}
+\markboth{foo}{Hogg \& Villar / A new formula for the area of a triangle}
 
-\begin{document}
+\begin{document}\thispagestyle{empty}
 
@@
 For each triple we attach a label, which is the squared area, and we require the linear combination of fourth-order, Deep Sets monomials to reproduce the labels to machine precision.
+The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
 
@@
 Indeed, the formulae \eqref{eq:school} through \eqref{eq:Heron} either (implicitly) involve square roots of polynomials, or else absolute value signs, such that the corresponding pure polynomial expression would be for the squared area.
-We restricted to monomials that contain four powers of positional coordinates because any expression for the squared area must be units covariant (equivariant to changes of length units).
+We restricted to monomials that contain four powers of positional coordinates because any expression for the squared area must be units covariant (equivariant to changes of the units used for the measurements of positions or lengths).
 
@@
 
-Why not get this with computer algebra?
+What happens if you hold the formula fixed, but use more than $N=3$ points?
+The formula still returns a squared area.
+That's the area of what?
 

### 2024-03-21 paper/main.tex
typo

@@
 It cannot contain any differences, or any sums over pairs of points (or edges).
-Thus $(\sum_{i=1}^2 x_i^2)\,(\sum_{i=1}^3 x_i\,y_i)$ is a permitted monomial, but nothing of the form $(\sum_{i=1}^3\sum_{j=1}^3 x_i^2\,y_j^2)$ is not permitted.
+Thus $(\sum_{i=1}^2 x_i^2)\,(\sum_{i=1}^3 x_i\,y_i)$ is a permitted monomial, but $(\sum_{i=1}^3\sum_{j=1}^3 x_i^2\,y_j^2)$ is not a permitted monomial.
 These monomials are both correct from a units perspective, but the second is not in Deep Sets form, because it contains products of position coordinates from different points.

### 2025-01-23 paper/main.tex
Updates from Overleaf

@@
 We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in Deep Sets form.
-It is the square-root of a permutation-invariant quartic polynomial.
-No insights of any kind were gained by this work.
+It is the square-root of a permutation-invariant quartic polynomial involving the integer 368.
+No insights of any kind were gained from this work.
 
@@
 
+HAINJE: Given data of form $(x, y)$, invariant factors are $F_{m,n} = \sum_i x_i^m y_i^n$, and the invariant monomials are any products of these factors $\prod_{i} F_{m_i,n_i}$.
+Do we want/need to be more specific than that?
+
 In detail, the formula for the area (or area squared) of a triangle must be invariant to permutations, translations, rotations, and reflections.
@@
 
+HAINJE: Here it is:
+\begin{gather}
+    \phi(r_i) = (x_i, y_i, x_iy_i, x_i^2, y_i^2, x_i y_i^2, x_i^2 y_i, x_i^2 y_i^2) \\
+    \begin{aligned}
+    \rho(\textstyle\sum_{i=1}^{3}\phi(r_i)\equiv \vec{u}) = &
+    - \tfrac{54}{368} u_8
+    - \tfrac{258}{368} u_3^2
+    + \tfrac{285}{368} u_4 u_5
+    + \tfrac{36}{368} u_2 u_7
+    + \tfrac{36}{368} u_1 u_6 \\ &
+    - \tfrac{101}{368} u_2^2 u_4
+    + \tfrac{148}{368} u_1 u_2 u_3
+    - \tfrac{101}{368} u_1^2 u_5
+    + \tfrac{9}{368} u_1^2 u_2^2.
+    \end{aligned}
+\end{gather}
+
 Note the symmetries....! See the $36/368$ terms, or the $-101/368$ terms
@@
 
+\newpage
+By inspection, we have also found a simpler form of the same formula, which is given by
+\begin{align}
+    4 \, A^2 = & \
+        2 \, (\sum_{i=1}^{3} x_i ) \, (\sum_{i=1}^{3} y_i ) \, (\sum_{i=1}^{3} x_i \, y_i )
+        + 3 \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i^2 ) \nonumber \\ &
+        - \, (\sum_{i=1}^{3} x_i )^2 \, (\sum_{i=1}^{3} y_i^2 )
+        - \, (\sum_{i=1}^{3} y_i )^2 \, (\sum_{i=1}^{3} x_i^2 ) \nonumber \\ &
+        - 3 \, (\sum_{i=1}^{3} x_i \, y_i )^2.
+\end{align}
+In explicit Deep Sets notation, this can be written
+\begin{gather}
+    \phi(r_i) = (
+        x_i,
+        y_i,
+        x_i y_i,
+        x_i^2,
+        y_i^2
+    ), \\
+    \rho(\textstyle\sum_{i=1}^{3} \phi(r_i) \equiv \vec{u})
+    = \tfrac{2}{4} u_1 u_2 u_3
+    + \tfrac{3}{4} u_4 u_5
+    - \tfrac{1}{4} u_1^2 u_5
+    - \tfrac{1}{4} u_2^2 u_4
+    - \tfrac{3}{4} u_3^2.
+\end{gather}
+
+The two are related by
+\begin{equation}
+    4 u_1 u_6 + 4 u_2 u_7 - 6 u_8 = 
+    -u_1^2 u_2^2 + 4 u_1 u_2 u_3 - 2 u_3^2 + u_2^2 u_4 + u_1^2 u_5 - u_4 u_5 ;
+\end{equation}
+namely, plugging this into \eqref{eq:result} causes it to reduce immediately into this simpler form.
+
+This form was not found by the symbolic regression of \secref{sec:method} because the regression makes use of \verb+lstsq+ with more parameters than data points.
+In this regime, the solution that minimizes the $\mathcal{L}_2$ norm of the weights (the coefficients of the invariant monomials) is returned.
+The $\mathcal{L}_2$ norm of the weights of our initial formula is $2127 / 1472 \approx 1.445$, whereas the $\mathcal{L}_2$ norm of the weights of this simpler formula is $3/2 = 1.5$.
+
+HAINJE: Would lasso regression return my formula? Is there a yet simpler formula?
+
 \end{document}

### 2025-03-01 paper/main.tex
Hainje formula; to-do items for us.

--- 2024-03-21 paper/main.tex
@@
-% Note to self: Please don't work on this, ever.
+% Copyright 2025 the authors.
 
@@
 % -----------
-% - Show that there is NOT an exact solution at polynomial degree 2 order.
-% - Make the code output something that could be executed by Mathematica. Then prove using Mathematica that the formula reduces to Heron's formula.
-% - Make the code output explicit expressions for phi and rho, and include them in section 3.
-% - Needs a clear tikz figure showing the corners and notation.
-% - What are the 3-dimensional and d-dimensional generalizations of this formula? Are they hard to find? YES I BET THEY ARE.
-% - Discuss alternatives like Novara's; And alternatives like Kaze's.
+% - CH: Include the d=3, d=4 solutions and comments on d>4.
+% - CH: Rewrite to remove the Hogg solution.
+% - CH: Show that there is NOT an exact solution at polynomial degree 2 order.
+% - CH: Needs a clear tikz figure showing the corners and notation. Like an infographic.
+% - HOGG: Discuss alternatives like Novara's; And alternatives like Kaze's.
+Positional encodings!
+CH: Describe the connection to positional encodings (in, eg, transformers).
+Oh and cosmology and power spectrum and bispectrum.
 
@@
 \newcommand{\secref}[1]{\sectionname~\ref{#1}}
+\setlength{\textwidth}{5.50in}
+\setlength{\oddsidemargin}{3.25in}\addtolength{\oddsidemargin}{-0.5\textwidth}
 \addtolength{\topmargin}{-0.20in}
@@
 \pagestyle{myheadings}
-\markboth{foo}{Hogg \& Villar / A new formula for the area of a triangle}
+\markboth{foo}{\sffamily Hainje, Hogg, \& Villar / A new formula for the area of a triangle}
+
+\usepackage{natbib}
+\usepackage{aas_macros}
+\bibliographystyle{aasjournal}
+\setcitestyle{aysep={}}
+\usepackage{setspace}
+\setlength{\bibsep}{0pt plus 0.3ex}
 
@@
 A new formula for the area of a triangle:
-Useless, but explicitly in Deep Sets and map--reduce form}
+Useless, but explicitly in Deep Sets and map--reduce form\footnote{%
+The authors thank Alexander Novara (NYU), Ben Wandelt (Sorbonne), and Kaze Wong (Flatiron) for valuable discussions.}}
 
+\noindent \textbf{Connor Hainje},\footnote{Email at \href{mailto:connor.hainje@nyu.edu}{connor.hainje@nyu.edu}.
+CH is at the Center for Cosmology and Particle Physics, Department of Physics, New York University.}
 \noindent \textbf{David W.
-Hogg}\footnote{The authors thank Alexander Novara (NYU), Ben Wandelt (Sorbonne), and Kaze Wong (Flatiron) for valuable discussions.}\footnote{Email at \texttt{<david.hogg@nyu.edu>}.
-DWH is at the Center for Cosmology and Particle Physics, Department of Physics, New York University; the Max-Planck-Insitut f\"ur Astronomie; and the Center for Computational Astrophysics, Flatiron Institute.
+Hogg},\footnote{DWH is at the Center for Cosmology and Particle Physics, Department of Physics, New York University; the Max-Planck-Insitut f\"ur Astronomie; and the Center for Computational Astrophysics, Flatiron Institute.
 The Flatiron Institute is a division of the Simons Foundation.} and
@@
 We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in Deep Sets form.
-It is the square-root of a permutation-invariant quartic polynomial.
-No insights of any kind were gained by this work.
+% It is the square-root of a permutation-invariant quartic polynomial involving the integer 368.
+No insights of any kind were gained from this work.
 
@@
 
-The theory of permutation-invariant functions says (CITE) that any per\-mu\-tation-invariant function (under conditions??) of $N$ objects $r_i$ (with $1\leq i\leq N$) can be approximated (to arbitrary precision??) with an expression of the form
+The theory of permutation-invariant functions says that any per\-mu\-tation-invariant function (under conditions??) of $N$ objects $r_i$ (with $1\leq i\leq N$) can be approximated (to arbitrary precision??) with an expression of the form \citep{zaheer+17-Deep-Sets}
 \begin{equation}
@@
 Although the contribution of this paper is minimal, or even zero, it was inspired by real problems in physics.
-For one very specific example, all known formulae for the gravitational potential energy of a system of $N$ point masses involve sums over pairs of points; they have not ever been written in the Deep Sets form \eqref{eq:DeepSets}.
-In cosmology, all known estimators for $k$-point functions (correlation functions, bispectra, and so on) involve sums over $k$-tuples; once again they have not been written in the Deep Sets form.
-Importantly, these kinds of problems cannot currently be performed (not even approximately) in any map--reduce framework; indeed they are very time-consuming because they involve sums over all elements of $N$-choose-$k$, which scales (badly) as $N^k$.
+For one very specific example, all known exact formulae for the gravitational potential energy of a system of $N$ point masses involve sums over pairs of points (and also require taking differences of points); they have not ever been written in Deep Sets form \eqref{eq:DeepSets} to our knowledge.
+CH: THAT SAID, THERE ARE MULTIPOLE-LIKE EXPRESSIONS THAT MIGHT BE IN DEEPSETS AND APPROXIMATE THIS CALCULATION; SEE THE DISCUSSION?
+In cosmology, standard estimators for $k$-point functions (correlation functions, bispectra, and so on) involve sums over $k$-tuples; once again they have not been written in the Deep Sets form.
+CH: DO THE PHILCOX PAPERS ADDRESS THIS PREVIOUS POINT?
+Importantly, when these statistics are computed in their naive forms, they are very time-consuming because they involve sums over all elements of $N$-choose-$k$, which scales (badly) as $N^k$.
 Because 3-point functions involve the shapes and areas of triangles, consideration of the 3-point function led to the question answered in this \documentname:
@@
 
-HOGG: How to construct all the invariant polynomials.
+In order to generate all the possible invariant monomials, we first note that they are made of products of factors of the form $F_{m,n} = \sum_i x_i^m y_i^n$.
+The invariant monomials of degree $d$ are then all products $\prod_i F_{m_i, n_i}$ such that $\sum_i m_i + n_i = d$.
+
+% HOGG: How to construct all the invariant polynomials.
+
+% HAINJE: Given data of form $(x, y)$, invariant factors are $F_{m,n} = \sum_i x_i^m y_i^n$, and the invariant monomials are any products of these factors $\prod_{i} F_{m_i,n_i}$.
+Do we want/need to be more specific than that?
 
@@
 
+HAINJE: Here it is:
+\begin{gather}
+    \phi(r_i) = (x_i, y_i, x_i\,y_i, x_i^2, y_i^2, x_i\,y_i^2, x_i^2\,y_i, x_i^2\,y_i^2) \\
+    \begin{aligned}
+    \rho(\textstyle\sum_{i=1}^{3}\phi(r_i)\equiv \vec{u}) = &
+    - \tfrac{54}{368} u_8
+    - \tfrac{258}{368} u_3^2
+    + \tfrac{285}{368} u_4\,u_5
+    + \tfrac{36}{368} u_2\,u_7
+    + \tfrac{36}{368} u_1\,u_6 \\ &
+    - \tfrac{101}{368} u_2^2\,u_4
+    + \tfrac{148}{368} u_1\,u_2\,u_3
+    - \tfrac{101}{368} u_1^2\,u_5
+    + \tfrac{9}{368} u_1^2\,u_2^2.
+    \end{aligned}
+\end{gather}
+
 Note the symmetries....! See the $36/368$ terms, or the $-101/368$ terms
@@
 
-Why not simplify the prefactors?
-
-What's the generalization to $d$ dimensions?
-
-Where does the factor of 23 come from?
-
-What happens if we try to do $A$ as a quadratic instead of $A^2$ as a quartic?
-
-What happens if you hold the formula fixed, but use more than $N=3$ points?
+HOGG SAY: CH: EXPAND THIS INTO AN OPEN QUESTIONS LIST:
+\begin{itemize}
+\item CH should answer this question: What's the generalization to $d$ dimensions? That is, what if the three points are each in d dimensions?
+\item questionable: What happens if we try to do $A$ as a quadratic instead of $A^2$ as a quartic?
+\item What happens if you hold the formula fixed, but use more than $N=3$ points?
 The formula still returns a squared area.
 That's the area of what?
+\item What about areas of quadrangles or larger polygons?
+\item What about volumes and hypervolumes of larger objects?
+\end{itemize}
+
+\newpage
+By inspection, we have also found a simpler form of the same formula, which is given by
+\begin{align}
+    4 \, A^2 = & \
+        2 \, (\sum_{i=1}^{3} x_i ) \, (\sum_{i=1}^{3} y_i ) \, (\sum_{i=1}^{3} x_i \, y_i )
+        \nonumber \\ &
+        + 3 \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i^2 )
+        - 3 \, (\sum_{i=1}^{3} x_i \, y_i )^2.
+        \nonumber \\ &
+        - \, (\sum_{i=1}^{3} x_i )^2 \, (\sum_{i=1}^{3} y_i^2 )
+        - \, (\sum_{i=1}^{3} y_i )^2 \, (\sum_{i=1}^{3} x_i^2 )
+\end{align}
+In explicit Deep Sets notation, this can be written
+\begin{align}
+    \phi(r_i) &= \begin{bmatrix}
+        x_i & y_i & x_i\,y_i & x_i^2 & y_i^2 \end{bmatrix}^\top
+        \equiv \vec{u}^\top, \\
+    \rho(\sum_{i=1}^{3} \phi(r_i))
+    &= \tfrac{2}{4} u_1\,u_2\,u_3
+    + \tfrac{3}{4} u_4\,u_5
+    - \tfrac{1}{4} u_1^2\,u_5
+    - \tfrac{1}{4} u_2^2\,u_4
+    - \tfrac{3}{4} u_3^2.
+\end{align}
+
+The two are related by
+\begin{equation}
+    4\,u_1\,u_6 + 4\,u_2\,u_7 - 6\,u_8 = 
+    -u_1^2 u_2^2 + 4 u_1 u_2 u_3 - 2 u_3^2 + u_2^2 u_4 + u_1^2 u_5 - u_4 u_5 ;
+\end{equation}
+namely, plugging this into \eqref{eq:result} causes it to reduce immediately into this simpler form.
+
+This form was not found by the symbolic regression of \secref{sec:method} because the regression makes use of \verb+lstsq+ with more parameters than data points.
+In this regime, the solution that minimizes the $\mathcal{L}_2$ norm of the weights (the coefficients of the invariant monomials) is returned.
+The $\mathcal{L}_2$ norm of the weights of our initial formula is $2127 / 1472 \approx 1.445$, whereas the $\mathcal{L}_2$ norm of the weights of this simpler formula is $3/2 = 1.5$.
+
+HAINJE: Would lasso regression return my formula? Is there a yet simpler formula?
+
+\newpage
+
+Define the ``invariant factor'' in $d$ dimensions
+\begin{equation}
+    F_{m_1 \, m_2 \, \cdots} \equiv \sum_{1=1}^{3} \prod_{k=1}^{d} (\vec{r}_{i})_{k}^{m_k}.
+\end{equation}
+This is the sum over all three particles of the product of the $d$ components of each $\vec{r}_i$, each component raised to a power. For example, in 2D, $d=2$ and this expands to
+\begin{equation}
+    F_{\ell m} = \sum_{i=1}^{3} x_i^\ell \, y_i^m.
+\end{equation}
+In 3D, $d=3$ and this gives
+\begin{equation}
+    F_{\ell m n} = \sum_{i=1}^{3} x_i^\ell \, y_i^m \, z_i^n.
+\end{equation}
+Then, the area formula found above can be extended to more dimensions with the following formulae.
+
+For reference, the 2D formula was
+\begin{equation}
+4 A_{\rm 2D}^2 = +3 F_{02} F_{20} -3 F_{11}^2 - F_{20} F_{01}^2 +2 F_{11} F_{10} F_{01} - F_{02} F_{10}^2.
+\end{equation}
+In 3D, we have
+\begin{equation}
+\begin{aligned}
+4 A_{\rm 3D}^2 = &
++3 F_{020} F_{002} 
++3 F_{200} F_{002} 
++3 F_{020} F_{200} \\&
+-3 F_{011}^2 
+-3 F_{101}^2 
+-3 F_{110}^2 \\&
++2 F_{011} F_{001} F_{010} 
++2 F_{100} F_{001} F_{101}
++2 F_{100} F_{010} F_{110} \\&
+-F_{002} F_{010}^2 
+-F_{002} F_{100}^2 
+-F_{020} F_{001}^2 
+-F_{020} F_{100}^2
+-F_{200} F_{010}^2
+-F_{200} F_{001}^2. 
+\end{aligned}
+\end{equation}
+And in 4D, we have
+\begin{equation}
+\begin{aligned}
+4 A_{\rm 4D}^2 = &
++3 F_{0002} F_{0020} 
++3 F_{0002} F_{0200} 
++3 F_{0002} F_{2000} 
+\\&
++3 F_{0200} F_{2000} 
++3 F_{0200} F_{0020} 
++3 F_{0020} F_{2000} 
+\\&
+-3 F_{0011}^2
+-3 F_{0101}^2
+-3 F_{0110}^2
+-3 F_{1001}^2 
+-3 F_{1010}^2 
+-3 F_{1100}^2
+\\&
++2 F_{0010} F_{0011} F_{0001} 
++2 F_{0101} F_{0100} F_{0001} 
++2 F_{1001} F_{1000} F_{0001}
+\\&
++2 F_{0010} F_{0100} F_{0110} 
++2 F_{0010} F_{1010} F_{1000}
++2 F_{1100} F_{0100} F_{1000} 
+\\&
+- F_{0020} F_{0001}^2
+- F_{0200} F_{0001}^2
+- F_{2000} F_{0001}^2
+- F_{0002} F_{0010}^2
+- F_{0200} F_{0010}^2
+- F_{2000} F_{0010}^2
+\\&
+- F_{0002} F_{0100}^2 
+- F_{0020} F_{0100}^2 
+- F_{2000} F_{0100}^2
+- F_{0002} F_{1000}^2 
+- F_{0020} F_{1000}^2 
+- F_{0200} F_{1000}^2
+\end{aligned}
+\end{equation}
+
+\newpage
+{\small\singlespacing\bibliography{main}}
 

### 2025-03-02 paper/main.tex
tweak

@@
 Useless, but explicitly in Deep Sets and map--reduce form\footnote{%
-The authors thank Alexander Novara (NYU), Ben Wandelt (Sorbonne), and Kaze Wong (Flatiron) for valuable discussions.}}
+The authors thank Alexander Novara (NYU), Ben Wandelt (Sorbonne), and Kaze Wong (JHU) for valuable discussions.}}
 

### 2025-03-03 paper/main.tex
hypervolumes

@@
 \usepackage[hidelinks]{hyperref}
-\usepackage{amsmath}
+\usepackage{amsmath,amsfonts}
 
@@
 
-The theory of permutation-invariant functions says that any per\-mu\-tation-invariant function (under conditions??) of $N$ objects $r_i$ (with $1\leq i\leq N$) can be approximated (to arbitrary precision??) with an expression of the form \citep{zaheer+17-Deep-Sets}
+Theorem 2 of \citet{zaheer+17-Deep-Sets} states that a function $f(X)$ operating on a set $X$ is invariant to the permutation of instances in $X$ if and only if it can be decomposed in the form
 \begin{equation}
-    \rho(\sum_{i=1}^N\phi(r_i)) \label{eq:DeepSets}
+    \textstyle f(X) = \rho \left( \sum_{x \in X} \phi(x) \right), \label{eq:DeepSets}
 \end{equation}
-where $\rho$ and $\phi$ are (possibly very nonlinear) functions,
-and the $r_i$ are the data points that can be permuted without changing the result.
-This result---which we associate with the subject \emph{Deep Sets} (CITE)---is not at all obvious:
+for some functions $\rho$ and $\phi$.
+This is proven for the cases where the elements of $X$ are from a countable universe and the size of $X$ is unlimited, and for the case where the elements of $X$ are from an uncountable universe but the size of $X$ is finite.
+This result---often known by the name of the machine learning architecture derived from it, \emph{Deep Sets} \citep{zaheer+17-Deep-Sets}---is not at all obvious:
 For example, even though Heron's formula \eqref{eq:Heron} is permutation-invariant, it is not actually in the form \eqref{eq:DeepSets}, if the triangle is seen as being defined by its corners (or even its sides).
@@
 \begin{itemize}
-    \item The function $\phi$ is universal to all the points $r_i$; its form won't depend on label $i$ or on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
+    \item The function $\phi$ is universal to all the elements $x$; its form won't depend on label $i$ or on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
     \item The point properties $r_i$ and the output of the function $\phi$ might both be very large.
@@
 These monomials are both correct from a units perspective, but the second is not in Deep Sets form, because it contains products of position coordinates from different points. 
-
 In order to generate all the possible invariant monomials, we first note that they are made of products of factors of the form $F_{m,n} = \sum_i x_i^m y_i^n$.
@@
 The symbolic regression produces a linear combination of monomials with real-valued coefficients.
-Trial-and-error delivered the integer 368, which makes all of the coefficients obviously rational.
-We have no theory of this number; we leave it as a second exercise to the reader to explain this large, non-prime integer.
+We systematically search for an integer which, when multiplied by all of the coefficients, brings all coefficients very near integers.
+We then round the coefficients to these nearest integers and check that we reproduce $({\rm denom}) A^2$ to better precision than the least-square result.
+% Trial-and-error delivered the integer 368, which makes all of the coefficients obviously rational.
+% We have no theory of this number; we leave it as a second exercise to the reader to explain this large, non-prime integer.
 
-Finally, the expression found is output by the code in LaTeX form, to create the typesetting commands for \eqref{eq:result} and also output by the code in Mathematica form, to permit computational algebra operations.
-A Mathematica \texttt{simplify} operation [HOGG NO NOT REALLY] is used to show that the Deep Sets expression \eqref{eq:result} is identical to the classic formulae for a triangle's squared area $A^2$.
+% Finally, the expression found is output by the code in LaTeX form, to create the typesetting commands for \eqref{eq:result} and also output by the code in Mathematica form, to permit computational algebra operations.
+% A Mathematica \texttt{simplify} operation [HOGG NO NOT REALLY] is used to show that the Deep Sets expression \eqref{eq:result} is identical to the classic formulae for a triangle's squared area $A^2$.
 
 \section{Result and discussion}\label{sec:result}
+
 Our result---and only contribution---is that the area $A$ of a triangle is given by
+% \begin{align}
+% A^2 = &
+% -\frac{ 54}{368}\,(\sum_{i=1}^3 x_i^2\,y_i^2) 
+% -\frac{258}{368}\,(\sum_{i=1}^3 x_i\,y_i)^2 \nonumber \\ &
+% +\frac{285}{368}\,(\sum_{i=1}^3 x_i^2)\,(\sum_{i=1}^3 y_i^2) 
+% +\frac{ 36}{368}\,(\sum_{i=1}^3 y_i)\,(\sum_{i=1}^3 x_i^2\,y_i) \nonumber \\ &
+% +\frac{ 36}{368}\,(\sum_{i=1}^3 x_i)\,(\sum_{i=1}^3 x_i\,y_i^2) 
+% -\frac{101}{368}\,(\sum_{i=1}^3 y_i)^2\,(\sum_{i=1}^3 x_i^2) \nonumber \\ &
+% +\frac{148}{368}\,(\sum_{i=1}^3 x_i)\,(\sum_{i=1}^3 y_i)\,(\sum_{i=1}^3 x_i\,y_i) 
+% -\frac{101}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i^2) \nonumber \\ &
+% +\frac{  9}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i)^2 ~, \label{eq:result}
+% \end{align}
 \begin{align}
-A^2 = &
--\frac{ 54}{368}\,(\sum_{i=1}^3 x_i^2\,y_i^2) 
--\frac{258}{368}\,(\sum_{i=1}^3 x_i\,y_i)^2 \nonumber \\ &
-+\frac{285}{368}\,(\sum_{i=1}^3 x_i^2)\,(\sum_{i=1}^3 y_i^2) 
-+\frac{ 36}{368}\,(\sum_{i=1}^3 y_i)\,(\sum_{i=1}^3 x_i^2\,y_i) \nonumber \\ &
-+\frac{ 36}{368}\,(\sum_{i=1}^3 x_i)\,(\sum_{i=1}^3 x_i\,y_i^2) 
--\frac{101}{368}\,(\sum_{i=1}^3 y_i)^2\,(\sum_{i=1}^3 x_i^2) \nonumber \\ &
-+\frac{148}{368}\,(\sum_{i=1}^3 x_i)\,(\sum_{i=1}^3 y_i)\,(\sum_{i=1}^3 x_i\,y_i) 
--\frac{101}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i^2) \nonumber \\ &
-+\frac{  9}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i)^2 ~, \label{eq:result}
+    A^2 = & \
+    \frac{2}{4} \, (\sum_{i=1}^{3} x_i ) \, (\sum_{i=1}^{3} y_i ) \, (\sum_{i=1}^{3} x_i \, y_i )
+    \nonumber \\ &
+    + \frac{3}{4} \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i^2 )
+    - \frac{3}{4} \, (\sum_{i=1}^{3} x_i \, y_i )^2.
+    \nonumber \\ &
+    - \, \frac{1}{4} \, (\sum_{i=1}^{3} x_i )^2 \, (\sum_{i=1}^{3} y_i^2 )
+    - \, \frac{1}{4} \, (\sum_{i=1}^{3} y_i )^2 \, (\sum_{i=1}^{3} x_i^2 )
+     ~, \label{eq:result}
 \end{align}
@@
 This answer is in Deep Sets form:
-Every expression in parentheses in \eqref{eq:result} is a pure sum over all three points.
+Every expression in parentheses in \eqref{eq:result} is a pure sum over all three points and can be written in the form of $F_{mn}$ for some choice of $m$ and $n$.
 There are no sums over differences, and there is no ordering of the points.
 Being very explicit, this can be written in the Deep Sets form \eqref{eq:DeepSets} if we define $\rho$ and $\phi$ as follows:
-
-HOGG: Explicit vector $\phi()$ and explicit expression $\rho()$.
-
-HAINJE: Here it is:
 \begin{gather}
-    \phi(r_i) = (x_i, y_i, x_i\,y_i, x_i^2, y_i^2, x_i\,y_i^2, x_i^2\,y_i, x_i^2\,y_i^2) \\
-    \begin{aligned}
-    \rho(\textstyle\sum_{i=1}^{3}\phi(r_i)\equiv \vec{u}) = &
-    - \tfrac{54}{368} u_8
-    - \tfrac{258}{368} u_3^2
-    + \tfrac{285}{368} u_4\,u_5
-    + \tfrac{36}{368} u_2\,u_7
-    + \tfrac{36}{368} u_1\,u_6 \\ &
-    - \tfrac{101}{368} u_2^2\,u_4
-    + \tfrac{148}{368} u_1\,u_2\,u_3
-    - \tfrac{101}{368} u_1^2\,u_5
-    + \tfrac{9}{368} u_1^2\,u_2^2.
-    \end{aligned}
+    \phi(\vec{r}_i) = (
+        x_i, \,
+        y_i, \,
+        x_i^2, \,
+        x_i \, y_i, \,
+        y_i^2
+    ), \\
+    {\textstyle \sum_i} \phi(\vec{r}_i) = (
+        F_{10}, \,
+        F_{01}, \,
+        F_{20}, \,
+        F_{11}, \,
+        F_{02}
+    ), \\
+    \rho \left( {\textstyle \sum_i} \phi(\vec{r}_i) \right)
+    = \tfrac{2}{4} \, F_{10} \, F_{01} \, F_{11}
+    + \tfrac{3}{4} \, F_{20} \, F_{02}
+    - \tfrac{3}{4} \, F_{11}^2
+    - \tfrac{1}{4} \, F_{10}^2 \, F_{02}
+    - \tfrac{1}{4} \, F_{01}^2 \, F_{20}.
 \end{gather}
 
-Note the symmetries....! See the $36/368$ terms, or the $-101/368$ terms
+% HOGG: Explicit vector $\phi()$ and explicit expression $\rho()$.
+
+% HAINJE: Here it is:
+% \begin{gather}
+%     \phi(r_i) = (x_i, y_i, x_i\,y_i, x_i^2, y_i^2, x_i\,y_i^2, x_i^2\,y_i, x_i^2\,y_i^2) \\
+%     \begin{aligned}
+%     \rho(\textstyle\sum_{i=1}^{3}\phi(r_i)\equiv \vec{u}) = &
+%     - \tfrac{54}{368} u_8
+%     - \tfrac{258}{368} u_3^2
+%     + \tfrac{285}{368} u_4\,u_5
+%     + \tfrac{36}{368} u_2\,u_7
+%     + \tfrac{36}{368} u_1\,u_6 \\ &
+%     - \tfrac{101}{368} u_2^2\,u_4
+%     + \tfrac{148}{368} u_1\,u_2\,u_3
+%     - \tfrac{101}{368} u_1^2\,u_5
+%     + \tfrac{9}{368} u_1^2\,u_2^2.
+%     \end{aligned}
+% \end{gather}
+
+% Note the symmetries....! See the $36/368$ terms, or the $-101/368$ terms
 
@@
 Formulae \eqref{eq:cross}, \eqref{eq:sine}, and \eqref{eq:Heron} are obviously rotation-invariant and translation-invariant.
+
+If the triangle is embedded in a higher-dimensional space with the positions of the three vertices given by $\vec{r}_i \in \mathbb{R}^{d}$ for $d \geq 2$, we note that the area of the triangle can be determined by summing in quadrature the areas of the triangles made by all projections to two axes.
+For example, in three dimensions, the area of the triangle is given by
+\begin{equation}
+    A_{\rm 3D}^2 =
+    A_{12}^2 + A_{13}^2 + A_{23}^2,
+\end{equation}
+and, in four dimensions, it is given by
+\begin{equation}
+    A_{\rm 4D}^2 =
+    A_{12}^2 + A_{13}^2 + A_{14}^2
+    + A_{23}^2 + A_{24}^2
+    + A_{34}^2,
+\end{equation}
+where we have introduced $A_{jk}$ the area of the triangle formed by just the $j^{\rm th}$ and $k^{\rm th}$ components of the position vectors.
+
+Our method also extends to the hypervolume of a $d$-simplex.
+Our findings are simplified by the realization that our result is a polynomial with \emph{integer} coefficients for the squared area of a parallelogram.
+Expecting this to generalize, we note that the hypervolume of the $d$-simplex ($S$) formed by vectors $\{ \vec{v}_{0}, \vec{v}_{1}, \cdots, \vec{v}_{d} \}$ is related to the hypervolume of the $d$-parallelotope ($P$) spanned by the vectors $\{ \vec{v}_{1} - \vec{v}_0, \cdots, \vec{v}_d - \vec{v}_0 \}$ by
+\begin{equation}
+    {\rm volume}(P)
+    = (d!) \, {\rm volume}(S)
+\end{equation}
+The squared hypervolume of a $d$-parallelotope is given by the determinant of the Gram matrix of its constituent vectors, e.g.,
+\begin{equation}
+    {\rm volume}(P)^2
+    = \det G, \quad
+    G_{ij} \equiv (\vec{v}_i - \vec{v}_0) \cdot (\vec{v}_j - \vec{v}_0).
+\end{equation}
+In the two-dimensional case, our formula \eqref{eq:result} gives an integer polynomial for the area of the parallelogram, equal to $4 A^2$ for $A$ the area of the triangle.
+To find the hypervolumes of higher-dimensional simplexes, we modify our symbolic regression to seek integer polynomials for the squared volume of the related parallelotope.
+
+In three dimensions, four vectors form the vertices of a simplex. We find an analogous Deep Sets-form formula for the volume of the simplex to be
+\begin{align}
+(6 V)^2 = &
++8 \, F_{1100} \, F_{1010} \, F_{0110}
++4 \, F_{200} \, F_{020} \, F_{0020}
+\nonumber \\ &
+-4 \, F_{200} \, F_{011}^2
+-4 \, F_{020} \, F_{101}^2
+-4 \, F_{002} \, F_{110}^2
+\nonumber \\ &
++2 \, F_{200} \, F_{011} \, F_{010} \, F_{001}
++2 \, F_{020} \, F_{101} \, F_{100} \, F_{001}
++2 \, F_{002} \, F_{110} \, F_{100} \, F_{010}
+\nonumber \\ &
+-2 \, F_{110} \, F_{101} \, F_{010} \, F_{001}
+-2 \, F_{110} \, F_{011} \, F_{100} \, F_{001}
+-2 \, F_{101} \, F_{011} \, F_{100} \, F_{010}
+\nonumber \\ &
++1 \, F_{110}^2 \, F_{001}^2
++1 \, F_{101}^2 \, F_{010}^2
++1 \, F_{011}^2 \, F_{100}^2
+\nonumber \\ &
+-1 \, F_{200} \, F_{020} \, F_{001}^2
+-1 \, F_{200} \, F_{002} \, F_{010}^2
+-1 \, F_{020} \, F_{002} \, F_{100}^2
+\label{eq:simplex}
+\end{align}
 
@@
 \newpage
-By inspection, we have also found a simpler form of the same formula, which is given by
-\begin{align}
-    4 \, A^2 = & \
-        2 \, (\sum_{i=1}^{3} x_i ) \, (\sum_{i=1}^{3} y_i ) \, (\sum_{i=1}^{3} x_i \, y_i )
-        \nonumber \\ &
-        + 3 \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i^2 )
-        - 3 \, (\sum_{i=1}^{3} x_i \, y_i )^2.
-        \nonumber \\ &
-        - \, (\sum_{i=1}^{3} x_i )^2 \, (\sum_{i=1}^{3} y_i^2 )
-        - \, (\sum_{i=1}^{3} y_i )^2 \, (\sum_{i=1}^{3} x_i^2 )
-\end{align}
-In explicit Deep Sets notation, this can be written
-\begin{align}
-    \phi(r_i) &= \begin{bmatrix}
-        x_i & y_i & x_i\,y_i & x_i^2 & y_i^2 \end{bmatrix}^\top
-        \equiv \vec{u}^\top, \\
-    \rho(\sum_{i=1}^{3} \phi(r_i))
-    &= \tfrac{2}{4} u_1\,u_2\,u_3
-    + \tfrac{3}{4} u_4\,u_5
-    - \tfrac{1}{4} u_1^2\,u_5
-    - \tfrac{1}{4} u_2^2\,u_4
-    - \tfrac{3}{4} u_3^2.
-\end{align}
-
-The two are related by
-\begin{equation}
-    4\,u_1\,u_6 + 4\,u_2\,u_7 - 6\,u_8 = 
-    -u_1^2 u_2^2 + 4 u_1 u_2 u_3 - 2 u_3^2 + u_2^2 u_4 + u_1^2 u_5 - u_4 u_5 ;
-\end{equation}
-namely, plugging this into \eqref{eq:result} causes it to reduce immediately into this simpler form.
-
-This form was not found by the symbolic regression of \secref{sec:method} because the regression makes use of \verb+lstsq+ with more parameters than data points.
-In this regime, the solution that minimizes the $\mathcal{L}_2$ norm of the weights (the coefficients of the invariant monomials) is returned.
-The $\mathcal{L}_2$ norm of the weights of our initial formula is $2127 / 1472 \approx 1.445$, whereas the $\mathcal{L}_2$ norm of the weights of this simpler formula is $3/2 = 1.5$.
-
-HAINJE: Would lasso regression return my formula? Is there a yet simpler formula?
 
@@
 
-For reference, the 2D formula was
-\begin{equation}
-4 A_{\rm 2D}^2 = +3 F_{02} F_{20} -3 F_{11}^2 - F_{20} F_{01}^2 +2 F_{11} F_{10} F_{01} - F_{02} F_{10}^2.
-\end{equation}
-In 3D, we have
 \begin{equation}
 \begin{aligned}
-4 A_{\rm 3D}^2 = &
-+3 F_{020} F_{002} 
-+3 F_{200} F_{002} 
-+3 F_{020} F_{200} \\&
--3 F_{011}^2 
--3 F_{101}^2 
--3 F_{110}^2 \\&
-+2 F_{011} F_{001} F_{010} 
-+2 F_{100} F_{001} F_{101}
-+2 F_{100} F_{010} F_{110} \\&
--F_{002} F_{010}^2 
--F_{002} F_{100}^2 
--F_{020} F_{001}^2 
--F_{020} F_{100}^2
--F_{200} F_{010}^2
--F_{200} F_{001}^2. 
+V^2 = &
++8 F_{110} F_{101} F_{011}
++4 F_{200} F_{020} F_{002}
+\\ &
+-4 F_{200} F_{011}^2
+-4 F_{020} F_{101}^2
+-4 F_{002} F_{110}^2
+\\ &
++2 F_{200} F_{011} F_{010} F_{001}
++2 F_{020} F_{101} F_{100} F_{001}
++2 F_{002} F_{110} F_{100} F_{010}
+\\ &
+-2 F_{110} F_{101} F_{010} F_{001}
+-2 F_{110} F_{011} F_{100} F_{001}
+-2 F_{101} F_{011} F_{100} F_{010}
+\\ &
++1 F_{110}^2 F_{001}^2
++1 F_{101}^2 F_{010}^2
++1 F_{011}^2 F_{100}^2
+\\ &
+-1 F_{200} F_{020} F_{001}^2
+-1 F_{200} F_{002} F_{010}^2
+-1 F_{020} F_{002} F_{100}^2
 \end{aligned}
 \end{equation}
-And in 4D, we have
+
 \begin{equation}
 \begin{aligned}
-4 A_{\rm 4D}^2 = &
-+3 F_{0002} F_{0020} 
-+3 F_{0002} F_{0200} 
-+3 F_{0002} F_{2000} 
-\\&
-+3 F_{0200} F_{2000} 
-+3 F_{0200} F_{0020} 
-+3 F_{0020} F_{2000} 
-\\&
--3 F_{0011}^2
--3 F_{0101}^2
--3 F_{0110}^2
--3 F_{1001}^2 
--3 F_{1010}^2 
--3 F_{1100}^2
-\\&
-+2 F_{0010} F_{0011} F_{0001} 
-+2 F_{0101} F_{0100} F_{0001} 
-+2 F_{1001} F_{1000} F_{0001}
-\\&
-+2 F_{0010} F_{0100} F_{0110} 
-+2 F_{0010} F_{1010} F_{1000}
-+2 F_{1100} F_{0100} F_{1000} 
-\\&
-- F_{0020} F_{0001}^2
-- F_{0200} F_{0001}^2
-- F_{2000} F_{0001}^2
-- F_{0002} F_{0010}^2
-- F_{0200} F_{0010}^2
-- F_{2000} F_{0010}^2
-\\&
-- F_{0002} F_{0100}^2 
-- F_{0020} F_{0100}^2 
-- F_{2000} F_{0100}^2
-- F_{0002} F_{1000}^2 
-- F_{0020} F_{1000}^2 
-- F_{0200} F_{1000}^2
+V^2 = &
++8 F_{1100} F_{1010} F_{0110}
++4 F_{2000} F_{0200} F_{0020}
+\\ &
+-4 F_{2000} F_{0110}^2
+-4 F_{0200} F_{1010}^2
+-4 F_{0020} F_{1100}^2
+\\ &
++2 F_{2000} F_{0110} F_{0100} F_{0010}
++2 F_{0200} F_{1010} F_{1000} F_{0010}
++2 F_{0020} F_{1100} F_{1000} F_{0100}
+\\ &
+-2 F_{1100} F_{1010} F_{0100} F_{0010}
+-2 F_{1100} F_{0110} F_{1000} F_{0010}
+-2 F_{1010} F_{0110} F_{1000} F_{0100}
+\\ &
++1 F_{1100}^2 F_{0010}^2
++1 F_{1010}^2 F_{0100}^2
++1 F_{0110}^2 F_{1000}^2
+\\ &
+-1 F_{2000} F_{0200} F_{0010}^2
+-1 F_{2000} F_{0020} F_{0100}^2
+-1 F_{0200} F_{0020} F_{1000}^2
+\\ &
+\\ &
++8 F_{1100} F_{1001} F_{0101}
+\\ &
++4 F_{2000} F_{0200} F_{0002}
+-4 F_{2000} F_{0101}^2
+\\ &
+-4 F_{1100}^2 F_{0002}
+-4 F_{0200} F_{1001}^2
+\\ &
++2 F_{2000} F_{0001} F_{0100} F_{0101}
+-2 F_{1100} F_{1001} F_{0001} F_{0100}
+\\ &
++2 F_{1100} F_{1000} F_{0100} F_{0002}
+-2 F_{1001} F_{1000} F_{0100} F_{0101}
+\\ &
+-2 F_{1000} F_{1100} F_{0001} F_{0101}
++2 F_{1000} F_{0200} F_{1001} F_{0001}
+\\ &
+-1 F_{2000} F_{0200} F_{0001}^2
+-1 F_{2000} F_{0100}^2 F_{0002}
+\\ &
++1 F_{1100}^2 F_{0001}^2
++1 F_{1001}^2 F_{0100}^2
+\\ &
++1 F_{1000}^2 F_{0101}^2
+-1 F_{0200} F_{1000}^2 F_{0002}
+\\ &
+\\ &
++8 F_{1010} F_{1001} F_{0011}
+\\ &
++4 F_{2000} F_{0020} F_{0002}
+-4 F_{2000} F_{0011}^2
+\\ &
+-4 F_{1010}^2 F_{0002}
+-4 F_{1001}^2 F_{0020}
+\\ &
++2 F_{2000} F_{0010} F_{0001} F_{0011}
++2 F_{1010} F_{0010} F_{1000} F_{0002}
+\\ &
+-2 F_{1000} F_{1010} F_{0001} F_{0011}
++2 F_{1000} F_{1001} F_{0001} F_{0020}
+\\ &
+-2 F_{0010} F_{1000} F_{1001} F_{0011}
+-2 F_{0010} F_{0001} F_{1001} F_{1010}
+\\ &
+-1 F_{2000} F_{0010}^2 F_{0002}
+-1 F_{2000} F_{0001}^2 F_{0020}
+\\ &
++1 F_{1000}^2 F_{0011}^2
+-1 F_{0020} F_{1000}^2 F_{0002}
+\\ &
++1 F_{0010}^2 F_{1001}^2
++1 F_{0001}^2 F_{1010}^2
+\\ &
+\\ &
++8 F_{0101} F_{0110} F_{0011}
+\\ &
++4 F_{0200} F_{0020} F_{0002}
+-4 F_{0200} F_{0011}^2
+\\ &
+-4 F_{0110}^2 F_{0002}
+-4 F_{0020} F_{0101}^2
+\\ &
++2 F_{0200} F_{0010} F_{0001} F_{0011}
+-2 F_{0110} F_{0010} F_{0001} F_{0101}
+\\ &
+-2 F_{0110} F_{0001} F_{0100} F_{0011}
++2 F_{0101} F_{0001} F_{0100} F_{0020}
+\\ &
++2 F_{0010} F_{0110} F_{0100} F_{0002}
+-2 F_{0010} F_{0101} F_{0100} F_{0011}
+\\ &
+-1 F_{0200} F_{0010}^2 F_{0002}
+-1 F_{0200} F_{0001}^2 F_{0020}
+\\ &
++1 F_{0100}^2 F_{0011}^2
+-1 F_{0020} F_{0100}^2 F_{0002}
+\\ &
++1 F_{0010}^2 F_{0101}^2
++1 F_{0001}^2 F_{0110}^2
 \end{aligned}

### 2025-03-06 paper/main.tex
content, abstract

@@
 \usepackage[hidelinks]{hyperref}
-\usepackage{amsmath,amsfonts}
+\usepackage{amsmath,amsfonts,tikz}
+\usetikzlibrary{angles,calc,quotes}
 
@@
 
+\usepackage{xcolor}
+\newcommand{\CH}[1]{{\color{blue} (CH: #1)}}
+
 \begin{document}\thispagestyle{empty}
@@
 A new formula for the area of a triangle:
-Useless, but explicitly in Deep Sets and map--reduce form\footnote{%
+Useless, but explicitly in Deep Sets form\footnote{%
 The authors thank Alexander Novara (NYU), Ben Wandelt (Sorbonne), and Kaze Wong (JHU) for valuable discussions.}}
@@
 where $\rho$ and $\phi$ are (probably nonlinear) functions, and the function $\phi$ might output a high-dimensional latent encoding of each point $r_i$.
-This form---known in the machine-learning literature as Deep Sets---is a map--reduce form, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically or in any order.
+This form---known in the machine-learning literature as Deep Sets---is also a map--reduce, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically or in any order.
 The area of a triangle is a permutation-invariant function of the locations $r_i$ of the three corners $1\leq i\leq 3$.
 We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in Deep Sets form.
-% It is the square-root of a permutation-invariant quartic polynomial involving the integer 368.
-No insights of any kind were gained from this work.
+This project was motivated by questions about the fundamental computational complexity of n-point statistics in cosmology; that said, no insights of any kind were gained from these results.
 
 \section{Introduction}
-The area $A$ of a triangle can be written in many ways.
+
+\begin{figure}
+    \centering
+    \begin{tikzpicture}
+    \coordinate (O) at (0,0);
+    \coordinate (A) at (3,4);
+    \coordinate (B) at (5,1);
+
+    \pgfmathsetmacro{\a}{sqrt(3*3 + 4*4)}
+    \pgfmathsetmacro{\b}{sqrt(5*5 + 1*1)}
+    \pgfmathsetmacro{\c}{sqrt(2*2 + 3*3)}
+    \pgfmathsetmacro{\s}{(\a+\b+\c)/2}
+    \pgfmathsetmacro{\h}{2*sqrt(\s*(\s-\a)*(\s-\b)*(\s-\c))/\b}
+
+    \pgfmathsetmacro{\DeltaX}{5 - 0}  % Bx - Ox
+    \pgfmathsetmacro{\DeltaY}{1 - 0}  % By - Oy
+    \pgfmathsetmacro{\len}{sqrt(\DeltaX*\DeltaX + \DeltaY*\DeltaY)}
+
+    \def\bLen{0.5mm}
+    \def\hLen{0.3mm}
+    \pgfmathsetmacro{\bdx}{-\DeltaY/\len * \bLen}
+    \pgfmathsetmacro{\bdy}{\DeltaX/\len * \bLen}
+    \pgfmathsetmacro{\hdx}{\DeltaX/\len * \hLen}
+    \pgfmathsetmacro{\hdy}{\DeltaY/\len * \hLen}
+    \pgfmathsetmacro{\hx}{-\DeltaY/\len * \h}
+    \pgfmathsetmacro{\hy}{\DeltaX/\len * \h}
+
+    \coordinate (BaseO) at ($(O) - (\bdx,\bdy)$);
+    \coordinate (BaseB) at ($(B) - (\bdx,\bdy)$);
+    \coordinate (HeightO) at ($(B) + (\hdx,\hdy)$);
+    \coordinate (HeightA) at ($(B) + (\hdx,\hdy) + (\hx,\hy)$);
+
+    % mark points
+    \fill (O) circle (1.2pt) node[below right]{$(x_1, y_1)$};
+    \fill (A) circle (1.2pt) node[above, yshift=+1mm]{$(x_2, y_2)$};
+    \fill (B) circle (1.2pt) node[below right]{$(x_3, y_3)$};
+
+    % draw vectors
+    \def\pad{1.5mm}
+    \draw[->, line width=0.4mm]
+        ($(O)!\pad!(A)$) -- ($(A)!\pad!(O)$)
+        node[midway, left, xshift=-1mm, yshift=+1mm] {$\vec{a}$};
+    \draw[->, line width=0.4mm]
+        ($(O)!\pad!(B)$) -- ($(B)!\pad!(O)$)
+        node[midway, below, xshift=+1mm, yshift=-0.5mm] {$\vec{b}$};
+    \draw[->, line width=0.4mm]
+        ($(B)!\pad!(A)$) -- ($(A)!\pad!(B)$)
+        node[midway, right, xshift=+1mm, yshift=+1mm] {$\vec{c}$};
+
+    % draw arc for angle
+    \draw pic["$C$", draw=black, angle radius=12mm, angle eccentricity=1.3] {angle=B--O--A};
+
+    % draw base, height guide lines
+    \draw ($(O)!\pad!(BaseO)$) -- ($(BaseO)$);
+    \draw ($(B)!\pad!(BaseB)$) -- ($(BaseB)$);
+    \draw ($(B)!\pad!(HeightO)$) -- ($(HeightO)$);
+    \draw ($(A)!\pad!(HeightA)$) -- ($(HeightA)$);
+
+    % draw base, height size lines
+    \draw[<->]
+        ($(BaseO)!\pad!(O)!\pad!(BaseB)$) -- ($(BaseB)!\pad!(B)!\pad!(BaseO)$)
+        node[midway, below, xshift=+2mm, yshift=-1mm] {base};
+    \draw[<->]
+        ($(HeightO)!\pad!(O)!\pad!(HeightA)$) -- ($(HeightA)!\pad!(A)!\pad!(HeightO)$)
+        node[midway, right, xshift=+1mm, yshift=+1mm] {height};
+
+    \end{tikzpicture}
+    \caption{A triangle.}
+    \label{fig:triangle}
+\end{figure}
+
+In Figure~\ref{fig:triangle}, we present a triangle.
+The area $\Delta$ of this triangle can be written in many ways.
 Here are a few:
 \begin{align}
-A &= \frac{1}{2}\,\text{(base)}\,\text{(height)} \label{eq:school} \\
-  &= \frac{1}{2}\,|(\vec{r}_3 - \vec{r}_1) \times (\vec{r}_2 - \vec{r}_1)| \label{eq:cross} \\
+\Delta 
+&= \frac{1}{2} \, \text{(base)} \, \text{(height)} \label{eq:school} \\
+&= \frac{1}{2}\, \| \vec{a} \times \vec{b} \| \label{eq:cross} \\
 &= \frac{1}{2}\,a\,b\,\sin(C) \label{eq:sine} \\
@@
 where the first formula \eqref{eq:school} is the usual introduction to the subject,
-$\vec{r}_1, \vec{r}_2, \vec{r}_3$ are three vectors pointing from an origin to the corners of the triangle,
-$\times$ represents the cross-product operator (or its generalization to higher dimensions),
-$a, b, c$ are side lengths of the triangle,
-$C$ is the angle between the $a$ and $b$ sides,
+$\times$ represents the cross-product operator,
+$a = \| \vec{a} \|$, $b = \| \vec{b} \|$, $c = \| \vec{c} \|$ are side lengths of the triangle,
+% $C$ is the angle between $\vec{a}$ and $\vec{b}$,
 and $s$ is the semi-perimeter $(a + b + c)/2$.
-Below, we present a new formula for the area of a triangle, which we believe may be novel.
+For a rather exhaustive list of 110 unique formulae, see \citet{baker1885a,baker1885b}.
 
-While the area of a triangle does not depend on how its sides or corners are labeled,
+While the area of the triangle does not depend on how its sides or corners are labeled,
 only the last of these formulae---Heron's formula \eqref{eq:Heron}---is explicitly permutation-invariant.
 
-Theorem 2 of \citet{zaheer+17-Deep-Sets} states that a function $f(X)$ operating on a set $X$ is invariant to the permutation of instances in $X$ if and only if it can be decomposed in the form
+Permutation invariance is an important symmetry in a number of fields for problems defined on \textit{sets}.
+In machine learning, one area of relevance is deep learning on point clouds (e.g., sets of three-dimensional position vectors) \citep{guo2020pointclouds}, which is particularly useful for problems such as 3D shape classification and object detection.
+In physics, many fundamental problems are invariant to permutation.
+For example, the gravitational/electric potential is invariant to the labeling of point masses/charges \CH{citation needed?}.
+Further, permutation invariance lies at the heart of all modern cosmological statistics, such as the power spectrum and the bispectrum \CH{cite}.
+
+There is a simple and powerful result which is useful for constructing and characterizing permutation-invariant functions.
+Given a function $f(r_1, \cdots, r_N)$, Theorem 2 of \citet{zaheer+17deepsets} states that $f$ is invariant to permutation of its arguments if and only if it can be decomposed in the form
 \begin{equation}
-    \textstyle f(X) = \rho \left( \sum_{x \in X} \phi(x) \right), \label{eq:DeepSets}
+    f(r_1, \cdots, r_N) = \rho \left( {\textstyle \sum_{i=1}^{N}} \, \phi(r_i) \right), \label{eq:DeepSets}
 \end{equation}
 for some functions $\rho$ and $\phi$.
-This is proven for the cases where the elements of $X$ are from a countable universe and the size of $X$ is unlimited, and for the case where the elements of $X$ are from an uncountable universe but the size of $X$ is finite.
-This result---often known by the name of the machine learning architecture derived from it, \emph{Deep Sets} \citep{zaheer+17-Deep-Sets}---is not at all obvious:
-For example, even though Heron's formula \eqref{eq:Heron} is permutation-invariant, it is not actually in the form \eqref{eq:DeepSets}, if the triangle is seen as being defined by its corners (or even its sides).
+
+This result---often known by the name of the deep learning architecture derived from it, \emph{Deep Sets} \citep{zaheer+17deepsets}---is not at all obvious:
+For example, even though Heron's formula \eqref{eq:Heron} is permutation-invariant, it is not actually in the form of \eqref{eq:DeepSets}. Consider instead the formula
+\begin{equation}
+    \Delta = \frac{1}{4} \sqrt{
+        \left( a^2 + b^2 + c^2 \right)^2
+        - 2 \left( a^4 + b^4 + c^4 \right)
+    }.
+\end{equation}
+It is simple to show that this formula is exactly equivalent to Heron's formula when expanded, but it is now precisely in Deep Sets form. Explicitly,
+\begin{equation}
+    \phi(x) = (x^2, x^4), \quad
+    \vec{u} = {\textstyle \sum_{x \in \{ a, b, c \}}} \, \phi(x), \quad
+    \Delta = \rho(\vec{u}) = \frac{1}{4} \sqrt{ u_1^2 - 2 \, u_2 }.
+\end{equation}
 
@@
 \begin{itemize}
-    \item The function $\phi$ is universal to all the elements $x$; its form won't depend on label $i$ or on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
+    \item The function $\phi$ is universal to all the elements $r_i$; its form won't depend on label $i$ or on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
     \item The point properties $r_i$ and the output of the function $\phi$ might both be very large.
@@
     The output of the function $\phi$ is some kind of informative, latent encoding of the point $r_i$.
-    \item If a function can be written in Deep-Sets form \eqref{eq:DeepSets}, then that function can be executed in a map--reduce framework, in which the function $\phi$ is executed independently and asynchronously on each data point (maybe even in the data center) and the sums are performed hierarchically or in any order.
+    \item If a function can be written in Deep Sets form \eqref{eq:DeepSets}, then that function can be executed in a map--reduce framework, in which the function $\phi$ is executed independently and asynchronously on each data point (maybe even in the data center) and the sums are performed hierarchically or in any order.
     That means that for large problems executed on large data sets in large data centers, problems in this form \eqref{eq:DeepSets} can often be performed in wall-clock time that scales as $\log N$.
@@
 These monomials are both correct from a units perspective, but the second is not in Deep Sets form, because it contains products of position coordinates from different points. 
-In order to generate all the possible invariant monomials, we first note that they are made of products of factors of the form $F_{m,n} = \sum_i x_i^m y_i^n$.
-The invariant monomials of degree $d$ are then all products $\prod_i F_{m_i, n_i}$ such that $\sum_i m_i + n_i = d$.
-
-% HOGG: How to construct all the invariant polynomials.
-
-% HAINJE: Given data of form $(x, y)$, invariant factors are $F_{m,n} = \sum_i x_i^m y_i^n$, and the invariant monomials are any products of these factors $\prod_{i} F_{m_i,n_i}$.
-Do we want/need to be more specific than that?
+All the possible monomials are made of permutation-invariant factors of the form 
+\begin{equation}
+F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n.
+\end{equation}
+The invariant monomials of degree $d$ are then all products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
 
@@
 
-Our result---and only contribution---is that the area $A$ of a triangle is given by
+Our result is that the area $A$ of a triangle is given by
 % \begin{align}
@@
     A^2 = & \
-    \frac{2}{4} \, (\sum_{i=1}^{3} x_i ) \, (\sum_{i=1}^{3} y_i ) \, (\sum_{i=1}^{3} x_i \, y_i )
+    \frac{2}{4} \, (\sum_{i=1}^{3} x_i \, y_i ) \, (\sum_{i=1}^{3} x_i ) \, (\sum_{i=1}^{3} y_i )
     \nonumber \\ &
@@
     \nonumber \\ &
-    - \, \frac{1}{4} \, (\sum_{i=1}^{3} x_i )^2 \, (\sum_{i=1}^{3} y_i^2 )
-    - \, \frac{1}{4} \, (\sum_{i=1}^{3} y_i )^2 \, (\sum_{i=1}^{3} x_i^2 )
+    - \, \frac{1}{4} \, (\sum_{i=1}^{3} y_i^2 ) \, (\sum_{i=1}^{3} x_i )^2
+    - \, \frac{1}{4} \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i )^2
      ~, \label{eq:result}
@@
 There are no sums over differences, and there is no ordering of the points.
-Being very explicit, this can be written in the Deep Sets form \eqref{eq:DeepSets} if we define $\rho$ and $\phi$ as follows:
+Being very explicit, this is written in the Deep Sets form \eqref{eq:DeepSets} for $\rho$ and $\phi$ defined as follows:
 \begin{gather}
@@
     \rho \left( {\textstyle \sum_i} \phi(\vec{r}_i) \right)
-    = \tfrac{2}{4} \, F_{10} \, F_{01} \, F_{11}
+    = \tfrac{2}{4} \, F_{11} \, F_{10} \, F_{01}
     + \tfrac{3}{4} \, F_{20} \, F_{02}
     - \tfrac{3}{4} \, F_{11}^2
-    - \tfrac{1}{4} \, F_{10}^2 \, F_{02}
-    - \tfrac{1}{4} \, F_{01}^2 \, F_{20}.
+    - \tfrac{1}{4} \, F_{02} \, F_{10}^2
+    - \tfrac{1}{4} \, F_{20} \, F_{01}^2.
 \end{gather}
@@
 
-But note that it isn't obviously invariant to arbitrary rotations.
-And it isn't obviously invariant to translations.
-Formulae \eqref{eq:cross}, \eqref{eq:sine}, and \eqref{eq:Heron} are obviously rotation-invariant and translation-invariant.
+% But note that it isn't obviously invariant to arbitrary rotations.
+% And it isn't obviously invariant to translations.
+% Formulae \eqref{eq:cross}, \eqref{eq:sine}, and \eqref{eq:Heron} are obviously rotation-invariant and translation-invariant.
 
-If the triangle is embedded in a higher-dimensional space with the positions of the three vertices given by $\vec{r}_i \in \mathbb{R}^{d}$ for $d \geq 2$, we note that the area of the triangle can be determined by summing in quadrature the areas of the triangles made by all projections to two axes.
+Note that a triangle defined by three points has the one-half the area of the parallelogram spanned by the vectors $\vec{r}_2 - \vec{r}_1$ and $\vec{r}_3 - \vec{r}_1$.
+Multiplying both sides by 4 then reveals that \eqref{eq:result} is an integer polynomial of invariant factors for the area of the parallelogram.
+
+If the triangle is embedded in a higher-dimensional space with the positions of the three vertices given by $\vec{r}_i \in \mathbb{R}^{d}$ for $d \geq 2$, the area of the triangle can be determined by summing in quadrature the areas of the triangles made by all projections to two axes.
 For example, in three dimensions, the area of the triangle is given by
@@
 \end{equation}
-where we have introduced $A_{jk}$ the area of the triangle formed by just the $j^{\rm th}$ and $k^{\rm th}$ components of the position vectors.
+where we have introduced $A_{jk}$, the area of the triangle formed by just the $j^{\rm th}$ and $k^{\rm th}$ components of the position vectors.
+By using \eqref{eq:result} for each of the two-dimensional sub-areas, each can be computed in a permutation-invariant way and the entire result is thus permutation invariant (and in Deep Sets form).
 
-Our method also extends to the hypervolume of a $d$-simplex.
-Our findings are simplified by the realization that our result is a polynomial with \emph{integer} coefficients for the squared area of a parallelogram.
-Expecting this to generalize, we note that the hypervolume of the $d$-simplex ($S$) formed by vectors $\{ \vec{v}_{0}, \vec{v}_{1}, \cdots, \vec{v}_{d} \}$ is related to the hypervolume of the $d$-parallelotope ($P$) spanned by the vectors $\{ \vec{v}_{1} - \vec{v}_0, \cdots, \vec{v}_d - \vec{v}_0 \}$ by
-\begin{equation}
-    {\rm volume}(P)
-    = (d!) \, {\rm volume}(S)
-\end{equation}
-The squared hypervolume of a $d$-parallelotope is given by the determinant of the Gram matrix of its constituent vectors, e.g.,
-\begin{equation}
-    {\rm volume}(P)^2
-    = \det G, \quad
-    G_{ij} \equiv (\vec{v}_i - \vec{v}_0) \cdot (\vec{v}_j - \vec{v}_0).
-\end{equation}
-In the two-dimensional case, our formula \eqref{eq:result} gives an integer polynomial for the area of the parallelogram, equal to $4 A^2$ for $A$ the area of the triangle.
-To find the hypervolumes of higher-dimensional simplexes, we modify our symbolic regression to seek integer polynomials for the squared volume of the related parallelotope.
-
-In three dimensions, four vectors form the vertices of a simplex. We find an analogous Deep Sets-form formula for the volume of the simplex to be
+Our method also extends to the volume of a simplex. In three dimensions, four vectors $\{ \vec{r}_1, \vec{r}_2, \vec{r}_3, \vec{r}_4 \}$ determine the vertices of a simplex. 
+Like the area of a triangle is related to the area of a parallelogram, the volume of a simplex is one-sixth the volume of the parallelepiped spanned by the vectors $\{ \vec{r}_2 - \vec{r}_1, \vec{r}_3 - \vec{r}_1, \vec{r}_4 - \vec{r}_1 \}$.
+Hoping for an integer solution once again, we modify our symbolic regression to search for polynomials of the invariant factors which yield the volume of the parallelepiped.
+This gives an analogous Deep Sets formula for the volume of a simplex to be
 \begin{align}
 (6 V)^2 = &
-+8 \, F_{1100} \, F_{1010} \, F_{0110}
-+4 \, F_{200} \, F_{020} \, F_{0020}
++8 \, F_{110} \, F_{101} \, F_{011}
++4 \, F_{200} \, F_{020} \, F_{002}
 \nonumber \\ &
@@
 -1 \, F_{200} \, F_{002} \, F_{010}^2
--1 \, F_{020} \, F_{002} \, F_{100}^2
+-1 \, F_{020} \, F_{002} \, F_{100}^2 \, .
 \label{eq:simplex}
@@
 
-HOGG SAY: CH: EXPAND THIS INTO AN OPEN QUESTIONS LIST:
+% HOGG SAY: CH: EXPAND THIS INTO AN OPEN QUESTIONS LIST:
+% \begin{itemize}
+% \item questionable: What happens if we try to do $A$ as a quadratic instead of $A^2$ as a quartic?
+% \item What happens if you hold the formula fixed, but use more than $N=3$ points?
+% The formula still returns a squared area.
+% That's the area of what?
+% \item What about areas of quadrangles or larger polygons?
+% \item What about volumes and hypervolumes of larger objects?
+% \end{itemize}
+
+We leave the reader with the following open questions:
 \begin{itemize}
-\item CH should answer this question: What's the generalization to $d$ dimensions? That is, what if the three points are each in d dimensions?
-\item questionable: What happens if we try to do $A$ as a quadratic instead of $A^2$ as a quartic?
-\item What happens if you hold the formula fixed, but use more than $N=3$ points?
-The formula still returns a squared area.
-That's the area of what?
-\item What about areas of quadrangles or larger polygons?
-\item What about volumes and hypervolumes of larger objects?
+    \item It is trivial to take the form of \eqref{eq:result} and evaluate the sums for more than three vectors. Does the result have any meaning?
+
+    \item Our formula for the volume of a simplex \eqref{eq:simplex} bears many similarities to that for the area of triangle \eqref{eq:result}.
+We take this to imply that a similar formula should exist for the hypervolume of a $d$-simplex.
+What is this result?
+
+    \item In both the triangle area and simplex volume formulae, the only invariant factors used are of degree $2$, and the factors never appear with exponents larger than $2$.
+Does this generalize?
+Can it be shown that higher order factors aren't needed?
 \end{itemize}
-
-\newpage
-
-\newpage
-
-Define the ``invariant factor'' in $d$ dimensions
-\begin{equation}
-    F_{m_1 \, m_2 \, \cdots} \equiv \sum_{1=1}^{3} \prod_{k=1}^{d} (\vec{r}_{i})_{k}^{m_k}.
-\end{equation}
-This is the sum over all three particles of the product of the $d$ components of each $\vec{r}_i$, each component raised to a power. For example, in 2D, $d=2$ and this expands to
-\begin{equation}
-    F_{\ell m} = \sum_{i=1}^{3} x_i^\ell \, y_i^m.
-\end{equation}
-In 3D, $d=3$ and this gives
-\begin{equation}
-    F_{\ell m n} = \sum_{i=1}^{3} x_i^\ell \, y_i^m \, z_i^n.
-\end{equation}
-Then, the area formula found above can be extended to more dimensions with the following formulae.
-
-\begin{equation}
-\begin{aligned}
-V^2 = &
-+8 F_{110} F_{101} F_{011}
-+4 F_{200} F_{020} F_{002}
-\\ &
--4 F_{200} F_{011}^2
--4 F_{020} F_{101}^2
--4 F_{002} F_{110}^2
-\\ &
-+2 F_{200} F_{011} F_{010} F_{001}
-+2 F_{020} F_{101} F_{100} F_{001}
-+2 F_{002} F_{110} F_{100} F_{010}
-\\ &
--2 F_{110} F_{101} F_{010} F_{001}
--2 F_{110} F_{011} F_{100} F_{001}
--2 F_{101} F_{011} F_{100} F_{010}
-\\ &
-+1 F_{110}^2 F_{001}^2
-+1 F_{101}^2 F_{010}^2
-+1 F_{011}^2 F_{100}^2
-\\ &
--1 F_{200} F_{020} F_{001}^2
--1 F_{200} F_{002} F_{010}^2
--1 F_{020} F_{002} F_{100}^2
-\end{aligned}
-\end{equation}
-
-\begin{equation}
-\begin{aligned}
-V^2 = &
-+8 F_{1100} F_{1010} F_{0110}
-+4 F_{2000} F_{0200} F_{0020}
-\\ &
--4 F_{2000} F_{0110}^2
--4 F_{0200} F_{1010}^2
--4 F_{0020} F_{1100}^2
-\\ &
-+2 F_{2000} F_{0110} F_{0100} F_{0010}
-+2 F_{0200} F_{1010} F_{1000} F_{0010}
-+2 F_{0020} F_{1100} F_{1000} F_{0100}
-\\ &
--2 F_{1100} F_{1010} F_{0100} F_{0010}
--2 F_{1100} F_{0110} F_{1000} F_{0010}
--2 F_{1010} F_{0110} F_{1000} F_{0100}
-\\ &
-+1 F_{1100}^2 F_{0010}^2
-+1 F_{1010}^2 F_{0100}^2
-+1 F_{0110}^2 F_{1000}^2
-\\ &
--1 F_{2000} F_{0200} F_{0010}^2
--1 F_{2000} F_{0020} F_{0100}^2
--1 F_{0200} F_{0020} F_{1000}^2
-\\ &
-\\ &
-+8 F_{1100} F_{1001} F_{0101}
-\\ &
-+4 F_{2000} F_{0200} F_{0002}
--4 F_{2000} F_{0101}^2
-\\ &
--4 F_{1100}^2 F_{0002}
--4 F_{0200} F_{1001}^2
-\\ &
-+2 F_{2000} F_{0001} F_{0100} F_{0101}
--2 F_{1100} F_{1001} F_{0001} F_{0100}
-\\ &
-+2 F_{1100} F_{1000} F_{0100} F_{0002}
--2 F_{1001} F_{1000} F_{0100} F_{0101}
-\\ &
--2 F_{1000} F_{1100} F_{0001} F_{0101}
-+2 F_{1000} F_{0200} F_{1001} F_{0001}
-\\ &
--1 F_{2000} F_{0200} F_{0001}^2
--1 F_{2000} F_{0100}^2 F_{0002}
-\\ &
-+1 F_{1100}^2 F_{0001}^2
-+1 F_{1001}^2 F_{0100}^2
-\\ &
-+1 F_{1000}^2 F_{0101}^2
--1 F_{0200} F_{1000}^2 F_{0002}
-\\ &
-\\ &
-+8 F_{1010} F_{1001} F_{0011}
-\\ &
-+4 F_{2000} F_{0020} F_{0002}
--4 F_{2000} F_{0011}^2
-\\ &
--4 F_{1010}^2 F_{0002}
--4 F_{1001}^2 F_{0020}
-\\ &
-+2 F_{2000} F_{0010} F_{0001} F_{0011}
-+2 F_{1010} F_{0010} F_{1000} F_{0002}
-\\ &
--2 F_{1000} F_{1010} F_{0001} F_{0011}
-+2 F_{1000} F_{1001} F_{0001} F_{0020}
-\\ &
--2 F_{0010} F_{1000} F_{1001} F_{0011}
--2 F_{0010} F_{0001} F_{1001} F_{1010}
-\\ &
--1 F_{2000} F_{0010}^2 F_{0002}
--1 F_{2000} F_{0001}^2 F_{0020}
-\\ &
-+1 F_{1000}^2 F_{0011}^2
--1 F_{0020} F_{1000}^2 F_{0002}
-\\ &
-+1 F_{0010}^2 F_{1001}^2
-+1 F_{0001}^2 F_{1010}^2
-\\ &
-\\ &
-+8 F_{0101} F_{0110} F_{0011}
-\\ &
-+4 F_{0200} F_{0020} F_{0002}
--4 F_{0200} F_{0011}^2
-\\ &
--4 F_{0110}^2 F_{0002}
--4 F_{0020} F_{0101}^2
-\\ &
-+2 F_{0200} F_{0010} F_{0001} F_{0011}
--2 F_{0110} F_{0010} F_{0001} F_{0101}
-\\ &
--2 F_{0110} F_{0001} F_{0100} F_{0011}
-+2 F_{0101} F_{0001} F_{0100} F_{0020}
-\\ &
-+2 F_{0010} F_{0110} F_{0100} F_{0002}
--2 F_{0010} F_{0101} F_{0100} F_{0011}
-\\ &
--1 F_{0200} F_{0010}^2 F_{0002}
--1 F_{0200} F_{0001}^2 F_{0020}
-\\ &
-+1 F_{0100}^2 F_{0011}^2
--1 F_{0020} F_{0100}^2 F_{0002}
-\\ &
-+1 F_{0010}^2 F_{0101}^2
-+1 F_{0001}^2 F_{0110}^2
-\end{aligned}
-\end{equation}
 

### 2025-03-08 paper/main.tex
Lots of changes by CH

@@
 % -----------
-% - CH: Include the d=3, d=4 solutions and comments on d>4.
-% - CH: Rewrite to remove the Hogg solution.
 % - CH: Show that there is NOT an exact solution at polynomial degree 2 order.
-% - CH: Needs a clear tikz figure showing the corners and notation. Like an infographic.
 % - HOGG: Discuss alternatives like Novara's; And alternatives like Kaze's.
@@
 \usetikzlibrary{angles,calc,quotes}
+
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
 
@@
 
-\usepackage{natbib}
-\usepackage{aas_macros}
+\renewcommand{\paragraph}[1]{\par\addvspace{1.5ex}\noindent\textbf{#1}~---}
+
+\usepackage{natbib,setspace,aas_macros}
 \bibliographystyle{aasjournal}
 \setcitestyle{aysep={}}
-\usepackage{setspace}
 \setlength{\bibsep}{0pt plus 0.3ex}
@@
 \usepackage{xcolor}
-\newcommand{\CH}[1]{{\color{blue} (CH: #1)}}
+\newcommand{\CH}[1]{{\color{blue} (CH says: #1)}}
+\newcommand{\Hogg}[1]{{\color{violet} (Hogg says: #1)}}
+\newcommand{\pseudosection}[1]
+{\noindent{\color{green!66!black} (\textit{#1})}\par}
+
+% define the affiliation symbols
+\newcommand{\aOne}{\textsuperscript{\textasteriskcentered}}
+\newcommand{\aTwo}{\textsuperscript{\textdagger}}
+\newcommand{\aThree}{\textsuperscript{\textdaggerdbl}}
+\newcommand{\aFour}{\textsuperscript{\textsection}}
+\newcommand{\aFive}{\textsuperscript{\textbardbl}}
+\newcommand{\aSix}{\textsuperscript{\textparagraph}}
 
@@
 A new formula for the area of a triangle:
-Useless, but explicitly in Deep Sets form\footnote{%
-The authors thank Alexander Novara (NYU), Ben Wandelt (Sorbonne), and Kaze Wong (JHU) for valuable discussions.}}
+Useless, but explicitly in Deep Sets form}%\footnote{%
+%The authors thank Alexander Novara (NYU), Ben Wandelt (JHU), Kaze Wong (JHU), and the Blanton--Hogg group meeting at NYU for valuable discussions.}}
 
-\noindent \textbf{Connor Hainje},\footnote{Email at \href{mailto:connor.hainje@nyu.edu}{connor.hainje@nyu.edu}.
+\noindent 
+\textbf{Connor Hainje}\aOne\aSix,
+\textbf{David W. Hogg}\aOne\aTwo\aThree, and
+\textbf{Soledad Villar}\aFour\aFive
+
+\vspace{0.5em}
+{\footnotesize\par\noindent \aOne\textsl{
+Center for Cosmology and Particle Physics, Department of Physics, New York University}}
+{\footnotesize\par\noindent \aTwo\textsl{
+Max-Planck-Insitut f\"ur Astronomie}}
+{\footnotesize\par\noindent \aThree\textsl{
+Center for Computational Astrophysics, Flatiron Institute}}
+{\footnotesize\par\noindent \aFour\textsl{
+Department of Applied Mathematics and Statistics, Johns Hopkins University}}
+{\footnotesize\par\noindent \aFive\textsl{
+Mathematical Institute for Data Science, Johns Hopkins University}}
+{\footnotesize\par\noindent \aSix\textsl{
+Email:} \texttt{\href{mailto:connor.hainje@nyu.edu}{connor.hainje@nyu.edu}}}
+{\footnotesize\par}  % needed to prevent some weird spacing issue
+
+% \noindent \textbf{Connor Hainje},\footnote{Email at \href{mailto:connor.hainje@nyu.edu}{connor.hainje@nyu.edu}.
 CH is at the Center for Cosmology and Particle Physics, Department of Physics, New York University.}
-\noindent \textbf{David W.
+% \noindent \textbf{David W.
 Hogg},\footnote{DWH is at the Center for Cosmology and Particle Physics, Department of Physics, New York University; the Max-Planck-Insitut f\"ur Astronomie; and the Center for Computational Astrophysics, Flatiron Institute.
 The Flatiron Institute is a division of the Simons Foundation.} and
-\textbf{Soledad Villar}\footnote{SV is at the Department of Applied Mathematics and Statistics, Johns Hopkins University; and the Mathematical Institute for Data Science, Johns Hopkins University.}
+% \textbf{Soledad Villar}\footnote{SV is at the Department of Applied Mathematics and Statistics, Johns Hopkins University; and the Mathematical Institute for Data Science, Johns Hopkins University.}
 
-\paragraph{Abstract:}
+\bigskip
+\paragraph{Abstract}
 It is known that any permutation-invariant function of data points $r_i$
 can be written in the form $\rho(\sum_i\phi(r_i))$,
-where $\rho$ and $\phi$ are (probably nonlinear) functions, and the function $\phi$ might output a high-dimensional latent encoding of each point $r_i$.
-This form---known in the machine-learning literature as Deep Sets---is also a map--reduce, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically or in any order.
+where $\rho$ and $\phi$ are functions.
+%, and the function $\phi$ might output a high-dimensional latent encoding of each point $r_i$.
+This form---known in the machine-learning literature as Deep Sets---is also a map--reduce algorithm, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically or in any order.
 The area of a triangle is a permutation-invariant function of the locations $r_i$ of the three corners $1\leq i\leq 3$.
-We find, possibly for the first time, the formula for the area of a triangle, given the positions of the three corners, that is explicitly in Deep Sets form.
-This project was motivated by questions about the fundamental computational complexity of n-point statistics in cosmology; that said, no insights of any kind were gained from these results.
+We find, possibly for the first time, the formula for the area of a triangle, given the Cartesian coordinates of the three corners, that is explicitly in Deep Sets form.
+This project was motivated by questions about the fundamental computational complexity of $n$-point statistics in cosmology; that said, no insights of any kind were gained from these results.
 
-\section{Introduction}
+\bigskip
+{\centering $\ast~\ast~\ast$ \par}
+\bigskip
 
-\begin{figure}
+\pseudosection{Areas}
+\noindent
+The area $\Delta$ of a triangle can be written in many ways.
+Here are a few:
+\begin{align}
+\Delta 
+&= \frac{1}{2} \, \text{(base)} \, \text{(height)} \label{eq:school} \\
+&= \frac{1}{2}\,a\,b\,\sin(C) \label{eq:sine} \\
+&= \frac{1}{2}\, \left| \vec{a} \times \vec{b} \right|
+    = \frac{1}{2}\, \left| \vec{a} \wedge \vec{b} \right| \label{eq:cross} \\
+&= \sqrt{s\,(s-a)\,(s-b)\,(s-c)} \label{eq:Heron} ~,
+\end{align}
+where the quantities used are defined in Figure~\ref{fig:triangle}, apart from the side lengths of the triangle
+$a = | \vec{a} |$, $b = | \vec{b} |$, $c = | \vec{c} |$
+and the semi-perimeter $s = (a + b + c)/2$.
+\eqref{eq:school} is the usual introduction to the subject.
+% \eqref{eq:sine} is the side-angle-side formula.
+\eqref{eq:cross} generalizes to triangles in higher dimensional spaces, using the fact that the cross/wedge product has magnitude equal to the area of the parallelogram spanned by $\vec{a}$ and $\vec{b}$, which is twice the area of the triangle with $\vec{a}$ and $\vec{b}$ as sides.
+Lastly, \eqref{eq:Heron} is Heron's formula, which elegantly computes the area from the side lengths alone.
+For a collection of 110 unique triangle area formulae, see \citet{baker1885a,baker1885b}.
+
+\begin{figure}[t!]
     \centering
@@
     \coordinate (O) at (0,0);
-    \coordinate (A) at (3,4);
-    \coordinate (B) at (5,1);
+    \coordinate (A) at (3.5,4);
+    \coordinate (B) at (6,1);
 
@@
     \draw[->, line width=0.4mm]
-        ($(B)!\pad!(A)$) -- ($(A)!\pad!(B)$)
+        ($(A)!\pad!(B)$) -- ($(B)!\pad!(A)$)
         node[midway, right, xshift=+1mm, yshift=+1mm] {$\vec{c}$};
@@
 
-In Figure~\ref{fig:triangle}, we present a triangle.
-The area $\Delta$ of this triangle can be written in many ways.
-Here are a few:
-\begin{align}
-\Delta 
-&= \frac{1}{2} \, \text{(base)} \, \text{(height)} \label{eq:school} \\
-&= \frac{1}{2}\, \| \vec{a} \times \vec{b} \| \label{eq:cross} \\
-&= \frac{1}{2}\,a\,b\,\sin(C) \label{eq:sine} \\
-&= \sqrt{s\,(s-a)\,(s-b)\,(s-c)} \label{eq:Heron} ~,
-\end{align}
-where the first formula \eqref{eq:school} is the usual introduction to the subject,
-$\times$ represents the cross-product operator,
-$a = \| \vec{a} \|$, $b = \| \vec{b} \|$, $c = \| \vec{c} \|$ are side lengths of the triangle,
-% $C$ is the angle between $\vec{a}$ and $\vec{b}$,
-and $s$ is the semi-perimeter $(a + b + c)/2$.
-For a rather exhaustive list of 110 unique formulae, see \citet{baker1885a,baker1885b}.
+While the area of the triangle does not depend on how its sides or corners are labeled,
+only Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant.
 
-While the area of the triangle does not depend on how its sides or corners are labeled,
-only the last of these formulae---Heron's formula \eqref{eq:Heron}---is explicitly permutation-invariant.
+\CH{struggling to find a segue here} \Hogg{We could call out symmetries: These formulae are for a quantity that obeys permutation, translation, rotation, and reflection symmetries.
+But these symmetries aren't always visible.
+Then go to the point (above) about Heron's formula.}
 
-Permutation invariance is an important symmetry in a number of fields for problems defined on \textit{sets}.
-In machine learning, one area of relevance is deep learning on point clouds (e.g., sets of three-dimensional position vectors) \citep{guo2020pointclouds}, which is particularly useful for problems such as 3D shape classification and object detection.
-In physics, many fundamental problems are invariant to permutation.
-For example, the gravitational/electric potential is invariant to the labeling of point masses/charges \CH{citation needed?}.
-Further, permutation invariance lies at the heart of all modern cosmological statistics, such as the power spectrum and the bispectrum \CH{cite}.
-
-There is a simple and powerful result which is useful for constructing and characterizing permutation-invariant functions.
-Given a function $f(r_1, \cdots, r_N)$, Theorem 2 of \citet{zaheer+17deepsets} states that $f$ is invariant to permutation of its arguments if and only if it can be decomposed in the form
+\pseudosection{Deep Sets} There is a simple result that is useful for constructing, characterizing, or learning permutation-invariant functions:
+given a function $f(r_1, \cdots, r_N)$, Theorem 2 of \citet{zaheer+17deepsets} states that $f$ is invariant to permutation of its arguments if and only if it can be decomposed in the form
 \begin{equation}
@@
 \end{equation}
-for some functions $\rho$ and $\phi$.
+for some functions $\rho$ and $\phi$. This result---often known by the name of the deep learning architecture derived from it, \emph{Deep Sets} \citep{zaheer+17deepsets}---is not at all obvious.
+For example, even though Heron's formula \eqref{eq:Heron} is permutation-invariant, it is not actually in the form of \eqref{eq:DeepSets}.
 
-This result---often known by the name of the deep learning architecture derived from it, \emph{Deep Sets} \citep{zaheer+17deepsets}---is not at all obvious:
-For example, even though Heron's formula \eqref{eq:Heron} is permutation-invariant, it is not actually in the form of \eqref{eq:DeepSets}. Consider instead the formula
-\begin{equation}
-    \Delta = \frac{1}{4} \sqrt{
-        \left( a^2 + b^2 + c^2 \right)^2
-        - 2 \left( a^4 + b^4 + c^4 \right)
-    }.
-\end{equation}
-It is simple to show that this formula is exactly equivalent to Heron's formula when expanded, but it is now precisely in Deep Sets form. Explicitly,
-\begin{equation}
-    \phi(x) = (x^2, x^4), \quad
-    \vec{u} = {\textstyle \sum_{x \in \{ a, b, c \}}} \, \phi(x), \quad
-    \Delta = \rho(\vec{u}) = \frac{1}{4} \sqrt{ u_1^2 - 2 \, u_2 }.
-\end{equation}
+% A few comments about this Deep Sets form \eqref{eq:DeepSets} we'd make are the following:
+% \begin{itemize}
+%     \item The function $\phi$ is universal to all the elements $r_i$; its form won't depend on label $i$ or on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
+%     \item The point properties $r_i$ and the output of the function $\phi$ might both be very large.
+%     That is, each point $r_i$ might be a vector or even an ordered list of vectors and tensors and images and so on, and the output of the function $\phi$ might also be very large and complicated.
+%     The output of the function $\phi$ is some kind of informative, latent encoding of the point $r_i$.
+%     \item If a function can be written in Deep Sets form \eqref{eq:DeepSets}, then that function can be executed in a map--reduce framework, in which the function $\phi$ is executed independently and asynchronously on each data point (maybe even in the data center) and the sums are performed hierarchically or in any order.
+%     That means that for large problems executed on large data sets in large data centers, problems in this form \eqref{eq:DeepSets} can often be performed in wall-clock time that scales as $\log N$.
+% \end{itemize}
 
-A few comments about this Deep Sets form \eqref{eq:DeepSets} we'd make are the following:
-\begin{itemize}
-    \item The function $\phi$ is universal to all the elements $r_i$; its form won't depend on label $i$ or on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
-    \item The point properties $r_i$ and the output of the function $\phi$ might both be very large.
-    That is, each point $r_i$ might be a vector or even an ordered list of vectors and tensors and images and so on, and the output of the function $\phi$ might also be very large and complicated.
-    The output of the function $\phi$ is some kind of informative, latent encoding of the point $r_i$.
-    \item If a function can be written in Deep Sets form \eqref{eq:DeepSets}, then that function can be executed in a map--reduce framework, in which the function $\phi$ is executed independently and asynchronously on each data point (maybe even in the data center) and the sums are performed hierarchically or in any order.
-    That means that for large problems executed on large data sets in large data centers, problems in this form \eqref{eq:DeepSets} can often be performed in wall-clock time that scales as $\log N$.
-\end{itemize}
+\pseudosection{Computational motivations}
 
-Although the contribution of this paper is minimal, or even zero, it was inspired by real problems in physics.
-For one very specific example, all known exact formulae for the gravitational potential energy of a system of $N$ point masses involve sums over pairs of points (and also require taking differences of points); they have not ever been written in Deep Sets form \eqref{eq:DeepSets} to our knowledge.
-CH: THAT SAID, THERE ARE MULTIPOLE-LIKE EXPRESSIONS THAT MIGHT BE IN DEEPSETS AND APPROXIMATE THIS CALCULATION; SEE THE DISCUSSION?
-In cosmology, standard estimators for $k$-point functions (correlation functions, bispectra, and so on) involve sums over $k$-tuples; once again they have not been written in the Deep Sets form.
-CH: DO THE PHILCOX PAPERS ADDRESS THIS PREVIOUS POINT?
-Importantly, when these statistics are computed in their naive forms, they are very time-consuming because they involve sums over all elements of $N$-choose-$k$, which scales (badly) as $N^k$.
-Because 3-point functions involve the shapes and areas of triangles, consideration of the 3-point function led to the question answered in this \documentname:
-\emph{What is the Deep Sets (or map--reduce) expression for the area of a triangle?}
-We deliver in \secref{sec:result}.
+Although the contribution of this paper is minimal, or even zero, it was inspired by real problems in machine learning and physics. \CH{another bad segue}
 
-\section{A note on our method}\label{sec:method}
-How did we find the formula \eqref{eq:result} presented below in \secref{sec:result}?
-Fundamentally, our method took the form of a symbolic regression, followed by a confirmation with computational algebra.
+Permutation invariance is important for problems defined on \emph{sets}.
+In machine learning, one area of relevance is deep learning on point clouds (\textit{e.g.,} sets of three-dimensional position vectors) \citep{guo2020pointclouds}, which is used for problems such as three-dimensional shape classification and object detection.
 
-For the symbolic regression, we considered a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
-We fit a linear combination of the monomials to a data set of generated triples of points, drawn from a two-dimensional Gaussian whose mean was also drawn from a Gaussian in a two-dimensional space.
-For each triple we attach a label, which is the squared area, and we require the linear combination of fourth-order, Deep Sets monomials to reproduce the labels to machine precision.
-The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
+\pseudosection{Cosmology motivations}
 
-We choose to work with squared area $A^2$ and not area $A$ because, geometrically, areas are signed quantities.
-Indeed, the formulae \eqref{eq:school} through \eqref{eq:Heron} either (implicitly) involve square roots of polynomials, or else absolute value signs, such that the corresponding pure polynomial expression would be for the squared area.
-We restricted to monomials that contain four powers of positional coordinates because any expression for the squared area must be units covariant (equivariant to changes of the units used for the measurements of positions or lengths).
+In physics, many fundamental problems are invariant to permutation.
+For example, the gravitational/electric potential is invariant to the labeling of point masses/charges \CH{citation needed?}.
+Further, and important to our motivations, cosmological clustering statistics such as 
+the correlation function \citep{},
+the power spectrum \citep{},
+the bispectrum \citep{},
+and higher-order statistics \citep{}
+used for precision measurement in cosmology \citep{}
+are all permutation-invariant.
+Na\"ively, k-point statistics involve sums that scale as $N^k$, where $N$ is the number of points in the set.
 
-Perhaps non-obviously, for a monomial to be in Deep Sets form, it must consist only of products of sums over the three points.
-It cannot contain any differences, or any sums over pairs of points (or edges).
-Thus $(\sum_{i=1}^2 x_i^2)\,(\sum_{i=1}^3 x_i\,y_i)$ is a permitted monomial, but $(\sum_{i=1}^3\sum_{j=1}^3 x_i^2\,y_j^2)$ is not a permitted monomial.
-These monomials are both correct from a units perspective, but the second is not in Deep Sets form, because it contains products of position coordinates from different points. 
-All the possible monomials are made of permutation-invariant factors of the form 
-\begin{equation}
-F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n.
-\end{equation}
-The invariant monomials of degree $d$ are then all products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
+\pseudosection{Scalings}
 
-In detail, the formula for the area (or area squared) of a triangle must be invariant to permutations, translations, rotations, and reflections.
-The Deep Sets form enforces only permutation-invariance.
-Some of the monomials are invariant to reflections, but no individual monomial is invariant to translations or rotations.
-It is left as an exercise to the reader to show that our solution \eqref{eq:result} below is invariant to translations, rotations, and reflections.
+Na\"ively, Deep Sets form \eqref{eq:DeepSets} seems to imply that the function $f$ can be computed in linear time. 
+For problems like the computation of the gravitational potential for a system of point masses, all known solutions scale as the square of the number of points, and yet they are permutation-invariant functions and can thus be put in Deep Sets form.
+How can this be?
+There is no free lunch.
+The resolution must be either that $\phi$ will be very large in dimension---its dimensionality may even depend directly on $N$---or else that the complexity of $rho$ may grow rapidly with $N$.
 
-The symbolic regression produces a linear combination of monomials with real-valued coefficients.
-We systematically search for an integer which, when multiplied by all of the coefficients, brings all coefficients very near integers.
-We then round the coefficients to these nearest integers and check that we reproduce $({\rm denom}) A^2$ to better precision than the least-square result.
+Still, Deep Sets form \eqref{eq:DeepSets} has a computational advantage in that it is explicitly in map--reduce form.
+Map--reduce is a model for parallelizing and distributed a calculation over a large dataset.
+In our case, $\phi$ is the ``map'':
+It can be executed independently and asynchronously on each data point.
+The sums are then the ``reduce'', which can be performed hierarchically across a data center such that the runtime of the sums scales only logarithmically with the number of points. 
+(Granted, a triangle will never have more than three points!)
+
+% HOGG TO CH: THIS PARAGRAPH REPEATS CONTENT FROM ABOVE. COMBINE HERE I THINK; THAT IS, MOVE IT DOWN TO HERE, BELOW DEEP SETS.
+% For one very specific example, all known exact formulae for the gravitational potential energy of a system of $N$ point masses involve sums over pairs of points (and also require taking differences of points); they have not ever been written in Deep Sets form \eqref{eq:DeepSets} to our knowledge.
+% CH: THAT SAID, THERE ARE MULTIPOLE-LIKE EXPRESSIONS THAT MIGHT BE IN DEEPSETS AND APPROXIMATE THIS CALCULATION; SEE THE DISCUSSION?
+% In cosmology, standard estimators for $k$-point functions (correlation functions, bispectra, and so on) involve sums over $k$-tuples; once again they have not been written in the Deep Sets form.
+% CH: DO THE PHILCOX PAPERS ADDRESS THIS PREVIOUS POINT?
+% Importantly, when these statistics are computed in their naive forms, they are very time-consuming because they involve sums over all elements of $N$-choose-$k$, which scales (badly) as $N^k$.
+% Because 3-point functions involve the shapes and areas of triangles, consideration of the 3-point function led to the question answered in this \documentname:
+% \emph{What is the Deep Sets (or map--reduce) expression for the area of a triangle?}
+% We deliver in \secref{sec:result}.
+
+% \section{A note on our method}\label{sec:method}
+% How did we find the formula \eqref{eq:result} presented below in \secref{sec:result}?
+% Fundamentally, our method took the form of a symbolic regression, followed by a confirmation with computational algebra.
+% CH THIS ISN'T TRUE ANY MORE?? OR IS IT?
+
+% For the symbolic regression, we considered a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
+% We fit a linear combination of the monomials to a data set of generated triples of points, drawn from a two-dimensional Gaussian whose mean was also drawn from a Gaussian in a two-dimensional space.
+% For each triple we attach a label, which is the squared area, and we require the linear combination of fourth-order, Deep Sets monomials to reproduce the labels to machine precision.
+% The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
+
+% We choose to work with squared area $A^2$ and not area $A$ because, geometrically, areas are signed quantities.
+% Indeed, the formulae \eqref{eq:school} through \eqref{eq:Heron} either (implicitly) involve square roots of polynomials, or else absolute value signs, such that the corresponding pure polynomial expression would be for the squared area.
+% We restricted to monomials that contain four powers of positional coordinates because any expression for the squared area must be units covariant (equivariant to changes of the units used for the measurements of positions or lengths).
+
+% Perhaps non-obviously, for a monomial to be in Deep Sets form, it must consist only of products of sums over the three points.
+% It cannot contain any differences, or any sums over pairs of points (or edges).
+% Thus $(\sum_{i=1}^2 x_i^2)\,(\sum_{i=1}^3 x_i\,y_i)$ is a permitted monomial, but $(\sum_{i=1}^3\sum_{j=1}^3 x_i^2\,y_j^2)$ is not a permitted monomial.
+% These monomials are both correct from a units perspective, but the second is not in Deep Sets form, because it contains products of position coordinates from different points. 
+% All the possible monomials are made of permutation-invariant factors of the form 
+% \begin{equation}
+% F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n.
+% \end{equation}
+% The invariant monomials of degree $d$ are then all products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
+
+% The symbolic regression produces a linear combination of monomials with real-valued coefficients.
+% We systematically search for an integer which, when multiplied by all of the coefficients, brings all coefficients very near integers.
+% We then round the coefficients to these nearest integers and check that we reproduce $({\rm denom}) A^2$ to better precision than the least-square result.
 % Trial-and-error delivered the integer 368, which makes all of the coefficients obviously rational.
@@
 
-\section{Result and discussion}\label{sec:result}
+\pseudosection{Result}
 
-Our result is that the area $A$ of a triangle is given by
-% \begin{align}
-% A^2 = &
-% -\frac{ 54}{368}\,(\sum_{i=1}^3 x_i^2\,y_i^2) 
-% -\frac{258}{368}\,(\sum_{i=1}^3 x_i\,y_i)^2 \nonumber \\ &
-% +\frac{285}{368}\,(\sum_{i=1}^3 x_i^2)\,(\sum_{i=1}^3 y_i^2) 
-% +\frac{ 36}{368}\,(\sum_{i=1}^3 y_i)\,(\sum_{i=1}^3 x_i^2\,y_i) \nonumber \\ &
-% +\frac{ 36}{368}\,(\sum_{i=1}^3 x_i)\,(\sum_{i=1}^3 x_i\,y_i^2) 
-% -\frac{101}{368}\,(\sum_{i=1}^3 y_i)^2\,(\sum_{i=1}^3 x_i^2) \nonumber \\ &
-% +\frac{148}{368}\,(\sum_{i=1}^3 x_i)\,(\sum_{i=1}^3 y_i)\,(\sum_{i=1}^3 x_i\,y_i) 
-% -\frac{101}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i^2) \nonumber \\ &
-% +\frac{  9}{368}\,(\sum_{i=1}^3 x_i)^2\,(\sum_{i=1}^3 y_i)^2 ~, \label{eq:result}
-% \end{align}
-\begin{align}
-    A^2 = & \
-    \frac{2}{4} \, (\sum_{i=1}^{3} x_i \, y_i ) \, (\sum_{i=1}^{3} x_i ) \, (\sum_{i=1}^{3} y_i )
-    \nonumber \\ &
-    + \frac{3}{4} \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i^2 )
-    - \frac{3}{4} \, (\sum_{i=1}^{3} x_i \, y_i )^2.
-    \nonumber \\ &
+Our primary result is that area $\Delta$ of a triangle can be written in Deep Sets form as
+\begin{multline} \label{eq:result}
+    \Delta^2 = \
+    \frac{3}{4} \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i^2 )
+    - \frac{3}{4} \, (\sum_{i=1}^{3} x_i \, y_i )^2 
+    + \frac{2}{4} \, (\sum_{i=1}^{3} x_i \, y_i ) \, (\sum_{i=1}^{3} x_i ) \, (\sum_{i=1}^{3} y_i )
+    \\
     - \, \frac{1}{4} \, (\sum_{i=1}^{3} y_i^2 ) \, (\sum_{i=1}^{3} x_i )^2
     - \, \frac{1}{4} \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i )^2
-     ~, \label{eq:result}
-\end{align}
+     ~,
+\end{multline}
 where the two-dimensional position of each of the three corners $i$ is $r_i \equiv (x_i, y_i)$.
-This answer is in Deep Sets form:
-Every expression in parentheses in \eqref{eq:result} is a pure sum over all three points and can be written in the form of $F_{mn}$ for some choice of $m$ and $n$.
-There are no sums over differences, and there is no ordering of the points.
-Being very explicit, this is written in the Deep Sets form \eqref{eq:DeepSets} for $\rho$ and $\phi$ defined as follows:
+We verified this equation by using computational algebra to expand the formula and check its equivalence to \eqref{eq:cross}.
+
+This formula is in Deep Sets form, since
+every expression in parentheses in \eqref{eq:result} is a pure sum over all three points.
+% There are no sums over differences, and there is no ordering of the points.
+Being very explicit, it matches \eqref{eq:DeepSets} if we define $\rho$ and $\phi$ as follows:
 \begin{gather}
@@
     \rho \left( {\textstyle \sum_i} \phi(\vec{r}_i) \right)
-    = \tfrac{2}{4} \, F_{11} \, F_{10} \, F_{01}
-    + \tfrac{3}{4} \, F_{20} \, F_{02}
+    = \tfrac{3}{4} \, F_{20} \, F_{02}
     - \tfrac{3}{4} \, F_{11}^2
+    + \tfrac{2}{4} \, F_{11} \, F_{10} \, F_{01}
     - \tfrac{1}{4} \, F_{02} \, F_{10}^2
@@
 
-% HOGG: Explicit vector $\phi()$ and explicit expression $\rho()$.
-
-% HAINJE: Here it is:
-% \begin{gather}
-%     \phi(r_i) = (x_i, y_i, x_i\,y_i, x_i^2, y_i^2, x_i\,y_i^2, x_i^2\,y_i, x_i^2\,y_i^2) \\
-%     \begin{aligned}
-%     \rho(\textstyle\sum_{i=1}^{3}\phi(r_i)\equiv \vec{u}) = &
-%     - \tfrac{54}{368} u_8
-%     - \tfrac{258}{368} u_3^2
-%     + \tfrac{285}{368} u_4\,u_5
-%     + \tfrac{36}{368} u_2\,u_7
-%     + \tfrac{36}{368} u_1\,u_6 \\ &
-%     - \tfrac{101}{368} u_2^2\,u_4
-%     + \tfrac{148}{368} u_1\,u_2\,u_3
-%     - \tfrac{101}{368} u_1^2\,u_5
-%     + \tfrac{9}{368} u_1^2\,u_2^2.
-%     \end{aligned}
-% \end{gather}
-
-% Note the symmetries....! See the $36/368$ terms, or the $-101/368$ terms
-
-% But note that it isn't obviously invariant to arbitrary rotations.
-% And it isn't obviously invariant to translations.
-% Formulae \eqref{eq:cross}, \eqref{eq:sine}, and \eqref{eq:Heron} are obviously rotation-invariant and translation-invariant.
+\pseudosection{A note on symbolic regression}
 
@@
 Multiplying both sides by 4 then reveals that \eqref{eq:result} is an integer polynomial of invariant factors for the area of the parallelogram.
+Using this fact, we have written a program to perform a symbolic regression which recovers our result.
+
+For the symbolic regression, we considered a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
+All the possible monomials are made of permutation-invariant factors of the form 
+\begin{equation}
+F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n.
+\end{equation}
+Such a factor has degree $m + n$.
+The invariant monomials of degree $d$ are then products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
+We generated all factors $F_{mn}$ with degree $1 \leq m + n \leq 4$, then formed all products of these factors with total degree four.
+However, these monomials are not all linearly-independent;
+the five monomials made of a single degree-four factor ($F_{40}$, $F_{31}$, $F_{22}$, $F_{13}$, $F_{04}$) can be written as linear combinations of the others and are thus removed.
+This leaves \CH{number?} monomials.
+
+We fit a linear combination of the monomials to a data set of generated triples of points, drawn from a two-dimensional Gaussian whose mean was also drawn from a Gaussian in a two-dimensional space.
+Each triple defines a parallelogram whose squared area we use as a label for regression.
+We then perform least-squares regression to fit a linear combination of the invariant monomials.
+We require that the labels are reproduced to machine precision after rounding the coefficients to the nearest integers.
+% For each triple we attach a label, which is the squared area of the parallelogram spanned by the two difference vectors \CH{?}, and we require the linear combination of fourth-order, Deep Sets monomials to reproduce the labels to machine precision with integer coefficients.
+The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
+
+\pseudosection{Generalizations: higher dimensions}
 
@@
 \begin{equation}
-    A_{\rm 3D}^2 =
-    A_{12}^2 + A_{13}^2 + A_{23}^2,
+    \Delta_{\rm 3D}^2 =
+    \Delta_{12}^2 + \Delta_{13}^2 + \Delta_{23}^2,
 \end{equation}
@@
 \begin{equation}
-    A_{\rm 4D}^2 =
-    A_{12}^2 + A_{13}^2 + A_{14}^2
-    + A_{23}^2 + A_{24}^2
-    + A_{34}^2,
+    \Delta_{\rm 4D}^2 =
+    \Delta_{12}^2 + \Delta_{13}^2 + \Delta_{14}^2
+    + \Delta_{23}^2 + A_{24}^2
+    + \Delta_{34}^2,
 \end{equation}
-where we have introduced $A_{jk}$, the area of the triangle formed by just the $j^{\rm th}$ and $k^{\rm th}$ components of the position vectors.
-By using \eqref{eq:result} for each of the two-dimensional sub-areas, each can be computed in a permutation-invariant way and the entire result is thus permutation invariant (and in Deep Sets form).
+where we have introduced $\Delta_{jk}$, the area of the triangle formed by just the $j^{\rm th}$ and $k^{\rm th}$ components of the position vectors.
+By using \eqref{eq:result} for each of the two-dimensional sub-areas, each can be computed in a permutation-invariant way and the entire result is thus permutation invariant and in Deep Sets form. 
+\Hogg{This last result somehow is related to the meaning of the square of the magnitude of a bivector.}
+
+\pseudosection{Generalizations: simplex volume}
 
 Our method also extends to the volume of a simplex. In three dimensions, four vectors $\{ \vec{r}_1, \vec{r}_2, \vec{r}_3, \vec{r}_4 \}$ determine the vertices of a simplex. 
-Like the area of a triangle is related to the area of a parallelogram, the volume of a simplex is one-sixth the volume of the parallelepiped spanned by the vectors $\{ \vec{r}_2 - \vec{r}_1, \vec{r}_3 - \vec{r}_1, \vec{r}_4 - \vec{r}_1 \}$.
-Hoping for an integer solution once again, we modify our symbolic regression to search for polynomials of the invariant factors which yield the volume of the parallelepiped.
-This gives an analogous Deep Sets formula for the volume of a simplex to be
+Like the area of a triangle is one-half the area of a related parallelogram, the volume of a simplex is one-sixth the volume of the parallelepiped spanned by the vectors $\{ \vec{r}_2 - \vec{r}_1, \vec{r}_3 - \vec{r}_1, \vec{r}_4 - \vec{r}_1 \}$.
+
+Expecting an integer solution once again, we modified our symbolic regression to generate 4-tuples of points, using the squared volumes of the parallelepipeds they produce as labels. 
+The invariant factors then have three indices, $F_{\ell m n} = \sum_i x_i^\ell \, y_i^m \, z_i^n$, and we form all invariant monomials of degree six.
+The result, which we have verified by computational algebra, is that the volume of the simplex $V$ is given by
 \begin{align}
@@
 
+\pseudosection{Generalizations: Heron symmetrized}
+
+We have stated above that Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, and yet it is not represented in Deep Sets form. Consider instead
+\begin{equation}
+\label{eq:HeronDeepSets}
+    \Delta = \frac{1}{4} \, \sqrt{
+        \left( a^2 + b^2 + c^2 \right)^2
+        - 2 \left( a^4 + b^4 + c^4 \right)
+    }.
+\end{equation}
+It is simple to show that this is exactly equivalent to Heron's formula when expanded, but it is now in Deep Sets form. Explicitly,
+\begin{equation}
+    \phi(x) = (x^2, x^4), \quad
+    \vec{u} = {\textstyle \sum_{x \in \{ a, b, c \}}} \, \phi(x), \quad
+    \Delta = \rho(\vec{u}) = \frac{1}{4} \sqrt{ u_1^2 - 2 \, u_2 }.
+\end{equation}
+
+This is not novel. Heron's formula is a special case of Brahmagupta's formula, which gives the area $K$ of a convex cyclic quadrilateral to be \citep{coolidge1939}
+\begin{equation}
+    K = \sqrt{
+        (s - a) \,
+        (s - b) \,
+        (s - c) \,
+        (s - d)
+    }~,
+\end{equation}
+where $a, b, c, d$ are the four side lengths and $s = (a+b+c+d)/2$ is the semi-perimeter. This formula has been alternatively written as \CH{seems like someone just put this on Wikipedia?}
+\begin{equation}
+    K = \frac{1}{4} \, \sqrt{
+        \left( a^2 + b^2 + c^2 + d^2 \right)^2
+        + 8 \, a\,b\,c\,d
+        - 2 \left( a^4 + b^4 + c^4 + d^4 \right)
+    }~,
+\end{equation}
+which is exactly \eqref{eq:HeronDeepSets} if the fourth side-length $d$ is zero.
+Note that this expression is in Deep Sets form if the product $a \, b \, c \, d$ is taken to be the exponential of the sum of logarithms.
+
+\pseudosection{Discussion: other symmetries}
+
+The area of a triangle is not only invariant to permutation, but also to translation, rotation, and reflection.
+Formulae \eqref{eq:sine} and \eqref{eq:Heron}, by virtue of referencing only side-lengths and angles, are obviously invariant to these other symmetries.
+While true, it is not obvious that our solution \eqref{eq:result} is invariant. An explicit proof of this is left to the reader.
+
+\pseudosection{Discussion: sines and cosines}
+
+\pseudosection{Discussion: open questions}
+
 % HOGG SAY: CH: EXPAND THIS INTO AN OPEN QUESTIONS LIST:
@@
 
-\newpage
+\CH{conclude}
+
+The authors thank Alexander Novara (NYU), Ben Wandelt (JHU), Kaze Wong (JHU), and the Blanton--Hogg group meeting at NYU for valuable discussions.
+The authors would also like to note that, while we believe that this result is novel, it is difficult to be sure, given the impossibility of conducting a complete literature review for such an ancient topic.
+
+\bigskip
+{\centering $\ast~\ast~\ast$ \par}
+
+\renewcommand{\section}[2]{}%
 {\small\singlespacing\bibliography{main}}

### 2025-03-14 paper/main.tex
content from CH mainly and DWH also

@@
 \usepackage[hidelinks]{hyperref}
-\usepackage{amsmath,amsfonts,tikz}
+\usepackage{amsmath,amsfonts,mathtools,tikz}
 \usetikzlibrary{angles,calc,quotes}
@@
 \newcommand{\pseudosection}[1]
-{\noindent{\color{green!66!black} (\textit{#1})}\par}
+{}
+% {\noindent{\color{green!66!black} (\textit{#1})}\par}
 
@@
 A new formula for the area of a triangle:
-Useless, but explicitly in Deep Sets form}%\footnote{%
-%The authors thank Alexander Novara (NYU), Ben Wandelt (JHU), Kaze Wong (JHU), and the Blanton--Hogg group meeting at NYU for valuable discussions.}}
+Useless, but explicitly in Deep Sets form}
 
@@
 
-% \noindent \textbf{Connor Hainje},\footnote{Email at \href{mailto:connor.hainje@nyu.edu}{connor.hainje@nyu.edu}.
-CH is at the Center for Cosmology and Particle Physics, Department of Physics, New York University.}
-% \noindent \textbf{David W.
-Hogg},\footnote{DWH is at the Center for Cosmology and Particle Physics, Department of Physics, New York University; the Max-Planck-Insitut f\"ur Astronomie; and the Center for Computational Astrophysics, Flatiron Institute.
-The Flatiron Institute is a division of the Simons Foundation.} and
-% \textbf{Soledad Villar}\footnote{SV is at the Department of Applied Mathematics and Statistics, Johns Hopkins University; and the Mathematical Institute for Data Science, Johns Hopkins University.}
-
 \bigskip
@@
 %, and the function $\phi$ might output a high-dimensional latent encoding of each point $r_i$.
-This form---known in the machine-learning literature as Deep Sets---is also a map--reduce algorithm, because the $\phi$ executions can be performed asynchronously and the sums can be performed hierarchically or in any order.
-The area of a triangle is a permutation-invariant function of the locations $r_i$ of the three corners $1\leq i\leq 3$.
-We find, possibly for the first time, the formula for the area of a triangle, given the Cartesian coordinates of the three corners, that is explicitly in Deep Sets form.
+This form---known in the machine-learning literature as Deep Sets---generates a map--reduce algorithm, because the $\phi$ executions can be performed asynchronously and the sums can be performed in any order.
+The area of a triangle is a permutation-invariant function of the locations $\vec{r}_i$ of the three corners $1\leq i\leq 3$.
+We find, possibly for the first time, the formula for the area of a triangle that is explicitly in Deep Sets form.
 This project was motivated by questions about the fundamental computational complexity of $n$-point statistics in cosmology; that said, no insights of any kind were gained from these results.
@@
 and the semi-perimeter $s = (a + b + c)/2$.
-\eqref{eq:school} is the usual introduction to the subject.
+\eqref{eq:school} is the usual introduction to the subject, traceable back even to Euclid.
 % \eqref{eq:sine} is the side-angle-side formula.
@@
 
-    \pgfmathsetmacro{\a}{sqrt(3*3 + 4*4)}
-    \pgfmathsetmacro{\b}{sqrt(5*5 + 1*1)}
-    \pgfmathsetmacro{\c}{sqrt(2*2 + 3*3)}
+    \pgfmathsetmacro{\a}{sqrt(3.5*3.5 + 4*4)}
+    \pgfmathsetmacro{\b}{sqrt(6*6 + 1*1)}
+    \pgfmathsetmacro{\c}{sqrt((6-3.5)*(6-3.5) + (4-1)*(4-1))}
     \pgfmathsetmacro{\s}{(\a+\b+\c)/2}
@@
 
-    \pgfmathsetmacro{\DeltaX}{5 - 0}  % Bx - Ox
+    \pgfmathsetmacro{\DeltaX}{6 - 0}  % Bx - Ox
     \pgfmathsetmacro{\DeltaY}{1 - 0}  % By - Oy
@@
 
-While the area of the triangle does not depend on how its sides or corners are labeled,
-only Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant.
+The area of a triangle has many symmetries.
+It is invariant to translation, rotation, and reflection, as well as to the labeling of sides or corners.
+However, these symmetries are not always obvious from a given formula; for example, of those listed above, only Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant.
 
-\CH{struggling to find a segue here} \Hogg{We could call out symmetries: These formulae are for a quantity that obeys permutation, translation, rotation, and reflection symmetries.
+% \CH{struggling to find a segue here} \Hogg{We could call out symmetries: These formulae are for a quantity that obeys permutation, translation, rotation, and reflection symmetries.
 But these symmetries aren't always visible.
@@
 
-\pseudosection{Deep Sets} There is a simple result that is useful for constructing, characterizing, or learning permutation-invariant functions:
-given a function $f(r_1, \cdots, r_N)$, Theorem 2 of \citet{zaheer+17deepsets} states that $f$ is invariant to permutation of its arguments if and only if it can be decomposed in the form
+\pseudosection{Deep Sets}
+
+A function $f(r_1, \dots, r_N)$ is invariant to permutation of its arguments if and only if it can be decomposed in the form
 \begin{equation}
-    f(r_1, \cdots, r_N) = \rho \left( {\textstyle \sum_{i=1}^{N}} \, \phi(r_i) \right), \label{eq:DeepSets}
+    f(r_1, \dots, r_N) = \rho \left( {\textstyle \sum_{i=1}^{N}} \, \phi(r_i) \right), \label{eq:DeepSets}
 \end{equation}
-for some functions $\rho$ and $\phi$. This result---often known by the name of the deep learning architecture derived from it, \emph{Deep Sets} \citep{zaheer+17deepsets}---is not at all obvious.
+for some functions $\rho$ and $\phi$.
+This result---often known by the name of the deep learning architecture derived from it, \emph{Deep Sets} (\citealt{zaheer+17deepsets}; see Theorem~2)---is not at all obvious.
 For example, even though Heron's formula \eqref{eq:Heron} is permutation-invariant, it is not actually in the form of \eqref{eq:DeepSets}. 
-
-% A few comments about this Deep Sets form \eqref{eq:DeepSets} we'd make are the following:
-% \begin{itemize}
-%     \item The function $\phi$ is universal to all the elements $r_i$; its form won't depend on label $i$ or on any properties of each point, nor any properties of the overall collection of points (the \emph{point cloud}, as it were).
-%     \item The point properties $r_i$ and the output of the function $\phi$ might both be very large.
-%     That is, each point $r_i$ might be a vector or even an ordered list of vectors and tensors and images and so on, and the output of the function $\phi$ might also be very large and complicated.
-%     The output of the function $\phi$ is some kind of informative, latent encoding of the point $r_i$.
-%     \item If a function can be written in Deep Sets form \eqref{eq:DeepSets}, then that function can be executed in a map--reduce framework, in which the function $\phi$ is executed independently and asynchronously on each data point (maybe even in the data center) and the sums are performed hierarchically or in any order.
-%     That means that for large problems executed on large data sets in large data centers, problems in this form \eqref{eq:DeepSets} can often be performed in wall-clock time that scales as $\log N$.
-% \end{itemize}
+Further, despite the triangle area being permutation-invariant, there is no existing formula for the area of triangle which is explicitly in this form.
 
@@
 
-Although the contribution of this paper is minimal, or even zero, it was inspired by real problems in machine learning and physics. \CH{another bad segue}
+% Although the contribution of this paper is minimal, or even zero, it was inspired by real problems in machine learning and physics. \CH{another bad segue}
 
-Permutation invariance is important for problems defined on \emph{sets}.
-In machine learning, one area of relevance is deep learning on point clouds (\textit{e.g.,} sets of three-dimensional position vectors) \citep{guo2020pointclouds}, which is used for problems such as three-dimensional shape classification and object detection.
+Although such a formula is of minimal, or even zero, practical interest, our search is motivated by real problems in machine learning and physics.
+For example, in machine learning, one area of relevance is deep learning on point clouds (e.g. sets of three-dimensional position vectors) \citep{guo2020pointclouds}.
+Such methods are used for problems such as three-dimensional shape classification and object detection.
 
@@
 
-In physics, many fundamental problems are invariant to permutation.
-For example, the gravitational/electric potential is invariant to the labeling of point masses/charges \CH{citation needed?}.
-Further, and important to our motivations, cosmological clustering statistics such as 
+The laws of classical physics are generally invariant to permutation.
+For example, the gravitational or electric potential energies of a set of point masses or point charges is invariant to the ordering or labeling of the points.
+Further, and important to our own motivations, cosmological clustering statistics such as 
 the correlation function \citep{},
@@
 are all permutation-invariant.
-Na\"ively, k-point statistics involve sums that scale as $N^k$, where $N$ is the number of points in the set.
+All of these statistics na\"ively involve loops over all $k$-tuples of points, which not only is inconsistent with Deep Sets form, but which has a runtime that scales as $N^k$, where $N$ the number of points in the set.
 
@@
 
-Na\"ively, Deep Sets form \eqref{eq:DeepSets} seems to imply that the function $f$ can be computed in linear time. 
-For problems like the computation of the gravitational potential for a system of point masses, all known solutions scale as the square of the number of points, and yet they are permutation-invariant functions and can thus be put in Deep Sets form.
-How can this be?
-There is no free lunch.
-The resolution must be either that $\phi$ will be very large in dimension---its dimensionality may even depend directly on $N$---or else that the complexity of $rho$ may grow rapidly with $N$.
+Because all these (computationally expensive) statistics are permutation-invariant, they must be possible to write in Deep Sets form.
+At the same time, Deep Sets form \eqref{eq:DeepSets} looks like it is computable in linear time. 
+What resolves this discrepancy?
+The resolution must be either that $\phi$ has a dimensionality that grows with $N$, or else that the complexity of $\rho$ grows with $N$.
 
-Still, Deep Sets form \eqref{eq:DeepSets} has a computational advantage in that it is explicitly in map--reduce form.
-Map--reduce is a model for parallelizing and distributed a calculation over a large dataset.
+Even if $\phi$ and $\rho$ grow rapidly with $N$, it might be that Deep Sets form \eqref{eq:DeepSets} delivers a computational advantage in that it is explicitly in map--reduce form.
+Map--reduce is a model for parallelizing and distributing a calculation over a large dataset.
 In our case, $\phi$ is the ``map'':
@@
 
-% HOGG TO CH: THIS PARAGRAPH REPEATS CONTENT FROM ABOVE. COMBINE HERE I THINK; THAT IS, MOVE IT DOWN TO HERE, BELOW DEEP SETS.
-% For one very specific example, all known exact formulae for the gravitational potential energy of a system of $N$ point masses involve sums over pairs of points (and also require taking differences of points); they have not ever been written in Deep Sets form \eqref{eq:DeepSets} to our knowledge.
-% CH: THAT SAID, THERE ARE MULTIPOLE-LIKE EXPRESSIONS THAT MIGHT BE IN DEEPSETS AND APPROXIMATE THIS CALCULATION; SEE THE DISCUSSION?
-% In cosmology, standard estimators for $k$-point functions (correlation functions, bispectra, and so on) involve sums over $k$-tuples; once again they have not been written in the Deep Sets form.
-% CH: DO THE PHILCOX PAPERS ADDRESS THIS PREVIOUS POINT?
-% Importantly, when these statistics are computed in their naive forms, they are very time-consuming because they involve sums over all elements of $N$-choose-$k$, which scales (badly) as $N^k$.
-% Because 3-point functions involve the shapes and areas of triangles, consideration of the 3-point function led to the question answered in this \documentname:
-% \emph{What is the Deep Sets (or map--reduce) expression for the area of a triangle?}
-% We deliver in \secref{sec:result}.
-
-% \section{A note on our method}\label{sec:method}
-% How did we find the formula \eqref{eq:result} presented below in \secref{sec:result}?
-% Fundamentally, our method took the form of a symbolic regression, followed by a confirmation with computational algebra.
-% CH THIS ISN'T TRUE ANY MORE?? OR IS IT?
-
-% For the symbolic regression, we considered a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
-% We fit a linear combination of the monomials to a data set of generated triples of points, drawn from a two-dimensional Gaussian whose mean was also drawn from a Gaussian in a two-dimensional space.
-% For each triple we attach a label, which is the squared area, and we require the linear combination of fourth-order, Deep Sets monomials to reproduce the labels to machine precision.
-% The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
-
-% We choose to work with squared area $A^2$ and not area $A$ because, geometrically, areas are signed quantities.
-% Indeed, the formulae \eqref{eq:school} through \eqref{eq:Heron} either (implicitly) involve square roots of polynomials, or else absolute value signs, such that the corresponding pure polynomial expression would be for the squared area.
-% We restricted to monomials that contain four powers of positional coordinates because any expression for the squared area must be units covariant (equivariant to changes of the units used for the measurements of positions or lengths).
-
-% Perhaps non-obviously, for a monomial to be in Deep Sets form, it must consist only of products of sums over the three points.
-% It cannot contain any differences, or any sums over pairs of points (or edges).
-% Thus $(\sum_{i=1}^2 x_i^2)\,(\sum_{i=1}^3 x_i\,y_i)$ is a permitted monomial, but $(\sum_{i=1}^3\sum_{j=1}^3 x_i^2\,y_j^2)$ is not a permitted monomial.
-% These monomials are both correct from a units perspective, but the second is not in Deep Sets form, because it contains products of position coordinates from different points. 
-% All the possible monomials are made of permutation-invariant factors of the form 
-% \begin{equation}
-% F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n.
-% \end{equation}
-% The invariant monomials of degree $d$ are then all products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
-
-% The symbolic regression produces a linear combination of monomials with real-valued coefficients.
-% We systematically search for an integer which, when multiplied by all of the coefficients, brings all coefficients very near integers.
-% We then round the coefficients to these nearest integers and check that we reproduce $({\rm denom}) A^2$ to better precision than the least-square result.
-% Trial-and-error delivered the integer 368, which makes all of the coefficients obviously rational.
-% We have no theory of this number; we leave it as a second exercise to the reader to explain this large, non-prime integer.
-
-% Finally, the expression found is output by the code in LaTeX form, to create the typesetting commands for \eqref{eq:result} and also output by the code in Mathematica form, to permit computational algebra operations.
-% A Mathematica \texttt{simplify} operation [HOGG NO NOT REALLY] is used to show that the Deep Sets expression \eqref{eq:result} is identical to the classic formulae for a triangle's squared area $A^2$.
-
 \pseudosection{Result}
 
-Our primary result is that area $\Delta$ of a triangle can be written in Deep Sets form as
-\begin{multline} \label{eq:result}
+These considerations led us to search for Deep Sets forms for problems that depend on pairs, triples, or $k$-tuples of points in a set. 
+The simplest such problem is the area of a triangle, which depends on pairs of points via
+side lengths, as in Heron's formula \eqref{eq:Heron};
+angles, as in \eqref{eq:sine};
+or point displacements, as in \eqref{eq:cross}.
+In this search, we find the following Deep Sets-form formula for the area $\Delta$ of a triangle:
+\begin{equation}
+\label{eq:result}
+\begin{multlined}
     \Delta^2 = \
@@
      ~,
-\end{multline}
-where the two-dimensional position of each of the three corners $i$ is $r_i \equiv (x_i, y_i)$.
-We verified this equation by using computational algebra to expand the formula and check its equivalence to \eqref{eq:cross}.
+\end{multlined}
+\end{equation}
+where the two-dimensional position of corner $i$ is $\vec{r}_i \equiv (x_i, y_i)$.
+This equation \eqref{eq:result} is the primary contribution of this \documentname.
 
-This formula is in Deep Sets form, since
+We verified equation \eqref{eq:result} via computational algebra, expanding and simplifying the formula to demonstrate its equivalence to \eqref{eq:cross}.
+\Hogg{Perhaps we should re-write cross to be in terms of differences again?
+Also, Figure~1 doesn't define the $\vec{r}_i$ variables, which in principle should be shown as displacements from an origin.}
+\Hogg{Try to make this text less narrative, and always stay in present tense! We are not telling a story of our lives.}
+The formula is in Deep Sets form, since
 every expression in parentheses in \eqref{eq:result} is a pure sum over all three points.
@@
 Being very explicit, it matches \eqref{eq:DeepSets} if we define $\rho$ and $\phi$ as follows:
-\begin{gather}
+\begin{equation}
+\begin{gathered}
     \phi(\vec{r}_i) = (
@@
     - \tfrac{1}{4} \, F_{20} \, F_{01}^2.
-\end{gather}
+\end{gathered}
+\end{equation}
 
@@
 
-Note that a triangle defined by three points has the one-half the area of the parallelogram spanned by the vectors $\vec{r}_2 - \vec{r}_1$ and $\vec{r}_3 - \vec{r}_1$.
-Multiplying both sides by 4 then reveals that \eqref{eq:result} is an integer polynomial of invariant factors for the area of the parallelogram.
+Note that a triangle defined by three points has one-half the area of the parallelogram spanned by the vectors $\vec{r}_2 - \vec{r}_1$ and $\vec{r}_3 - \vec{r}_1$.
+Multiplying both sides by 4 reveals that \eqref{eq:result} is an integer polynomial of invariant factors for the squared area of the parallelogram.
 Using this fact, we have written a program to perform a symbolic regression which recovers our result.
+\Hogg{Say here in a footnote maybe what the URL to the code is?}
 
-For the symbolic regression, we considered a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
+In this symbolic regression, we considered a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
 All the possible monomials are made of permutation-invariant factors of the form 
@@
 We require that the labels are reproduced to machine precision after rounding the coefficients to the nearest integers.
-% For each triple we attach a label, which is the squared area of the parallelogram spanned by the two difference vectors \CH{?}, and we require the linear combination of fourth-order, Deep Sets monomials to reproduce the labels to machine precision with integer coefficients.
 The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
@@
 If the triangle is embedded in a higher-dimensional space with the positions of the three vertices given by $\vec{r}_i \in \mathbb{R}^{d}$ for $d \geq 2$, the area of the triangle can be determined by summing in quadrature the areas of the triangles made by all projections to two axes.
-For example, in three dimensions, the area of the triangle is given by
+This is easiest to see starting from \eqref{eq:cross}, using the wedge product for vectors $\vec{a}, \vec{b}$ in a higher-dimensional space, as
 \begin{equation}
-    \Delta_{\rm 3D}^2 =
-    \Delta_{12}^2 + \Delta_{13}^2 + \Delta_{23}^2,
+\begin{aligned}
+    \Delta^2
+    &= \tfrac{1}{4}\, | \vec{a} \wedge \vec{b} |^2 \\
+    &= \tfrac{1}{4}\, | \vec{a} |^2 \, | \vec{b} |^2 - \tfrac{1}{4}\, ( \vec{a} \cdot \vec{b} )^2 \\
+    &= \tfrac{1}{4} \sum_{1 \leq j, k \leq d} ( a_j \, a_j \, b_k \, b_k - a_j \, b_j \, a_k \, b_k ) \\
+    &= \tfrac{1}{4} \sum_{1 \leq j < k \leq d} ( a_j \, b_k - a_k \, b_j)^2 \\
+    &= \sum_{1 \leq j < k \leq d} \Delta_{jk}^2,
+\end{aligned}
 \end{equation}
-and, in four dimensions, it is given by
-\begin{equation}
-    \Delta_{\rm 4D}^2 =
-    \Delta_{12}^2 + \Delta_{13}^2 + \Delta_{14}^2
-    + \Delta_{23}^2 + A_{24}^2
-    + \Delta_{34}^2,
-\end{equation}
-where we have introduced $\Delta_{jk}$, the area of the triangle formed by just the $j^{\rm th}$ and $k^{\rm th}$ components of the position vectors.
+where in the last line we have introduced $\Delta_{jk}$, the area of the triangle formed by just the $j^{\rm th}$ and $k^{\rm th}$ components of the position vectors.
+% For example, in three dimensions, the area of the triangle is given by
+% \begin{equation}
+%     \Delta_{\rm 3D}^2 =
+%     \Delta_{12}^2 + \Delta_{13}^2 + \Delta_{23}^2,
+% \end{equation}
+% and, in four dimensions, it is given by
+% \begin{equation}
+%     \Delta_{\rm 4D}^2 =
+%     \Delta_{12}^2 + \Delta_{13}^2 + \Delta_{14}^2
+%     + \Delta_{23}^2 + A_{24}^2
+%     + \Delta_{34}^2,
+% \end{equation}
 By using \eqref{eq:result} for each of the two-dimensional sub-areas, each can be computed in a permutation-invariant way and the entire result is thus permutation invariant and in Deep Sets form. 
-\Hogg{This last result somehow is related to the meaning of the square of the magnitude of a bivector.}
+% \Hogg{This last result somehow is related to the meaning of the square of the magnitude of a bivector.}
 
@@
 \end{align}
+There are many similarities between this and our result for the triangle area \eqref{eq:result}, which we believe may indicate a way to generalize this result to the hypervolume of a $d$-simplex.
 
@@
 
-We have stated above that Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, and yet it is not represented in Deep Sets form. Consider instead
+We stated above that Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, and yet it is not represented in Deep Sets form. Consider instead
 \begin{equation}
@@
         - 2 \left( a^4 + b^4 + c^4 \right)
-    }.
+    }~.
 \end{equation}
@@
     \phi(x) = (x^2, x^4), \quad
-    \vec{u} = {\textstyle \sum_{x \in \{ a, b, c \}}} \, \phi(x), \quad
-    \Delta = \rho(\vec{u}) = \frac{1}{4} \sqrt{ u_1^2 - 2 \, u_2 }.
+    {\textstyle \sum_{x \in \{ a, b, c \}}} \, \phi(x)
+    \equiv (F_2, F_4) , \quad
+    \Delta = \rho(\vec{u}) = \frac{1}{4} \sqrt{ F_2^2 - 2 \, F_4 }~.
 \end{equation}
@@
 
+\CH{add discussion on sines and cosines}
+
 \pseudosection{Discussion: open questions}
@@
 
-    \item Our formula for the volume of a simplex \eqref{eq:simplex} bears many similarities to that for the area of triangle \eqref{eq:result}.
+    \item Our triangle area \eqref{eq:result} and simplex volume \eqref{eq:simplex} formulae have many similarities.
 We take this to imply that a similar formula should exist for the hypervolume of a $d$-simplex.

### 2025-03-15 paper/main.tex
content, formatting tweaks, notes.

@@
 Oh and cosmology and power spectrum and bispectrum.
+
+% style notes
+% -----------
+% - Use \abs{} not |
+% - Put variable names next to the precise thing they label.
 
@@
 
-\renewcommand{\paragraph}[1]{\par\addvspace{1.5ex}\noindent\textbf{#1}~---}
-
 \usepackage{natbib,setspace,aas_macros}
@@
 \setlength{\bibsep}{0pt plus 0.3ex}
+
+\newcommand{\abs}[1]{|\,{#1}\,|}
+\renewcommand{\paragraph}[1]{\par\addvspace{1.5ex}\noindent\textsl{#1}~---}
 
@@
 \paragraph{Abstract}
-It is known that any permutation-invariant function of data points $r_i$
-can be written in the form $\rho(\sum_i\phi(r_i))$,
+It is known that any permutation-invariant function of data points $\vec{r}_i$
+can be written in the form $\rho(\sum_i\phi(\vec{r}_i))$,
 where $\rho$ and $\phi$ are functions.
@@
 &= \frac{1}{2}\,a\,b\,\sin(C) \label{eq:sine} \\
-&= \frac{1}{2}\, \left| \vec{a} \times \vec{b} \right|
-    = \frac{1}{2}\, \left| \vec{a} \wedge \vec{b} \right| \label{eq:cross} \\
+&= \frac{1}{2}\, \abs{\vec{a} \times \vec{b}}
+    = \frac{1}{2}\, \abs{\vec{a} \wedge \vec{b}}\label{eq:cross} \\
+&= \frac{1}{2}\, \abs{
+    x_1 \, y_2 - x_2 \, y_1 +
+    x_2 \, y_3 - x_3 \, y_2 +
+    x_3 \, y_1 - x_1 \, y_3
+}\label{eq:polynomial} \\
 &= \sqrt{s\,(s-a)\,(s-b)\,(s-c)} \label{eq:Heron} ~,
 \end{align}
-where the quantities used are defined in Figure~\ref{fig:triangle}, apart from the side lengths of the triangle
-$a = | \vec{a} |$, $b = | \vec{b} |$, $c = | \vec{c} |$
+where the quantities used are defined in Figure~\ref{fig:triangle}, apart from the side lengths
+$a = \abs{\vec{a}}$, $b = \abs{\vec{b}}$, $c = \abs{\vec{c}}$
 and the semi-perimeter $s = (a + b + c)/2$.
-\eqref{eq:school} is the usual introduction to the subject, traceable back even to Euclid.
+Formula~\eqref{eq:school} is the usual introduction to the subject, traceable back even to Euclid.
 % \eqref{eq:sine} is the side-angle-side formula.
-\eqref{eq:cross} generalizes to triangles in higher dimensional spaces, using the fact that the cross/wedge product has magnitude equal to the area of the parallelogram spanned by $\vec{a}$ and $\vec{b}$, which is twice the area of the triangle with $\vec{a}$ and $\vec{b}$ as sides.
-Lastly, \eqref{eq:Heron} is Heron's formula, which elegantly computes the area from the side lengths alone.
+Formula~\eqref{eq:cross} generalizes to triangles in higher dimensional spaces, using the fact that the cross or wedge product has a magnitude equal to the area of the parallelogram spanned by $\vec{a}$ and $\vec{b}$.
+Formula~\eqref{eq:polynomial} gives a polynomial in terms of the Cartesian coordinates of the corners in two dimensions; it is equal to an expansion of \eqref{eq:cross}.
+Formula~\eqref{eq:Heron} is Heron's formula, which elegantly computes the area from the side lengths alone.
 For a collection of 110 unique triangle area formulae, see \citet{baker1885a,baker1885b}.
@@
     \begin{tikzpicture}
+    \coordinate (F) at (-1.5,4.2);
     \coordinate (O) at (0,0);
@@
     \fill (B) circle (1.2pt) node[below right]{$(x_3, y_3)$};
+    \fill[gray] (F) circle (1.2pt) node[above left]{$(0, 0)$};
 
@@
 
+    \draw[->, line width=0.1mm, gray]
+        ($(F)!\pad!(O)$) -- ($(O)!2mm!(F)$)
+        node[near start, left, xshift=-1mm, yshift=0mm, black] {$\vec{r}_1$};
+    \draw[->, line width=0.1mm, gray]
+        ($(F)!\pad!(A)$) -- ($(A)!2mm!(F)$)
+        node[near start, above, xshift=+1mm, yshift=+1mm, black] {$\vec{r}_2$};
+    \draw[->, line width=0.1mm, gray]
+        ($(F)!\pad!(B)$) -- ($(B)!3mm!(F)$)
+        node[very near start, right, xshift=-2mm, yshift=-4mm, black] {$\vec{r}_3$};
+
     % draw arc for angle
@@
     % draw base, height guide lines
-    \draw ($(O)!\pad!(BaseO)$) -- ($(BaseO)$);
-    \draw ($(B)!\pad!(BaseB)$) -- ($(BaseB)$);
-    \draw ($(B)!\pad!(HeightO)$) -- ($(HeightO)$);
-    \draw ($(A)!\pad!(HeightA)$) -- ($(HeightA)$);
+    \draw[dotted, line width=0.3mm]($(O)!\pad!(BaseO)$) -- ($(BaseO)$);
+    \draw[dotted, line width=0.3mm]($(B)!\pad!(BaseB)$) -- ($(BaseB)$);
+    \draw[dotted, line width=0.3mm]($(B)!\pad!(HeightO)$) -- ($(HeightO)$);
+    \draw[dotted, line width=0.3mm]($(A)!\pad!(HeightA)$) -- ($(HeightA)$);
 
@@
 However, these symmetries are not always obvious from a given formula; for example, of those listed above, only Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant.
+\Hogg{Isn't formula \eqref{eq:polynomial} also explicitly permutation invariant?}
 
@@
 
-A function $f(r_1, \dots, r_N)$ is invariant to permutation of its arguments if and only if it can be decomposed in the form
+A function $f(\vec{r}_1, \dots, \vec{r}_N)$ is invariant to permutation of its arguments if and only if it can be decomposed in the form
 \begin{equation}
-    f(r_1, \dots, r_N) = \rho \left( {\textstyle \sum_{i=1}^{N}} \, \phi(r_i) \right), \label{eq:DeepSets}
+    f(r_1, \dots, r_N) = \rho \, \big( \sum_{i=1}^{N} \, \phi(\vec{r}_i) \big), \label{eq:DeepSets}
 \end{equation}
@@
 This result---often known by the name of the deep learning architecture derived from it, \emph{Deep Sets} (\citealt{zaheer+17deepsets}; see Theorem~2)---is not at all obvious.
-For example, even though Heron's formula \eqref{eq:Heron} is permutation-invariant, it is not actually in the form of \eqref{eq:DeepSets}. 
-Further, despite the triangle area being permutation-invariant, there is no existing formula for the area of triangle which is explicitly in this form.
+For example, even though Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, it is not actually in the form of \eqref{eq:DeepSets}. 
+Further, despite the triangle area being permutation-invariant, there is no existing \Hogg{previously known?} formula for the area of triangle which is explicitly in this form.
 
@@
 In this search, we find the following Deep Sets-form formula for the area $\Delta$ of a triangle:
-\begin{equation}
+\begin{align}
+    \Delta^2 = \ &
+    \frac{3}{4} \, \big( \sum_{i=1}^{3} x_i^2 \big) \, \big( \sum_{i=1}^{3} y_i^2 \big)
+    - \frac{3}{4} \, \big( \sum_{i=1}^{3} x_i \, y_i \big)^2 
+    + \frac{2}{4} \, \big( \sum_{i=1}^{3} x_i \, y_i \big) \, \big( \sum_{i=1}^{3} x_i \big) \, \big( \sum_{i=1}^{3} y_i \big)
+    \nonumber\\ &
+    - \, \frac{1}{4} \, \big( \sum_{i=1}^{3} y_i^2 \big) \, \big( \sum_{i=1}^{3} x_i \big)^2
+    - \, \frac{1}{4} \, \big( \sum_{i=1}^{3} x_i^2 \big) \, \big( \sum_{i=1}^{3} y_i \big)^2
+     ~,
 \label{eq:result}
-\begin{multlined}
-    \Delta^2 = \
-    \frac{3}{4} \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i^2 )
-    - \frac{3}{4} \, (\sum_{i=1}^{3} x_i \, y_i )^2 
-    + \frac{2}{4} \, (\sum_{i=1}^{3} x_i \, y_i ) \, (\sum_{i=1}^{3} x_i ) \, (\sum_{i=1}^{3} y_i )
-    \\
-    - \, \frac{1}{4} \, (\sum_{i=1}^{3} y_i^2 ) \, (\sum_{i=1}^{3} x_i )^2
-    - \, \frac{1}{4} \, (\sum_{i=1}^{3} x_i^2 ) \, (\sum_{i=1}^{3} y_i )^2
-     ~,
-\end{multlined}
-\end{equation}
+\end{align}
 where the two-dimensional position of corner $i$ is $\vec{r}_i \equiv (x_i, y_i)$.
@@
 
-We verified equation \eqref{eq:result} via computational algebra, expanding and simplifying the formula to demonstrate its equivalence to \eqref{eq:cross}.
-\Hogg{Perhaps we should re-write cross to be in terms of differences again?
+We verified equation \eqref{eq:result} via computational algebra, expanding and simplifying the formula to demonstrate its equivalence to \eqref{eq:polynomial}.
+% \Hogg{Perhaps we should re-write cross to be in terms of differences again?
 Also, Figure~1 doesn't define the $\vec{r}_i$ variables, which in principle should be shown as displacements from an origin.}
@@
 Being very explicit, it matches \eqref{eq:DeepSets} if we define $\rho$ and $\phi$ as follows:
-\begin{equation}
-\begin{gathered}
+\begin{gather}
     \phi(\vec{r}_i) = (
@@
         y_i^2
-    ), \\
-    {\textstyle \sum_i} \phi(\vec{r}_i) = (
+    ), \quad
+    (
         F_{10}, \,
@@
         F_{02}
-    ), \\
-    \rho \left( {\textstyle \sum_i} \phi(\vec{r}_i) \right)
+    ) = \sum_i \phi(\vec{r}_i),
+    \nonumber\\
+    \rho \, \big( \sum_i \phi(\vec{r}_i) \big)
     = \tfrac{3}{4} \, F_{20} \, F_{02}
@@
     - \tfrac{1}{4} \, F_{20} \, F_{01}^2.
-\end{gathered}
-\end{equation}
+\end{gather}
 
@@
 
-Note that a triangle defined by three points has one-half the area of the parallelogram spanned by the vectors $\vec{r}_2 - \vec{r}_1$ and $\vec{r}_3 - \vec{r}_1$.
+Note that a triangle defined by three points has one-half the area of the parallelogram spanned by the vectors $\vec{a} = \vec{r}_2 - \vec{r}_1$ and $\vec{b} = \vec{r}_3 - \vec{r}_1$.
 Multiplying both sides by 4 reveals that \eqref{eq:result} is an integer polynomial of invariant factors for the squared area of the parallelogram.
-Using this fact, we have written a program to perform a symbolic regression which recovers our result.
-\Hogg{Say here in a footnote maybe what the URL to the code is?}
+Using this fact, we have written a program to perform a symbolic regression which recovers our result.\footnote{
+    The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
+}
 
@@
 We require that the labels are reproduced to machine precision after rounding the coefficients to the nearest integers.
-The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
 
@@
 This is easiest to see starting from \eqref{eq:cross}, using the wedge product for vectors $\vec{a}, \vec{b}$ in a higher-dimensional space, as
-\begin{equation}
-\begin{aligned}
+\begin{align}
     \Delta^2
-    &= \tfrac{1}{4}\, | \vec{a} \wedge \vec{b} |^2 \\
-    &= \tfrac{1}{4}\, | \vec{a} |^2 \, | \vec{b} |^2 - \tfrac{1}{4}\, ( \vec{a} \cdot \vec{b} )^2 \\
-    &= \tfrac{1}{4} \sum_{1 \leq j, k \leq d} ( a_j \, a_j \, b_k \, b_k - a_j \, b_j \, a_k \, b_k ) \\
-    &= \tfrac{1}{4} \sum_{1 \leq j < k \leq d} ( a_j \, b_k - a_k \, b_j)^2 \\
+    &= \frac{1}{4}\, \abs{\vec{a} \wedge \vec{b}}^2 \nonumber\\
+    &= \frac{1}{4}\, \abs{\vec{a}}^2 \, \abs{\vec{b}}^2 - \frac{1}{4}\, \abs{\vec{a} \cdot \vec{b}}^2 \nonumber\\
+    &= \frac{1}{4} \sum_{1 \leq j, k \leq d} ( a_j \, a_j \, b_k \, b_k - a_j \, b_j \, a_k \, b_k ) \nonumber\\
+    &= \frac{1}{4} \sum_{1 \leq j < k \leq d} ( a_j \, b_k - a_k \, b_j)^2 \nonumber\\
     &= \sum_{1 \leq j < k \leq d} \Delta_{jk}^2,
-\end{aligned}
-\end{equation}
+\end{align}
 where in the last line we have introduced $\Delta_{jk}$, the area of the triangle formed by just the $j^{\rm th}$ and $k^{\rm th}$ components of the position vectors.
@@
 The invariant factors then have three indices, $F_{\ell m n} = \sum_i x_i^\ell \, y_i^m \, z_i^n$, and we form all invariant monomials of degree six.
-The result, which we have verified by computational algebra, is that the volume of the simplex $V$ is given by
+The result, which we have verified by computational algebra, is that the volume of the simplex $V$ is \Hogg{Since it is the volume that is $V$ not the simplex, it should say ``The volume $V$ of the simplex...''} given by
 \begin{align}
@@
 \end{equation}
-It is simple to show that this is exactly equivalent to Heron's formula when expanded, but it is now in Deep Sets form. Explicitly,
-\begin{equation}
-    \phi(x) = (x^2, x^4), \quad
-    {\textstyle \sum_{x \in \{ a, b, c \}}} \, \phi(x)
-    \equiv (F_2, F_4) , \quad
-    \Delta = \rho(\vec{u}) = \frac{1}{4} \sqrt{ F_2^2 - 2 \, F_4 }~.
-\end{equation}
-
-This is not novel. Heron's formula is a special case of Brahmagupta's formula, which gives the area $K$ of a convex cyclic quadrilateral to be \citep{coolidge1939}
-\begin{equation}
-    K = \sqrt{
-        (s - a) \,
-        (s - b) \,
-        (s - c) \,
-        (s - d)
-    }~,
-\end{equation}
-where $a, b, c, d$ are the four side lengths and $s = (a+b+c+d)/2$ is the semi-perimeter. This formula has been alternatively written as \CH{seems like someone just put this on Wikipedia?}
+It is simple to show that this is exactly equivalent to Heron's formula when expanded, but it is now in Deep Sets form.\footnote{
+    This is not novel.
+Heron's formula is a special case of Brahmagupta's formula, which gives the area $K$ of a convex cyclic quadrilateral in terms of its four side lengths $a, b, c, d$ \citep{coolidge1939}.
+    On Wikipedia, it has been written
     \begin{equation}
@@
     \end{equation}
-which is exactly \eqref{eq:HeronDeepSets} if the fourth side-length $d$ is zero.
-Note that this expression is in Deep Sets form if the product $a \, b \, c \, d$ is taken to be the exponential of the sum of logarithms.
+    which is exactly \eqref{eq:HeronDeepSets} if $d$ is zero.
+    This expression is in Deep Sets form, with the product $a \, b \, c \, d$ 
+    representable as $\exp( \sum_{x \in \{ a,b,c,d \}} \log x )$.
+    % taken to be the exponential of the sum of logarithms.
+}
 
@@
 The area of a triangle is not only invariant to permutation, but also to translation, rotation, and reflection.
-Formulae \eqref{eq:sine} and \eqref{eq:Heron}, by virtue of referencing only side-lengths and angles, are obviously invariant to these other symmetries.
+Formulae \eqref{eq:sine} and \eqref{eq:Heron}, by virtue of referencing only side-lengths and angles, are invariant to these other symmetries.
 While true, it is not obvious that our solution \eqref{eq:result} is invariant. An explicit proof of this is left to the reader.

### 2025-03-22 paper/bbs_mail.tex
various, including authorship

\documentclass{article}
\usepackage{amsfonts}
\usepackage[parfill]{parskip}

\begin{document}

...

2) The existence of the formula follows from general theory that long predates the DeepSets paper. This seems worth mentioning.

2a) It's known classically (by the late 18th century) that permutation-invariant polynomials on a set of 1-dimensional data are representable in terms of (in fact, are polynomials in) the power sums $\sum_i x_i^j$ for $j=1,\dots,n$ (where $n$ is the size of the data). The first written proof on the record is probably the one in Eduard Waring’s \textit{Meditationes Algebraicae} (1782).

2b) Weyl proved in his \textit{The Classical Groups} (the edition everybody cites is 1946 but I think there was an earlier edition?) that the permutation-invariant polynomials on a set of any-dimensional data are representable in terms of (in fact, polynomials in) the polarizations of the permutation-invariant polynomials on a corresponding set of 1-dimensional data.

Your $F_{mn}$'s are the 2-dimensional polarizations of the 1-dimensional power sums. So the existence of your formula follows from (2a) and (2b).

...

[D]efinition: if $f$ is a polynomial function of a column vector $v\in \mathbb{R}^n$, then the \textit{$d$-dimensional polarizations} of $f$ are functions on $d$ vectors $v_1,\dots,v_d\in \mathbb{R}^{n\times d}$ of the coefficients of each monomial in $a_1,\dots,a_d$ in the expansion of $f(a_1v_1 + \dots + a_dv_d)$.
In your setting, $v_1$ is $(x_1,x_2,x_3)^T$ and $v_2$ is $(y_1,y_2,y_3)^T$, $F_{10}$ and $F_{01}$ are the polarizations of the first power sum, and $F_{20}$, $F_{11}$, and $F_{02}$ are the polarizations of the second power sum.

\end{document}

### 2025-03-22 paper/main.tex
various, including authorship

@@
 % - Put variable names next to the precise thing they label.
+% - Put ~, or ~. at the end of equations that need punctuation at the end.
+% - Be careful about "not translationally invariant" -- we mean "not explicitly translationally invariant" and so on. Explicitness is what this paper is about.
 
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
 \usepackage[letterpaper]{geometry}
 \usepackage[hidelinks]{hyperref}
-\usepackage{amsmath,amsfonts,mathtools,tikz}
-\usetikzlibrary{angles,calc,quotes}
-
 \usepackage[utf8]{inputenc}
 \usepackage[T1]{fontenc}
-\usepackage{lmodern}
+\usepackage{amsmath,amsfonts,lmodern,mathtools,nameref,setspace,tikz,titlesec}
+\usetikzlibrary{angles,calc,quotes}
 
@@
 \newcommand{\secref}[1]{\sectionname~\ref{#1}}
-\setlength{\textwidth}{5.50in}
+\setlength{\textwidth}{4.75in}
 \setlength{\oddsidemargin}{3.25in}\addtolength{\oddsidemargin}{-0.5\textwidth}
@@
 \sloppy\sloppypar\raggedbottom\frenchspacing
+\setstretch{1.08}
 \pagestyle{myheadings}
-\markboth{foo}{\sffamily Hainje, Hogg, \& Villar / A new formula for the area of a triangle}
+\markboth{foo}{\scshape hainje \& hogg / A useless formula for the area of a triangle}
 
-\usepackage{natbib,setspace,aas_macros}
+\titleformat*{\subsection}{\normalfont\normalsize\slshape}
+\titlespacing*{\subsection}{0pt}{0pt}{0pt}
+
+\usepackage{natbib,aas_macros}
 \bibliographystyle{aasjournal}
@@
 \newcommand{\Hogg}[1]{{\color{violet} (Hogg says: #1)}}
-\newcommand{\pseudosection}[1]
-{}
-% {\noindent{\color{green!66!black} (\textit{#1})}\par}
 
-% define the affiliation symbols
+% affiliation symbols
 \newcommand{\aOne}{\textsuperscript{\textasteriskcentered}}
@@
 
+\renewcommand{\d}{\mathrm{d}}
+
+\newcommand{\secbreak}{\bigskip{\centering\footnotesize $\triangle~~~\triangle~~~\triangle$\par}\bigskip\noindent}
+
 \begin{document}\thispagestyle{empty}
+\setcounter{secnumdepth}{0}
 
-\section*{\raggedright
-A new formula for the area of a triangle:
-Useless, but explicitly in Deep Sets form}
+\section*{\centering\normalsize\uppercase{
+A formula for the area of a triangle:\\
+Useless, but explicitly in Deep Sets form}}
 
+% \bigskip
 \noindent
-\textbf{Connor Hainje}\aOne\aSix,
-\textbf{David W. Hogg}\aOne\aTwo\aThree, and
-\textbf{Soledad Villar}\aFour\aFive
+\textbf{Connor Hainje}\aOne\aSix{}
+\&
+\textbf{David W. Hogg}\aOne\aTwo\aThree%, and
+%\textbf{Soledad Villar}\aFour\aFive
 
-\vspace{0.5em}
+\medskip
 {\footnotesize\par\noindent \aOne\textsl{
@@
 Center for Computational Astrophysics, Flatiron Institute}}
-{\footnotesize\par\noindent \aFour\textsl{
-Department of Applied Mathematics and Statistics, Johns Hopkins University}}
-{\footnotesize\par\noindent \aFive\textsl{
-Mathematical Institute for Data Science, Johns Hopkins University}}
+%{\footnotesize\par\noindent \aFour\textsl{
+%Department of Applied Mathematics and Statistics, Johns Hopkins University}}
+%{\footnotesize\par\noindent \aFive\textsl{
+%Mathematical Institute for Data Science, Johns Hopkins University}}
 {\footnotesize\par\noindent \aSix\textsl{
@@
 
-\bigskip
+\medskip
 \paragraph{Abstract}
@@
 where $\rho$ and $\phi$ are functions.
-%, and the function $\phi$ might output a high-dimensional latent encoding of each point $r_i$.
-This form---known in the machine-learning literature as Deep Sets---generates a map--reduce algorithm, because the $\phi$ executions can be performed asynchronously and the sums can be performed in any order.
+This form---known in the machine-learning literature as Deep Sets---also directly generates a map--reduce algorithm.
 The area of a triangle is a permutation-invariant function of the locations $\vec{r}_i$ of the three corners $1\leq i\leq 3$.
-We find, possibly for the first time, the formula for the area of a triangle that is explicitly in Deep Sets form.
+We find the formula for the area of a triangle that is explicitly in Deep Sets form.
 This project was motivated by questions about the fundamental computational complexity of $n$-point statistics in cosmology; that said, no insights of any kind were gained from these results.
 
-\bigskip
-{\centering $\ast~\ast~\ast$ \par}
-\bigskip
-
-\pseudosection{Areas}
-\noindent
+\secbreak
 The area $\Delta$ of a triangle can be written in many ways.
@@
 \Delta 
-&= \frac{1}{2} \, \text{(base)} \, \text{(height)} \label{eq:school} \\
+&= \frac{1}{2} \, (\text{base}) \, (\text{height}) \label{eq:school} \\
 &= \frac{1}{2}\,a\,b\,\sin(C) \label{eq:sine} \\
@@
 and the semi-perimeter $s = (a + b + c)/2$.
-Formula~\eqref{eq:school} is the usual introduction to the subject, traceable back even to Euclid.
-% \eqref{eq:sine} is the side-angle-side formula.
+Formula~\eqref{eq:school} is the usual introduction to the subject, traceable back to \citet{Euclid300BC}.
+Formula~\eqref{eq:sine} makes use of the fact that the height can be found from $\sin(C) = (\text{height}) / a$, commonly seen in introductory geometry courses.
 Formula~\eqref{eq:cross} generalizes to triangles in higher dimensional spaces, using the fact that the cross or wedge product has a magnitude equal to the area of the parallelogram spanned by $\vec{a}$ and $\vec{b}$.
@@
 Formula~\eqref{eq:Heron} is Heron's formula, which elegantly computes the area from the side lengths alone.
-For a collection of 110 unique triangle area formulae, see \citet{baker1885a,baker1885b}.
+For a collection of 110 unique triangle area formulae, see \citet{Baker1885a,Baker1885b}.
 
@@
     \draw[->, line width=0.1mm, gray]
-        ($(F)!\pad!(B)$) -- ($(B)!3mm!(F)$)
+        ($(F)!\pad!(B)$) -- ($(B)!3.5mm!(F)$)
         node[very near start, right, xshift=-2mm, yshift=-4mm, black] {$\vec{r}_3$};
@@
     % draw base, height guide lines
-    \draw[dotted, line width=0.3mm]($(O)!\pad!(BaseO)$) -- ($(BaseO)$);
-    \draw[dotted, line width=0.3mm]($(B)!\pad!(BaseB)$) -- ($(BaseB)$);
-    \draw[dotted, line width=0.3mm]($(B)!\pad!(HeightO)$) -- ($(HeightO)$);
-    \draw[dotted, line width=0.3mm]($(A)!\pad!(HeightA)$) -- ($(HeightA)$);
+    \draw[dotted, line width=0.2mm]($(O)!\pad!(BaseO)$) -- ($(BaseO)$);
+    \draw[dotted, line width=0.2mm]($(B)!\pad!(BaseB)$) -- ($(BaseB)$);
+    \draw[dotted, line width=0.2mm]($(B)!\pad!(HeightO)$) -- ($(HeightO)$);
+    \draw[dotted, line width=0.2mm]($(A)!\pad!(HeightA)$) -- ($(HeightA)$);
 
@@
 It is invariant to translation, rotation, and reflection, as well as to the labeling of sides or corners.
-However, these symmetries are not always obvious from a given formula; for example, of those listed above, only Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant.
-\Hogg{Isn't formula \eqref{eq:polynomial} also explicitly permutation invariant?}
-
-% \CH{struggling to find a segue here} \Hogg{We could call out symmetries: These formulae are for a quantity that obeys permutation, translation, rotation, and reflection symmetries.
-But these symmetries aren't always visible.
-Then go to the point (above) about Heron's formula.}
-
-\pseudosection{Deep Sets}
+However, these symmetries are not always obvious from a given formula; for example, of those listed above, only the polynomial \eqref{eq:polynomial} and Heron's formula \eqref{eq:Heron} are explicitly permutation-invariant.
 
@@
 \begin{equation}
-    f(r_1, \dots, r_N) = \rho \, \big( \sum_{i=1}^{N} \, \phi(\vec{r}_i) \big), \label{eq:DeepSets}
+    f(r_1, \dots, r_N) = \rho \, \big( \sum_{i=1}^{N} \, \phi(\vec{r}_i) \big)~,\label{eq:DeepSets}
 \end{equation}
 for some functions $\rho$ and $\phi$.
-This result---often known by the name of the deep learning architecture derived from it, \emph{Deep Sets} (\citealt{zaheer+17deepsets}; see Theorem~2)---is not at all obvious.
+This result---often known by the name of the deep learning architecture derived from it, \emph{Deep Sets} (\citealt{Zaheer+17deepsets}; see Theorem~2)---is not at all obvious.
 For example, even though Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, it is not actually in the form of \eqref{eq:DeepSets}. 
-Further, despite the triangle area being permutation-invariant, there is no existing \Hogg{previously known?} formula for the area of triangle which is explicitly in this form.
+Further, despite the triangle area being permutation-invariant, there is no known (prior to this work) formula for the area of a triangle that is explicitly in Deep Sets form.
 
-\pseudosection{Computational motivations}
+While this result was recently popularized as Deep Sets, it follows from the classical theory of invariant polynomials.
+Specifically, it is known that the permutation-invariant polynomials on a set of any-dimensional data are representable in terms of the polarizations of the permutation-invariant polynomials on a corresponding set of one-dimensional data \citep{Weyl1939}.
+Further, the permutation-invariant polynomials on a set of one-dimensional data are representable in terms of the power sums $\sum_{i=1}^{n} r_i^j$ \citep{Waring1782}.
+This matches onto the Deep Sets form, as it states that a permutation-invariant polynomial of any-dimensional data $f(\{r_i\})$ can be decomposed into a polynomial ($\rho$) of polarizations of the power sums ($\sum_i \phi(\vec{r}_i)$).
+\CH{this is mostly plagiarized from BBS email; need to understand it better to massage it.} \Hogg{Important to emphasize that there is a proof of universality, and it's old.}
 
-% Although the contribution of this paper is minimal, or even zero, it was inspired by real problems in machine learning and physics. \CH{another bad segue}
+Although such a formula for the area of a triangle is of minimal, or even zero, practical interest, our search is motivated by real problems in physics.
+The laws of classical physics are generally invariant to permutation.
+For example, the gravitational (or electric) potential energy of a set of point masses (or charges) is invariant to the ordering or labeling of the points. 
+Specifically, for a system of $N$ point particles with masses $m_i$ at positions $\vec{r}_i$, the gravitational potential energy is
+\begin{equation}
+    U = -G \, \sum_{i=1}^{N} \sum_{j=i+1}^{N} \frac{m_i \, m_j}{\abs{\vec{r}_i - \vec{r}_j}} ~.
+\end{equation}
+This is explicitly permutation invariant since it involves a sum without repeats over all pairs of particles, but it is not in Deep Sets form \eqref{eq:DeepSets} because it involves a double sum (with $N^2$ computational scaling) of quantities computed from pairs of particles.
 
-Although such a formula is of minimal, or even zero, practical interest, our search is motivated by real problems in machine learning and physics.
-For example, in machine learning, one area of relevance is deep learning on point clouds (e.g. sets of three-dimensional position vectors) \citep{guo2020pointclouds}.
-Such methods are used for problems such as three-dimensional shape classification and object detection.
-
-\pseudosection{Cosmology motivations}
-
-The laws of classical physics are generally invariant to permutation.
-For example, the gravitational or electric potential energies of a set of point masses or point charges is invariant to the ordering or labeling of the points.
 Further, and important to our own motivations, cosmological clustering statistics such as 
-the correlation function \citep{},
-the power spectrum \citep{},
-the bispectrum \citep{},
-and higher-order statistics \citep{}
-used for precision measurement in cosmology \citep{}
+the correlation function \citep{Peebles1973},
+the power spectrum \citep{Peebles1973},
+the three-point correlation function \citep{PeeblesGroth1975},
+the bispectrum \citep{FrySeldner1982},
+and higher-order statistics \citep{Peebles1980book}
+used for precision measurement in cosmology \citep[e.g.,][]{Planck18CosmoParams,Planck18PNG,Planck18Inflation,Cabass+2022}
 are all permutation-invariant.
-All of these statistics na\"ively involve loops over all $k$-tuples of points, which not only is inconsistent with Deep Sets form, but which has a runtime that scales as $N^k$, where $N$ the number of points in the set.
-
-\pseudosection{Scalings}
+Calculating these statistics na\"ively involves loops over all $k$-tuplets of points, which is not only inconsistent with Deep Sets form, but also has a runtime that scales as $N^k$, where $N$ the number of points in the set.
 
@@
 The resolution must be either that $\phi$ has a dimensionality that grows with $N$, or else that the complexity of $\rho$ grows with $N$.
+Still, these time complexities may be better than the $N^k$ runtime of na\"ive implementations.
 
 Even if $\phi$ and $\rho$ grow rapidly with $N$, it might be that Deep Sets form \eqref{eq:DeepSets} delivers a computational advantage in that it is explicitly in map--reduce form.
-Map--reduce is a model for parallelizing and distributing a calculation over a large dataset.
+Map--reduce is a model for parallelizing and distributing a calculation over a large dataset \citep{DeanGhemawat2008,Lammel2008}.
 In our case, $\phi$ is the ``map'':
@@
 
-\pseudosection{Result}
-
-These considerations led us to search for Deep Sets forms for problems that depend on pairs, triples, or $k$-tuples of points in a set. 
+So, there may be benefits to finding Deep Sets forms for problems that depend on permutation-invariant pairs, triplets, or $k$-tuplets of points.
 The simplest such problem is the area of a triangle, which depends on pairs of points via
@@
 or point displacements, as in \eqref{eq:cross}.
-In this search, we find the following Deep Sets-form formula for the area $\Delta$ of a triangle:
+
+\secbreak
+We find this formula---in Deep Sets form---for the area $\Delta$ of a triangle:
 \begin{align}
@@
     - \, \frac{1}{4} \, \big( \sum_{i=1}^{3} x_i^2 \big) \, \big( \sum_{i=1}^{3} y_i \big)^2
-     ~,
+     ~.
 \label{eq:result}
 \end{align}
-where the two-dimensional position of corner $i$ is $\vec{r}_i \equiv (x_i, y_i)$.
-This equation \eqref{eq:result} is the primary contribution of this \documentname.
+This formula \eqref{eq:result} for the squared area\footnote{%
+    We choose to work with squared area $\Delta^2$ and not area $\Delta$ because, geometrically, areas are signed quantities.
+    Indeed, the formulae \eqref{eq:school} through \eqref{eq:Heron} either (implicitly) involve square roots of polynomials, or else absolute value signs, such that the corresponding pure polynomial expression would be for the squared area.}
+is the primary contribution of this \documentname, along with the generalization to higher-dimensional vectors $\vec{r}_i$ below in \eqref{eq:quadrature}.
 
-We verified equation \eqref{eq:result} via computational algebra, expanding and simplifying the formula to demonstrate its equivalence to \eqref{eq:polynomial}.
-% \Hogg{Perhaps we should re-write cross to be in terms of differences again?
-Also, Figure~1 doesn't define the $\vec{r}_i$ variables, which in principle should be shown as displacements from an origin.}
-\Hogg{Try to make this text less narrative, and always stay in present tense! We are not telling a story of our lives.}
+This equation \eqref{eq:result} is verified via computational algebra, expanding and simplifying the formula to demonstrate its equivalence to \eqref{eq:polynomial}.
 The formula is in Deep Sets form, since
 every expression in parentheses in \eqref{eq:result} is a pure sum over all three points.
-% There are no sums over differences, and there is no ordering of the points.
-Being very explicit, it matches \eqref{eq:DeepSets} if we define $\rho$ and $\phi$ as follows:
+Being very explicit, it matches \eqref{eq:DeepSets} for $\rho$ and $\phi$ defined as follows:
 \begin{gather}
@@
         F_{02}
-    ) = \sum_i \phi(\vec{r}_i),
-    \nonumber\\
-    \rho \, \big( \sum_i \phi(\vec{r}_i) \big)
-    = \tfrac{3}{4} \, F_{20} \, F_{02}
-    - \tfrac{3}{4} \, F_{11}^2
-    + \tfrac{2}{4} \, F_{11} \, F_{10} \, F_{01}
-    - \tfrac{1}{4} \, F_{02} \, F_{10}^2
-    - \tfrac{1}{4} \, F_{20} \, F_{01}^2.
+    ) = \sum_{i=1}^{3} \phi(\vec{r}_i),
+    \\
+    \rho \, \big( \sum_{i=1}^{3} \phi(\vec{r}_i) \big)
+    = \frac{3}{4} \, F_{20} \, F_{02}
+    - \frac{3}{4} \, F_{11}^2
+    + \frac{2}{4} \, F_{11} \, F_{10} \, F_{01}
+    - \frac{1}{4} \, F_{02} \, F_{10}^2
+    - \frac{1}{4} \, F_{20} \, F_{01}^2~.
+    \nonumber
 \end{gather}
 
-\pseudosection{A note on symbolic regression}
-
 Note that a triangle defined by three points has one-half the area of the parallelogram spanned by the vectors $\vec{a} = \vec{r}_2 - \vec{r}_1$ and $\vec{b} = \vec{r}_3 - \vec{r}_1$.
-Multiplying both sides by 4 reveals that \eqref{eq:result} is an integer polynomial of invariant factors for the squared area of the parallelogram.
-Using this fact, we have written a program to perform a symbolic regression which recovers our result.\footnote{
+Multiplying both sides by 4 reveals that \eqref{eq:result} is an integer polynomial of permutation-invariant factors for the squared area of the parallelogram.
+Using this fact, we have written code to perform a symbolic regression which recovers our result.\footnote{
     The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
@@
 
-In this symbolic regression, we considered a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
+In this symbolic regression, we consider a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
 All the possible monomials are made of permutation-invariant factors of the form 
 \begin{equation}
-F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n.
+F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n~.
 \end{equation}
@@
 The invariant monomials of degree $d$ are then products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
-We generated all factors $F_{mn}$ with degree $1 \leq m + n \leq 4$, then formed all products of these factors with total degree four.
+We generate all factors $F_{mn}$ with degree $1 \leq m + n \leq 4$, then form all products of these factors with total degree four.
 However, these monomials are not all linearly-independent;
-the five monomials made of a single degree-four factor ($F_{40}$, $F_{31}$, $F_{22}$, $F_{13}$, $F_{04}$) can be written as linear combinations of the others and are thus removed.
-This leaves \CH{number?} monomials.
+the five monomials made of a single degree-four factor ($F_{40}$, $F_{31}$, $F_{22}$, $F_{13}$, $F_{04}$) can be written as a linear combination of the other monomials and are thus removed.
+This leaves 28 linearly-independent monomials.
 
@@
 
-\pseudosection{Generalizations: higher dimensions}
-
-If the triangle is embedded in a higher-dimensional space with the positions of the three vertices given by $\vec{r}_i \in \mathbb{R}^{d}$ for $d \geq 2$, the area of the triangle can be determined by summing in quadrature the areas of the triangles made by all projections to two axes.
-This is easiest to see starting from \eqref{eq:cross}, using the wedge product for vectors $\vec{a}, \vec{b}$ in a higher-dimensional space, as
-\begin{align}
-    \Delta^2
-    &= \frac{1}{4}\, \abs{\vec{a} \wedge \vec{b}}^2 \nonumber\\
-    &= \frac{1}{4}\, \abs{\vec{a}}^2 \, \abs{\vec{b}}^2 - \frac{1}{4}\, \abs{\vec{a} \cdot \vec{b}}^2 \nonumber\\
-    &= \frac{1}{4} \sum_{1 \leq j, k \leq d} ( a_j \, a_j \, b_k \, b_k - a_j \, b_j \, a_k \, b_k ) \nonumber\\
-    &= \frac{1}{4} \sum_{1 \leq j < k \leq d} ( a_j \, b_k - a_k \, b_j)^2 \nonumber\\
-    &= \sum_{1 \leq j < k \leq d} \Delta_{jk}^2,
-\end{align}
-where in the last line we have introduced $\Delta_{jk}$, the area of the triangle formed by just the $j^{\rm th}$ and $k^{\rm th}$ components of the position vectors.
-% For example, in three dimensions, the area of the triangle is given by
-% \begin{equation}
-%     \Delta_{\rm 3D}^2 =
-%     \Delta_{12}^2 + \Delta_{13}^2 + \Delta_{23}^2,
-% \end{equation}
-% and, in four dimensions, it is given by
-% \begin{equation}
-%     \Delta_{\rm 4D}^2 =
-%     \Delta_{12}^2 + \Delta_{13}^2 + \Delta_{14}^2
-%     + \Delta_{23}^2 + A_{24}^2
-%     + \Delta_{34}^2,
-% \end{equation}
-By using \eqref{eq:result} for each of the two-dimensional sub-areas, each can be computed in a permutation-invariant way and the entire result is thus permutation invariant and in Deep Sets form. 
-% \Hogg{This last result somehow is related to the meaning of the square of the magnitude of a bivector.}
-
-\pseudosection{Generalizations: simplex volume}
+If the triangle is embedded in a higher-dimensional space with the positions of the three vertices given by $\vec{r}_i \in \mathbb{R}^{d}$ for $d \geq 2$, the area of the triangle can be determined by summing in quadrature the areas of the triangles made by all mutually orthogonal two-dimensional projections of the points:
+\begin{equation}
+    \Delta^2 = \sum_{1 \leq j < k \leq d} \Delta_{jk}^2~, 
+    \label{eq:quadrature}
+\end{equation}
+where $\Delta_{jk}$ is the area of the two-dimensional triangle formed by just the $j^{\rm th}$ and $k^{\rm th}$ components of the position vectors.
+A proof is given in the \nameref{sec:appendix}.
+By using \eqref{eq:result} for each of the two-dimensional sub-areas, each can be computed in a permutation-invariant way and the entire result is thus explicitly permutation invariant and in Deep Sets form.
 
@@
 Like the area of a triangle is one-half the area of a related parallelogram, the volume of a simplex is one-sixth the volume of the parallelepiped spanned by the vectors $\{ \vec{r}_2 - \vec{r}_1, \vec{r}_3 - \vec{r}_1, \vec{r}_4 - \vec{r}_1 \}$.
-
-Expecting an integer solution once again, we modified our symbolic regression to generate 4-tuples of points, using the squared volumes of the parallelepipeds they produce as labels. 
+Expecting an integer solution once again, we modify the symbolic regression to generate 4-tuplets of points, using the squared volumes of the parallelepipeds they produce as labels.
 The invariant factors then have three indices, $F_{\ell m n} = \sum_i x_i^\ell \, y_i^m \, z_i^n$, and we form all invariant monomials of degree six.
-The result, which we have verified by computational algebra, is that the volume of the simplex $V$ is \Hogg{Since it is the volume that is $V$ not the simplex, it should say ``The volume $V$ of the simplex...''} given by
+The result, verifiable by computational algebra, is that the volume $V$ of the simplex is given by
 \begin{align}
@@
 -1 \, F_{200} \, F_{002} \, F_{010}^2
--1 \, F_{020} \, F_{002} \, F_{100}^2 \, .
+-1 \, F_{020} \, F_{002} \, F_{100}^2~.
 \label{eq:simplex}
@@
 
-\pseudosection{Generalizations: Heron symmetrized}
-
+\secbreak
+We now consider two more applications of the Deep Sets form to Heron's formula and to the potential energy of a system of point masses.
 We stated above that Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, and yet it is not represented in Deep Sets form. Consider instead
@@
     representable as $\exp( \sum_{x \in \{ a,b,c,d \}} \log x )$.
-    % taken to be the exponential of the sum of logarithms.
 }
 
-\pseudosection{Discussion: other symmetries}
+\CH{feels out of place...}
+The area of a triangle is not only invariant to permutation, but also to translation, rotation, and reflection.
+Formulae \eqref{eq:sine} and \eqref{eq:Heron}, by virtue of referencing only side-lengths and angles, are explicitly invariant to these other symmetries.
+While true, it is not obvious that our solution \eqref{eq:result} is so invariant.
+An explicit proof of this is left to the reader.
 
-The area of a triangle is not only invariant to permutation, but also to translation, rotation, and reflection.
-Formulae \eqref{eq:sine} and \eqref{eq:Heron}, by virtue of referencing only side-lengths and angles, are invariant to these other symmetries.
-While true, it is not obvious that our solution \eqref{eq:result} is invariant. An explicit proof of this is left to the reader.
+While the triangle area is related to physics problems, the polynomial methods we employ to find the Deep Sets form for the triangle area \eqref{eq:result} are not relevant in a physical context.
+More relevant instead is to consider a $\phi$ which Fourier transforms the input data.
+In doing so, translation symmetry is encoded as well, since the Fourier basis $\exp (i\, \vec{k}\cdot\vec{x})$ is the eigenbasis of the translation operator.
+Below, we show that this can be used to produce a Deep Sets form for the gravitational potential energy of a system of point masses.
 
-\pseudosection{Discussion: sines and cosines}
+Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
+The total energy is given in Deep Sets form by
+\begin{align}
+    \phi_{\vec{k}}(m_j, \vec{r}_j)
+    &=m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}
+    ~,
+    \quad
+    \sum_{i=1}^{N} \phi_{\vec{k}}(m_i, \vec{r}_i)
+    = \mathcal{F}[\rho](\vec{k})~,
+    \nonumber\\
+    U \, \big( \sum_{i=1}^{N} \, \phi_{\vec{k}}(m_i, \vec{r}_i) \big)
+    &= -\frac{G}{4 \pi^2} \,
+        \int \d^3 r \,
+        \mathcal{F}^{-1} \left[
+            \mathcal{F}[\rho] \ast \left(
+            k^{-2} \, \mathcal{F}[\rho] \right)
+        \right]~.
+    \label{eq:EnergyDeepSets}
+\end{align}
+A derivation is given in the \nameref{sec:appendix}.
+In principle this requires $\phi$ to be evaluated at every possible wavenumber $\vec{k}$.
+To avoid making this infinite-dimensional $\phi$, one can instead evaluate $\mathcal{F}[\rho]$ at a finite set of wavenumbers $\vec{k}$ and live with approximations to the integrals necessary for the convolution and inverse Fourier transform.
+The approximation will improve as the size of $\phi$ grows.
+Approaches akin to this Fourier approach can be used \citep[and are used; e.g.,][]{Portillo+2018,Philcox+2022ENCORE} to estimate or approximate $k$-point statistics in cosmology, which was our motivating question.
 
-\CH{add discussion on sines and cosines}
+We leave the reader with the following questions. 
+It is trivial to take the form of \eqref{eq:result} and evaluate the sums for more than three vectors; does the result have any meaning?
+Is there a generalization of the triangle area \eqref{eq:result} and simplex volume \eqref{eq:simplex} results to a $d$-simplex?
+Our formulae \eqref{eq:result} and \eqref{eq:simplex} only used invariant factors of degree $2$, with exponents that are larger than $2$.
+Does this generalize to higher dimension simplexes, and can it be shown that higher order factors aren't needed?
+One can replace the sum in \eqref{eq:DeepSets} with any other permutation-invariant aggregator (e.g., max or product), though we do not know if there are any claims to universality in these cases.
+Does another area formula exist for these other aggregators?
+We have discussed the general value of a map function $\phi$ which Fourier transforms the points;\footnote{%
+\Hogg{The Deep Sets version of the Fourier transform is reminiscent of a positional encoding (CITE), similar to those used to preserve ordering information in (otherwise permutation invariant) transformers (CITE).}}
+does a formula like \eqref{eq:EnergyDeepSets} exist for the area of a triangle?
 
-\pseudosection{Discussion: open questions}
+% \Hogg{Maybe drop this and replace it with a negative conclusion about learning nothing.}
+% In this \documentname,
+% we presented a new Deep Sets formula for the area of a triangle \eqref{eq:result} given the coordinates of the triangle's vertices.
+% We extended it to triangles in higher-dimensional spaces \eqref{eq:quadrature}
+% and to compute the volume of a simplex \eqref{eq:simplex}.
+% We also considered other applications of the Deep Sets form, including to Heron's formula \eqref{eq:HeronDeepSets}
+% and the gravitational potential energy of a system of point masses \eqref{eq:EnergyDeepSets}.
 
-% HOGG SAY: CH: EXPAND THIS INTO AN OPEN QUESTIONS LIST:
-% \begin{itemize}
-% \item questionable: What happens if we try to do $A$ as a quadratic instead of $A^2$ as a quartic?
-% \item What happens if you hold the formula fixed, but use more than $N=3$ points?
-% The formula still returns a squared area.
-% That's the area of what?
-% \item What about areas of quadrangles or larger polygons?
-% \item What about volumes and hypervolumes of larger objects?
-% \end{itemize}
+{\footnotesize\par\bigskip\noindent
+This project would not have been possible without originating inspiration and discussions with Soledad Villar (JHU).
+The authors also thank Ben Blum-Smith (JHU), Jonathan Goodman (NYU), Wilson Gregory (JHU), Teresa Huang (Flatiron), Mike O'Neil (NYU), Alexander Novara (NYU), Ben Wandelt (JHU), Kaze Wong (JHU), and the Blanton--Hogg group meeting at NYU for valuable discussions.
+The authors also note that, while we believe the formula is (barely) novel, it is difficult to be sure, given the challenge of conducting a complete literature review for this ancient topic.\par}
 
-We leave the reader with the following open questions:
-\begin{itemize}
-    \item It is trivial to take the form of \eqref{eq:result} and evaluate the sums for more than three vectors. Does the result have any meaning?
+\secbreak
+% need a line break this time
 
-    \item Our triangle area \eqref{eq:result} and simplex volume \eqref{eq:simplex} formulae have many similarities.
-We take this to imply that a similar formula should exist for the hypervolume of a $d$-simplex.
-What is this result?
+\vspace{-\bigskipamount}
+\renewcommand{\section}[2]{}%
+{\small\singlespacing\bibliography{main}\par}
 
-    \item In both the triangle area and simplex volume formulae, the only invariant factors used are of degree $2$, and the factors never appear with exponents larger than $2$.
-Does this generalize?
-Can it be shown that higher order factors aren't needed?
-\end{itemize}
+\secbreak
+\appendix
+\subsection{Appendix}
+\label{sec:appendix}
 
-\CH{conclude}
+\paragraph{Proof that area sums in quadrature} We prove equation \eqref{eq:quadrature}, that the area of a triangle in $d$ dimensions is the sum in quadrature of the areas of the triangles made by all orthogonal projections to two axes.
 
-The authors thank Alexander Novara (NYU), Ben Wandelt (JHU), Kaze Wong (JHU), and the Blanton--Hogg group meeting at NYU for valuable discussions.
-The authors would also like to note that, while we believe that this result is novel, it is difficult to be sure, given the impossibility of conducting a complete literature review for such an ancient topic.
+From the three position vectors $\vec{r}_1, \vec{r}_2, \vec{r}_3 \in \mathbb{R}^d$ (for $d \geq 2$), we consider the two displacement vectors $\vec{a} \equiv \vec{r}_2 - \vec{r}_1$ and $\vec{b} \equiv \vec{r}_3 - \vec{r}_1$.
+Given these, we can write the two-dimensional triangle sub-area $\Delta_{jk}$ by first defining the projected two-dimensional displacement vectors $\vec{a}_{jk} = ( a_j, a_k )$, $\vec{b}_{jk} = ( b_j, b_k )$.
+Then,
+\begin{equation}
+    \Delta_{jk}
+    = \frac{1}{2} \, \abs{ \vec{a}_{jk} \wedge \vec{b}_{jk} }
+    = \frac{1}{2}
+        \, \abs{a_j \, b_k - b_j \, a_k}~,
+\end{equation}
+using formula \eqref{eq:cross}.
 
-\bigskip
-{\centering $\ast~\ast~\ast$ \par}
+Next, consider formula \eqref{eq:cross} with the full $d$-dimensional vectors:
+\begin{align}
+    \Delta^2
+    &= \frac{1}{4}\, \abs{\vec{a} \wedge \vec{b}}^2 \nonumber\\
+    &= \frac{1}{4}\, \abs{\vec{a}}^2 \, \abs{\vec{b}}^2 - \frac{1}{4}\, \abs{\vec{a} \cdot \vec{b}}^2 \nonumber\\
+    &= \frac{1}{4} \sum_{1 \leq j, k \leq d}
+        ( a_j \, a_j \, b_k \, b_k - a_j \, b_j \, a_k \, b_k ) \nonumber\\
+    &= \frac{1}{4} \sum_{1 \leq j < k \leq d}
+        ( a_j \, b_k - b_j \, a_k )^2 \nonumber\\
+    &= \sum_{1 \leq j < k \leq d} \Delta_{jk}^2~.
+\end{align}
 
-\renewcommand{\section}[2]{}%
-{\small\singlespacing\bibliography{main}}
+\paragraph{Derivation of Deep Sets energy} Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
+The energy density $\mathcal{E} = d U / d^3 x$ is given by
+\begin{equation}
+    \mathcal{E} = \frac{\rho \, \Phi,}{2}~,
+\end{equation}
+where $\rho$ is the mass density and $\Phi$ is the gravitational potential \citep{LandauLifshitzFields}.
+These are related by the Poisson equation
+\begin{equation}
+    \nabla^2 \Phi = 4 \pi G \, \rho~,
+\end{equation}
+which implies the following relationship between their Fourier transforms
+\begin{equation}
+    -k^2 \, \mathcal{F}[\Phi] = 4 \pi G \, \mathcal{F}[\rho]~,
+\end{equation}
+with $k^2 \equiv |\vec{k}|^2$.
+Using the convolution theorem, we then have
+\begin{align}
+    \mathcal{F}[\mathcal{E}] 
+    &= \frac{1}{2} \mathcal{F}[\rho \, \Phi] \nonumber\\
+    &= \frac{1}{2} \, \frac{\mathcal{F}[\rho] \ast \mathcal{F}[\Phi]}{(2\pi)^3} \nonumber\\
+    &= \frac{1}{2 \, (2\pi)^3} \, \mathcal{F}[\rho] \ast \left(
+        -4 \pi G \, k^{-2} \, \mathcal{F}[\rho]
+    \right)~.
+    \label{eq:ConvolutionThm}
+\end{align}
+This form of the convolution theorem is proven below.
+The Fourier transform of the mass density is
+\begin{equation}
+    \mathcal{F}[\rho](\vec{k})
+    = \int \d^3 r
+        \, e^{-i \, \vec{k} \cdot \vec{r}}
+        \, \sum_{j=1}^{N}
+        \, m_j
+        \, \delta(\vec{r} - \vec{r}_j)
+    = \sum_{j=1}^{N} \, m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}~,
+\end{equation}
+where we sum over $j$ and use $i=\sqrt{-1}$.
+Note that the value of the Fourier transform at a wavenumber $\vec{k}$ is of the required form $\sum_i \phi(m_i, \vec{r}_i)$ for Deep Sets features.
+Hence, the total energy is given in Deep Sets form by
+\begin{align}
+    \phi_{\vec{k}}(m_j, \vec{r}_j)
+    &=m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}
+    ~,
+    \quad
+    \sum_{i=1}^{N} \phi_{\vec{k}}(m_i, \vec{r}_i)
+    = \mathcal{F}[\rho](\vec{k})~,
+    \nonumber\\
+    U \, \big( \sum_{i=1}^{N} \, \phi_{\vec{k}}(m_i, \vec{r}_i) \big)
+    &= -\frac{G}{4 \pi^2} \,
+        \int \d^3 r \,
+        \mathcal{F}^{-1} \left[
+            \mathcal{F}[\rho] \ast \left(
+            k^{-2} \, \mathcal{F}[\rho] \right)
+        \right]~.
+\end{align}
+
+\paragraph{Proof of the convolution theorem} We prove the claim (equation \eqref{eq:ConvolutionThm}) that
+\begin{equation}
+    \mathcal{F}[\rho \, \Phi] 
+    = (2\pi)^{-3} \, \mathcal{F}[\rho] \ast \mathcal{F}[\Phi] ~.
+\end{equation}
+Note that we use the conventions:
+\begin{align}
+    \tilde{f}(\vec{k})
+    = \mathcal{F}[f(\vec{x})]
+    &= \int \d^3 x
+        \, e^{-i \, \vec{k} \cdot \vec{x}}
+        \, f(\vec{x})~, \\
+    f(\vec{x})
+    = \mathcal{F}^{-1}[\tilde{f}(\vec{k})]
+    &= \int \frac{\d^3 k}{(2\pi)^3}
+        \, e^{+i \, \vec{k} \cdot \vec{x}}
+        \, f(\vec{x})~.
+\end{align}
+Consider $\mathcal{F}^{-1}[\tilde{\rho} \ast \tilde{\Phi}](\vec{x})$. 
+\begin{align}
+    \mathcal{F}^{-1}[\tilde{\rho} \ast \tilde{\Phi}](\vec{x})
+    &= \int_{\mathbb{R}^3} \frac{\d^3 k}{(2\pi)^3}
+        \, e^{i \, \vec{k} \cdot \vec{x}} 
+        \, \big( \tilde{\rho} \ast \tilde{\Phi} \big) (\vec{k}) 
+        \nonumber\\
+    &= \int_{\mathbb{R}^3} \frac{\d^3 k}{(2\pi)^3}
+        \, e^{i \, \vec{k} \cdot \vec{x}} 
+        \, \int_{\mathbb{R}^3} \d^3 \vec{k}' 
+        \, \tilde{\rho}(\vec{k}') 
+        \, \tilde{\Phi}(\vec{k} - \vec{k}')
+        \nonumber\\
+    &= \int_{\mathbb{R}^3} \d^3 \vec{k}'
+        \, \tilde{\rho}(\vec{k}') 
+        \, \int_{\mathbb{R}^3} \frac{\d^3 k}{(2\pi)^3}
+        \, e^{i \, \vec{k} \cdot \vec{x}} 
+        \, \tilde{\Phi}(\vec{k} - \vec{k}')
+        \nonumber\\
+    &= \int_{\mathbb{R}^3} \d^3 \vec{k}'
+        \, \tilde{\rho}(\vec{k}') 
+        \, e^{i \, \vec{k}' \cdot \vec{x}}
+        \, \Phi(\vec{x})
+        \nonumber\\
+    &= (2\pi)^3
+        \, \rho(\vec{x}) 
+        \, \Phi(\vec{x})~.
+\end{align}
+Taking the Fourier transform of both sides gives the desired result.
 

### 2025-03-23 paper/main.tex
Fixed bibliography issues

@@
 \pagestyle{myheadings}
-\markboth{foo}{\scshape hainje \& hogg / A useless formula for the area of a triangle}
+\markboth{foo}{\scshape hainje {\footnotesize\&} hogg / useless formula for the area of a triangle}
+% \markboth{foo}{\scshape hainje et al. / useless formula for the area of a triangle}
 
@@
 
-\usepackage{natbib,aas_macros}
+\usepackage{natbib}
 \bibliographystyle{aasjournal}
@@
 
-% \bigskip
+\bigskip
 \noindent
@@
     \begin{tikzpicture}
-    \coordinate (F) at (-1.5,4.2);
+    \coordinate (F) at (-1.5,4.4);
     \coordinate (O) at (0,0);
@@
 
-Although such a formula for the area of a triangle is of minimal, or even zero, practical interest, our search is motivated by real problems in physics.
+Although a Deep Sets formula for the area of a triangle is of minimal, or even zero, practical interest, our search is motivated by real problems in physics.
 The laws of classical physics are generally invariant to permutation.
@@
 
-\CH{feels out of place...}
+\CH{feels out of place...} \Hogg{This can be deleted.}
 The area of a triangle is not only invariant to permutation, but also to translation, rotation, and reflection.
@@
 
-% \Hogg{Maybe drop this and replace it with a negative conclusion about learning nothing.}
-% In this \documentname,
-% we presented a new Deep Sets formula for the area of a triangle \eqref{eq:result} given the coordinates of the triangle's vertices.
-% We extended it to triangles in higher-dimensional spaces \eqref{eq:quadrature}
-% and to compute the volume of a simplex \eqref{eq:simplex}.
-% We also considered other applications of the Deep Sets form, including to Heron's formula \eqref{eq:HeronDeepSets}
-% and the gravitational potential energy of a system of point masses \eqref{eq:EnergyDeepSets}.
-
 {\footnotesize\par\bigskip\noindent
 This project would not have been possible without originating inspiration and discussions with Soledad Villar (JHU).
-The authors also thank Ben Blum-Smith (JHU), Jonathan Goodman (NYU), Wilson Gregory (JHU), Teresa Huang (Flatiron), Mike O'Neil (NYU), Alexander Novara (NYU), Ben Wandelt (JHU), Kaze Wong (JHU), and the Blanton--Hogg group meeting at NYU for valuable discussions.
+The authors thank Ben Blum-Smith (JHU), Jonathan Goodman (NYU), Wilson Gregory (JHU), Teresa Huang (Flatiron), Mike O'Neil (NYU), Alexander Novara (NYU), Ben Wandelt (JHU), Kaze Wong (JHU), and the Blanton--Hogg group meeting at NYU for valuable discussions.
 The authors also note that, while we believe the formula is (barely) novel, it is difficult to be sure, given the challenge of conducting a complete literature review for this ancient topic.\par}

### 2025-03-23 paper/main.tex
Notes, comments, tiny bibliographic formatting adjustments.

@@
 
-% typesetting issues
+% page typesetting issues
 \newcommand{\documentname}{\textsl{Note}}
-\newcommand{\sectionname}{Section}
-\newcommand{\secref}[1]{\sectionname~\ref{#1}}
 \setlength{\textwidth}{4.75in}
@@
 \setstretch{1.08}
+
+% document formatting and sectioning issues
 \pagestyle{myheadings}
+\renewcommand{\paragraph}[1]{\par\addvspace{1.5ex}\noindent\textsl{#1}~---}
+\newcommand{\secbreak}{\bigskip{\centering\footnotesize $\triangle~~~\triangle~~~\triangle$\par}\bigskip\noindent}
+\titleformat*{\subsection}{\normalfont\normalsize\slshape}
+\titlespacing*{\subsection}{0pt}{0pt}{0pt}
 \markboth{foo}{\scshape hainje {\footnotesize\&} hogg / useless formula for the area of a triangle}
@@
 
-\titleformat*{\subsection}{\normalfont\normalsize\slshape}
-\titlespacing*{\subsection}{0pt}{0pt}{0pt}
-
+% bibliography issues
 \usepackage{natbib}
@@
 
+% math issues
+\renewcommand{\d}{\mathrm{d}}
 \newcommand{\abs}[1]{|\,{#1}\,|}
-\renewcommand{\paragraph}[1]{\par\addvspace{1.5ex}\noindent\textsl{#1}~---}
 
+% issues
 \usepackage{xcolor}
@@
 
-\renewcommand{\d}{\mathrm{d}}
-
-\newcommand{\secbreak}{\bigskip{\centering\footnotesize $\triangle~~~\triangle~~~\triangle$\par}\bigskip\noindent}
-
-\begin{document}\thispagestyle{empty}
-\setcounter{secnumdepth}{0}
+\begin{document}\thispagestyle{empty}\setcounter{secnumdepth}{0}
 
@@
 
-\bigskip
+\medskip
 \noindent
 \textbf{Connor Hainje}\aOne\aSix{}
-\&
+\textsl{and}
 \textbf{David W. Hogg}\aOne\aTwo\aThree%, and
@@
 Multiplying both sides by 4 reveals that \eqref{eq:result} is an integer polynomial of permutation-invariant factors for the squared area of the parallelogram.
-Using this fact, we have written code to perform a symbolic regression which recovers our result.\footnote{
+Using this fact, we have written code to perform a symbolic regression that recovers our result.\footnote{
     The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
@@
 \begin{equation}
-F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n~.
+F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n~;
 \end{equation}
-Such a factor has degree $m + n$.
+each such a factor has degree $m + n$.
 The invariant monomials of degree $d$ are then products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
@@
 \end{equation}
-with $k^2 \equiv |\vec{k}|^2$.
+with $k^2 \equiv \abs{\vec{k}}^2$.
 Using the convolution theorem, we then have

### 2025-03-24 paper/main.tex
translator in bibtex
tiny things

@@
 \setlength{\textwidth}{4.75in}
+\setlength{\textheight}{9.20in}
 \setlength{\oddsidemargin}{3.25in}\addtolength{\oddsidemargin}{-0.5\textwidth}
-\addtolength{\topmargin}{-0.20in}
-\addtolength{\textheight}{1.40in}
-\sloppy\sloppypar\raggedbottom\frenchspacing
+\addtolength{\topmargin}{-0.30in}
+\sloppy\sloppypar\frenchspacing
 \setstretch{1.08}
@@
 \paragraph{Abstract}
-It is known that any permutation-invariant function of data points $\vec{r}_i$
+Any permutation-invariant function of data points $\vec{r}_i$
 can be written in the form $\rho(\sum_i\phi(\vec{r}_i))$,
 where $\rho$ and $\phi$ are functions.
-This form---known in the machine-learning literature as Deep Sets---also directly generates a map--reduce algorithm.
+This form---known in the machine-learning literature as Deep Sets---also generates a map--reduce algorithm.
 The area of a triangle is a permutation-invariant function of the locations $\vec{r}_i$ of the three corners $1\leq i\leq 3$.
@@
 where the quantities used are defined in Figure~\ref{fig:triangle}, apart from the side lengths
-$a = \abs{\vec{a}}$, $b = \abs{\vec{b}}$, $c = \abs{\vec{c}}$
+$a, b, c = \abs{\vec{a}}, \abs{\vec{b}}, \abs{\vec{c}}$
 and the semi-perimeter $s = (a + b + c)/2$.
@@
 The authors thank Ben Blum-Smith (JHU), Jonathan Goodman (NYU), Wilson Gregory (JHU), Teresa Huang (Flatiron), Mike O'Neil (NYU), Alexander Novara (NYU), Ben Wandelt (JHU), Kaze Wong (JHU), and the Blanton--Hogg group meeting at NYU for valuable discussions.
+The Flatiron Institute is a division of the Simons Foundation.
+\Hogg{CH: You need to acknowledge a grant number here, I believe.}
 The authors also note that, while we believe the formula is (barely) novel, it is difficult to be sure, given the challenge of conducting a complete literature review for this ancient topic.\par}

### 2025-03-25 paper/main.tex
CH content; fourier triangle area

@@
 % bibliography issues
-\usepackage{natbib}
+\usepackage{aas_macros,natbib}
 \bibliographystyle{aasjournal}
@@
 Any permutation-invariant function of data points $\vec{r}_i$
-can be written in the form $\rho(\sum_i\phi(\vec{r}_i))$,
-where $\rho$ and $\phi$ are functions.
+can be written in the form $\rho(\sum_i\phi(\vec{r}_i))$ for suitable functions $\rho$ and $\phi$.
 This form---known in the machine-learning literature as Deep Sets---also generates a map--reduce algorithm.
@@
 \begin{equation}
-    f(r_1, \dots, r_N) = \rho \, \big( \sum_{i=1}^{N} \, \phi(\vec{r}_i) \big)~,\label{eq:DeepSets}
+    f(\vec{r}_1, \dots, \vec{r}_N) = \rho \, \big( \sum_{i=1}^{N} \, \phi(\vec{r}_i) \big)~,\label{eq:DeepSets}
 \end{equation}
-for some functions $\rho$ and $\phi$.
+for some functions $\rho$ and $\phi$, where $\phi$ is usually vector valued.
 This result---often known by the name of the deep learning architecture derived from it, \emph{Deep Sets} (\citealt{Zaheer+17deepsets}; see Theorem~2)---is not at all obvious.
@@
 While this result was recently popularized as Deep Sets, it follows from the classical theory of invariant polynomials.
-Specifically, it is known that the permutation-invariant polynomials on a set of any-dimensional data are representable in terms of the polarizations of the permutation-invariant polynomials on a corresponding set of one-dimensional data \citep{Weyl1939}.
-Further, the permutation-invariant polynomials on a set of one-dimensional data are representable in terms of the power sums $\sum_{i=1}^{n} r_i^j$ \citep{Waring1782}.
-This matches onto the Deep Sets form, as it states that a permutation-invariant polynomial of any-dimensional data $f(\{r_i\})$ can be decomposed into a polynomial ($\rho$) of polarizations of the power sums ($\sum_i \phi(\vec{r}_i)$).
-\CH{this is mostly plagiarized from BBS email; need to understand it better to massage it.} \Hogg{Important to emphasize that there is a proof of universality, and it's old.}
+It has long been known that any permutation-invariant (or symmetric) polynomial of scalar data $f(\{r_i\})$ can be written as a polynomial of the power sums $\sum_i r_i^m$ \citep{Waring1782}.
+Further, symmetric polynomials of any-dimensional data are decomposable into \emph{polarizations} of the scalar symmetric polynomials \citep{Weyl1939}.
+Together, these imply that a permutation-invariant polynomial of vector data is a polynomial in the polarizations of the power sums; e.g., the power sums of the components of the vectors.
+This result is \emph{universal}, in the sense that it is true for all symmetric polynomials.
+It also matches onto the Deep Sets form, where the polarizations of the power sums are the components of the vector $\sum_i \phi(\vec{r}_i)$, which are combined in a polynomial $\rho$.
+
+% Specifically, it is known that the permutation-invariant polynomials on a set of any-dimensional data are representable in terms of the polarizations of the permutation-invariant polynomials on a corresponding set of one-dimensional data \citep{Weyl1939}.
+% Further, the permutation-invariant polynomials on a set of one-dimensional data are representable in terms of the power sums $\sum_{i=1}^{n} r_i^j$ \citep{Waring1782}.
+% Note that this decomposition is \emph{universal}, in the sense that it is true for \emph{all} symmetric polynomials of any-dimensional data. 
+% This also matches onto the Deep Sets form, as it states that a permutation-invariant polynomial of any-dimensional data $f(\{r_i\})$ can be decomposed into a polynomial ($\rho$) of polarizations of the power sums ($\sum_i \phi(\vec{r}_i)$).
+% \CH{this is mostly plagiarized from BBS email; need to understand it better to massage it.} 
+% \Hogg{Important to emphasize that there is a proof of universality, and it's old.}
 
@@
 and higher-order statistics \citep{Peebles1980book}
-used for precision measurement in cosmology \citep[e.g.,][]{Planck18CosmoParams,Planck18PNG,Planck18Inflation,Cabass+2022}
+used for precision measurement in cosmology \citep[e.g.,][]{Eisenstein+2005BAO,Planck18PNG,Planck18Inflation,Cabass+2022}
 are all permutation-invariant.
@@
 
+\CH{moved this to here from farther down}
+The area of a triangle is not only invariant to permutation, but also to translation, rotation, and reflection.
+Formulae \eqref{eq:sine} and \eqref{eq:Heron}, by virtue of referencing only side-lengths and angles, are explicitly invariant to these other symmetries.
+While true, it is not obvious that our solution \eqref{eq:result} is so invariant.
+An explicit proof of this is left to the reader.
+
 Note that a triangle defined by three points has one-half the area of the parallelogram spanned by the vectors $\vec{a} = \vec{r}_2 - \vec{r}_1$ and $\vec{b} = \vec{r}_3 - \vec{r}_1$.
@@
 \secbreak
-We now consider two more applications of the Deep Sets form to Heron's formula and to the potential energy of a system of point masses.
-We stated above that Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, and yet it is not represented in Deep Sets form. Consider instead
-\begin{equation}
-\label{eq:HeronDeepSets}
-    \Delta = \frac{1}{4} \, \sqrt{
-        \left( a^2 + b^2 + c^2 \right)^2
-        - 2 \left( a^4 + b^4 + c^4 \right)
-    }~.
-\end{equation}
-It is simple to show that this is exactly equivalent to Heron's formula when expanded, but it is now in Deep Sets form.\footnote{
-    This is not novel.
+% We now consider two more applications of the Deep Sets form to Heron's formula and to the potential energy of a system of point masses.
+
+% \CH{I want to cut this?}
+% We stated above that Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, and yet it is not represented in Deep Sets form. Consider instead
+% \begin{equation}
+% \label{eq:HeronDeepSets}
+%     \Delta = \frac{1}{4} \, \sqrt{
+%         \left( a^2 + b^2 + c^2 \right)^2
+%         - 2 \left( a^4 + b^4 + c^4 \right)
+%     }~.
+% \end{equation}
+% It is simple to show that this is exactly equivalent to Heron's formula when expanded, but it is now in Deep Sets form.\footnote{
+%     This is not novel.
 Heron's formula is a special case of Brahmagupta's formula, which gives the area $K$ of a convex cyclic quadrilateral in terms of its four side lengths $a, b, c, d$ \citep{coolidge1939}.
-    On Wikipedia, it has been written
-    \begin{equation}
-        K = \frac{1}{4} \, \sqrt{
-            \left( a^2 + b^2 + c^2 + d^2 \right)^2
-            + 8 \, a\,b\,c\,d
-            - 2 \left( a^4 + b^4 + c^4 + d^4 \right)
-        }~,
-    \end{equation}
-    which is exactly \eqref{eq:HeronDeepSets} if $d$ is zero.
-    This expression is in Deep Sets form, with the product $a \, b \, c \, d$ 
-    representable as $\exp( \sum_{x \in \{ a,b,c,d \}} \log x )$.
-}
+%     On Wikipedia, it has been written
+%     \begin{equation}
+%         K = \frac{1}{4} \, \sqrt{
+%             \left( a^2 + b^2 + c^2 + d^2 \right)^2
+%             + 8 \, a\,b\,c\,d
+%             - 2 \left( a^4 + b^4 + c^4 + d^4 \right)
+%         }~,
+%     \end{equation}
+%     which is exactly \eqref{eq:HeronDeepSets} if $d$ is zero.
+%     This expression is in Deep Sets form, with the product $a \, b \, c \, d$ 
+%     representable as $\exp( \sum_{x \in \{ a,b,c,d \}} \log x )$.
+% }
+% 
+While the triangle area is in a form similar to some physics problems, the polynomial methods we employ to find the Deep Sets form for the triangle area \eqref{eq:result} are not so relevant for those problems.
+More relevant instead is to consider a $\phi$ which Fourier transforms the input data.
+In doing so, translation symmetry is encoded as well as permutation symmetry, since the Fourier basis $\exp (i\, \vec{k}\cdot\vec{x})$ is the eigenbasis of the translation operator.
 
-\CH{feels out of place...} \Hogg{This can be deleted.}
-The area of a triangle is not only invariant to permutation, but also to translation, rotation, and reflection.
-Formulae \eqref{eq:sine} and \eqref{eq:Heron}, by virtue of referencing only side-lengths and angles, are explicitly invariant to these other symmetries.
-While true, it is not obvious that our solution \eqref{eq:result} is so invariant.
-An explicit proof of this is left to the reader.
-
-While the triangle area is related to physics problems, the polynomial methods we employ to find the Deep Sets form for the triangle area \eqref{eq:result} are not relevant in a physical context.
-More relevant instead is to consider a $\phi$ which Fourier transforms the input data.
-In doing so, translation symmetry is encoded as well, since the Fourier basis $\exp (i\, \vec{k}\cdot\vec{x})$ is the eigenbasis of the translation operator.
-Below, we show that this can be used to produce a Deep Sets form for the gravitational potential energy of a system of point masses.
-
-Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
+For example, this can be used to produce a Deep Sets form for the gravitational potential energy of a system of point masses.
+Explicitly, consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
 The total energy is given in Deep Sets form by
@@
 
+We can even use these methods to find another formula for the area $\Delta$ of a triangle.
+\begin{align}
+    \phi_{\vec{k}} (\vec{r}_j) &= e^{-i \, \vec{k} \cdot \vec{r}_j}~,
+    \quad
+    \mathcal{F}[\rho](\vec{k}) = \sum_{j=1}^{3} \phi_{\vec{k}}(\vec{r}_j)~,
+    \nonumber\\
+    \mathcal{P}(\vec{r}) &= (2\pi)^d \, \mathcal{F}^{-1} \left[ 
+    \abs{ \mathcal{F}[\rho] }^2 \right]~,
+    \quad
+    F_m = \int_{\mathbb{R}^d} \d^d r \, \mathcal{P}(\vec{r}) \, \frac{\abs{\vec{r}}^m}{2}~,
+    \nonumber\\
+    \Delta
+    &= \frac{1}{4} \, \sqrt{F_2^2 - 2 \, F_4}~.
+\end{align}
+A complete derivation is provided in the \nameref{sec:appendix}.
+Briefly, $\mathcal{P}$ is the Patterson function of X-ray crystallography \citep{Patterson1934} which has 6 delta function peaks corresponding to the three displacement vectors and their negatives. 
+Multiplying the Patterson function by a function of the magnitude $r \equiv \vec{r}$ and integrating over all space gives a sum of the values of the function evaluated at the three side lengths. 
+Choosing this function properly allows us to write the power sums of the side lengths $F_m = a^m+b^m+c^m$ in this form.
+
+To verify this formula, note that we can write it in terms of the side lengths as
+\begin{equation}
+\label{eq:HeronDeepSets}
+    \Delta = \frac{1}{4} \, \sqrt{
+        F_2^2
+        - 2 \, F_4
+    }
+    = \frac{1}{4} \, \sqrt{
+        \left( a^2 + b^2 + c^2 \right)^2
+        - 2 \, \left( a^4 + b^4 + c^4 \right)
+    }~.
+\end{equation}
+This formula is equivalent to Heron's formula \eqref{eq:Heron} by expansion.
+Note that it is also in Deep Sets form as a function of the side lengths.
+
+\CH{I don't like this ending ... thinking about other ways to conclude}
+
 We leave the reader with the following questions. 
@@
 Does another area formula exist for these other aggregators?
-We have discussed the general value of a map function $\phi$ which Fourier transforms the points;\footnote{%
-\Hogg{The Deep Sets version of the Fourier transform is reminiscent of a positional encoding (CITE), similar to those used to preserve ordering information in (otherwise permutation invariant) transformers (CITE).}}
-does a formula like \eqref{eq:EnergyDeepSets} exist for the area of a triangle?
+% We have discussed the general value of a map function $\phi$ which Fourier transforms the points;\footnote{%
+% \Hogg{The Deep Sets version of the Fourier transform is reminiscent of a positional encoding (CITE), similar to those used to preserve ordering information in (otherwise permutation invariant) transformers (CITE).}}
+% does a formula like \eqref{eq:EnergyDeepSets} exist for the area of a triangle?
 
@@
 This project would not have been possible without originating inspiration and discussions with Soledad Villar (JHU).
-The authors thank Ben Blum-Smith (JHU), Jonathan Goodman (NYU), Wilson Gregory (JHU), Teresa Huang (Flatiron), Mike O'Neil (NYU), Alexander Novara (NYU), Ben Wandelt (JHU), Kaze Wong (JHU), and the Blanton--Hogg group meeting at NYU for valuable discussions.
+The authors thank
+  Ben Blum-Smith (JHU),
+  Jonathan Goodman (NYU),
+  Wilson Gregory (JHU),
+  Teresa Huang (Flatiron),
+  Alexander Novara (NYU),
+  Mike O'Neil (NYU),
+  Oliver Philcox (Columbia),
+  Ben Wandelt (JHU),
+  Kaze Wong (JHU),
+and the Blanton--Hogg group meeting at NYU for valuable discussions.
 The Flatiron Institute is a division of the Simons Foundation.
-\Hogg{CH: You need to acknowledge a grant number here, I believe.}
+CH is supported by the National Science Foundation Graduate Research Fellowship under Grant Number DGE-2234660.
 The authors also note that, while we believe the formula is (barely) novel, it is difficult to be sure, given the challenge of conducting a complete literature review for this ancient topic.\par}
@@
 
+\paragraph{Fourier triangle area}
+Suppose our triangle is specified by three vertices $\vec{r}_1,\vec{r}_2,\vec{r}_3 \in \mathbb{R}^d$. We represent this distribution by the field
+\begin{equation}
+    \rho(\vec{r})
+    = \sum_i \, \delta(\vec{r} - \vec{r}_i)~.
+\end{equation}
+Fourier transforming this field gives
+\begin{equation}
+    \tilde{\rho}(\vec{k})
+    =
+    e^{-i \, \vec{k} \cdot \vec{r}_1} +
+    e^{-i \, \vec{k} \cdot \vec{r}_2} +
+    e^{-i \, \vec{k} \cdot \vec{r}_3}~.
+\end{equation}
+Multiplying by its conjugate to find the magnitude squared, we find
+\begin{equation}
+    \abs{\tilde{\rho}(\vec{k})}^2
+    =
+    3 +
+    e^{-i \, \vec{k} \cdot \vec{r}_{12}} +
+    e^{+i \, \vec{k} \cdot \vec{r}_{12}} +
+    e^{-i \, \vec{k} \cdot \vec{r}_{13}} +
+    e^{+i \, \vec{k} \cdot \vec{r}_{13}} +
+    e^{-i \, \vec{k} \cdot \vec{r}_{23}} +
+    e^{+i \, \vec{k} \cdot \vec{r}_{23}}~,
+\end{equation}
+defining $\vec{r}_{ij} \equiv \vec{r}_i - \vec{r}_j$.
+The inverse Fourier transform of this is
+\begin{align}
+    \mathcal{P}(\vec{r})
+    &= (2\pi)^d \, \mathcal{F}^{-1} \left[ \abs{\tilde{\rho}}^2 \right]
+    \nonumber\\ 
+    &= 3 \,\delta(\vec{r})
+    + \delta(\vec{r} - \vec{r}_{12})
+    + \delta(\vec{r} + \vec{r}_{12})
+    \nonumber\\
+    &\quad + \delta(\vec{r} - \vec{r}_{13})
+    + \delta(\vec{r} + \vec{r}_{13})
+    + \delta(\vec{r} - \vec{r}_{23})
+    + \delta(\vec{r} + \vec{r}_{23})~.
+\end{align}
+This is known as the Patterson function.
+
+By multiplying the Patterson function $\mathcal{P}$ by a function of $r \equiv \abs{\vec{r}}$ and integrating over all space, we can extract sums of functions of the side lengths. Explicitly,
+\begin{align}
+    \int_{\mathbb{R}^d} \d^d r \, \mathcal{P}(\vec{r}) \, f(\abs{\vec{r}})
+    &= 3 \, f(0) +
+        2 \, f(r_{12}) +
+        2 \, f(r_{13}) +
+        2 \, f(r_{23}) \nonumber\\
+    &= 3 \, f(0) +
+        2 \, f(a) +
+        2 \, f(b) +
+        2 \, f(c)~.
+\end{align}
+In particular, we can define the $k^{\rm th}$ power sum of the side lengths
+\begin{equation}
+    F_k
+    = \int_{\mathbb{R}^d} \d^d r \, \mathcal{P}(\vec{r}) \, \frac{\abs{\vec{r}}^k}{2}
+    = a^k + b^k + c^k~,
+\end{equation}
+which allows us to write the triangle area as
+\begin{equation}
+    \Delta
+    = \frac{1}{4} \, \sqrt{\left(a^2+b^2+c^2\right)^2 - 2\, \left(a^4+b^4+c^4\right)}
+    = \frac{1}{4} \, \sqrt{F_2^2 - 2\, F_4}~.
+\end{equation}
+
 \end{document}

### 2025-03-27 paper/main.tex
lots of improvements by ch

@@
 % -----------
-% - CH: Show that there is NOT an exact solution at polynomial degree 2 order.
-% - HOGG: Discuss alternatives like Novara's; And alternatives like Kaze's.
-Positional encodings!
-CH: Describe the connection to positional encodings (in, eg, transformers).
-Oh and cosmology and power spectrum and bispectrum.
+% - [x] CH will mention positional encodings.
+% - [x] CH will consider changing the outer function rho() to sigma().
+% - [x] CH should define \d as \mathrm{d} and use it for differentials.
 
@@
 % - Be careful about "not translationally invariant" -- we mean "not explicitly translationally invariant" and so on. Explicitness is what this paper is about.
+% - Think about brackets and parens?
+% - All matrices are \mathsf. All vectors are \vec.
 
@@
 \pagestyle{myheadings}
-\renewcommand{\paragraph}[1]{\par\addvspace{1.5ex}\noindent\textsl{#1}~---}
+\renewcommand{\paragraph}[1]{\par\addvspace{1em}\noindent\textsl{#1}~---}
 \newcommand{\secbreak}{\bigskip{\centering\footnotesize $\triangle~~~\triangle~~~\triangle$\par}\bigskip\noindent}
@@
 
-\medskip
+\smallskip
 \paragraph{Abstract}
@@
 The area of a triangle is a permutation-invariant function of the locations $\vec{r}_i$ of the three corners $1\leq i\leq 3$.
-We find the formula for the area of a triangle that is explicitly in Deep Sets form.
-This project was motivated by questions about the fundamental computational complexity of $n$-point statistics in cosmology; that said, no insights of any kind were gained from these results.
+We find the polynomial formula for the area of a triangle that is explicitly in Deep Sets form.
+This project was motivated by questions about the fundamental computational complexity of $n$-point statistics in cosmology;
+that said, no insights of any kind were gained from these results.
 
@@
 Formula~\eqref{eq:cross} generalizes to triangles in higher dimensional spaces, using the fact that the cross or wedge product has a magnitude equal to the area of the parallelogram spanned by $\vec{a}$ and $\vec{b}$.
-Formula~\eqref{eq:polynomial} gives a polynomial in terms of the Cartesian coordinates of the corners in two dimensions; it is equal to an expansion of \eqref{eq:cross}.
+Formula~\eqref{eq:polynomial} gives the absolute value of a polynomial in terms of the Cartesian coordinates of the corners in two dimensions; it is equal to an expansion of \eqref{eq:cross}.
 Formula~\eqref{eq:Heron} is Heron's formula, which elegantly computes the area from the side lengths alone.
@@
 
-The area of a triangle has many symmetries.
+The area of a triangle obeys many symmetries.
 It is invariant to translation, rotation, and reflection, as well as to the labeling of sides or corners.
@@
 It has long been known that any permutation-invariant (or symmetric) polynomial of scalar data $f(\{r_i\})$ can be written as a polynomial of the power sums $\sum_i r_i^m$ \citep{Waring1782}.
-Further, symmetric polynomials of any-dimensional data are decomposable into \emph{polarizations} of the scalar symmetric polynomials \citep{Weyl1939}.
-Together, these imply that a permutation-invariant polynomial of vector data is a polynomial in the polarizations of the power sums; e.g., the power sums of the components of the vectors.
+Further, symmetric polynomials of vector data $\{ \vec{r}_i \}$ are decomposable into polarizations of the scalar symmetric polynomials \citep{Weyl1939}, which look like sums of the form $\sum_i \, (\vec{r}_{i})_{j}^{m_j} \cdots (\vec{r}_{i})_{k}^{m_k}$, where $(\vec{r}_i)_j$ is the $j^{\rm th}$ component of $\vec{r}_i$.
+Together, these imply that a permutation-invariant polynomial of vector data is a polynomial in the polarizations of the power sums.
 This result is \emph{universal}, in the sense that it is true for all symmetric polynomials.
 It also matches onto the Deep Sets form, where the polarizations of the power sums are the components of the vector $\sum_i \phi(\vec{r}_i)$, which are combined in a polynomial $\rho$.
-
-% Specifically, it is known that the permutation-invariant polynomials on a set of any-dimensional data are representable in terms of the polarizations of the permutation-invariant polynomials on a corresponding set of one-dimensional data \citep{Weyl1939}.
-% Further, the permutation-invariant polynomials on a set of one-dimensional data are representable in terms of the power sums $\sum_{i=1}^{n} r_i^j$ \citep{Waring1782}.
-% Note that this decomposition is \emph{universal}, in the sense that it is true for \emph{all} symmetric polynomials of any-dimensional data. 
-% This also matches onto the Deep Sets form, as it states that a permutation-invariant polynomial of any-dimensional data $f(\{r_i\})$ can be decomposed into a polynomial ($\rho$) of polarizations of the power sums ($\sum_i \phi(\vec{r}_i)$).
-% \CH{this is mostly plagiarized from BBS email; need to understand it better to massage it.} 
-% \Hogg{Important to emphasize that there is a proof of universality, and it's old.}
 
@@
 \begin{equation}
+\label{eq:EnergyNaive}
     U = -G \, \sum_{i=1}^{N} \sum_{j=i+1}^{N} \frac{m_i \, m_j}{\abs{\vec{r}_i - \vec{r}_j}} ~.
@@
 
-Because all these (computationally expensive) statistics are permutation-invariant, they must be possible to write in Deep Sets form.
+Because all these (computationally expensive) statistics are permutation-invariant, it must be possible to write them in Deep Sets form.
 At the same time, Deep Sets form \eqref{eq:DeepSets} looks like it is computable in linear time.
@@
 So, there may be benefits to finding Deep Sets forms for problems that depend on permutation-invariant pairs, triplets, or $k$-tuplets of points.
-The simplest such problem is the area of a triangle, which depends on pairs of points via
+The simplest such problem is the area of a triangle, 
+which depends on pairs of points via
 side lengths, as in Heron's formula \eqref{eq:Heron};
 angles, as in \eqref{eq:sine};
-or point displacements, as in \eqref{eq:cross}.
+or point displacements, as in \eqref{eq:cross};
+or as a sum over $x_i \, y_j - x_j \, y_i$, as in \eqref{eq:polynomial}.
+% which can be written as a sum over pairs of points as
+% \begin{equation}
+%     \label{eq:pairwiseSums}
+%     \Delta = \frac{1}{2} \, \bigg|\,
+%     \sum_{i=1}^{3} \sum_{j=i+1}^{3}
+%     x_i \, y_j - x_j \, y_i
+%     \,\bigg|~,
+% \end{equation}
+% very similar in form to \eqref{eq:EnergyNaive}.
 
@@
     \Delta^2 = \ &
-    \frac{3}{4} \, \big( \sum_{i=1}^{3} x_i^2 \big) \, \big( \sum_{i=1}^{3} y_i^2 \big)
-    - \frac{3}{4} \, \big( \sum_{i=1}^{3} x_i \, y_i \big)^2 
-    + \frac{2}{4} \, \big( \sum_{i=1}^{3} x_i \, y_i \big) \, \big( \sum_{i=1}^{3} x_i \big) \, \big( \sum_{i=1}^{3} y_i \big)
+    \frac{3}{4}
+        \,\big( \sum_{i=1}^{3} x_i^2 \big)
+        \,\big( \sum_{i=1}^{3} y_i^2 \big)
+    - \frac{3}{4}
+        \,\big( \sum_{i=1}^{3} x_i \, y_i \big)^2 
+    + \frac{2}{4}
+        \,\big( \sum_{i=1}^{3} x_i \, y_i \big)
+        \,\big( \sum_{i=1}^{3} x_i \big)
+        \,\big( \sum_{i=1}^{3} y_i \big)
     \nonumber\\ &
-    - \, \frac{1}{4} \, \big( \sum_{i=1}^{3} y_i^2 \big) \, \big( \sum_{i=1}^{3} x_i \big)^2
-    - \, \frac{1}{4} \, \big( \sum_{i=1}^{3} x_i^2 \big) \, \big( \sum_{i=1}^{3} y_i \big)^2
+    - \frac{1}{4}
+        \,\big( \sum_{i=1}^{3} x_i^2 \big)
+        \,\big( \sum_{i=1}^{3} y_i \big)^2
+    - \frac{1}{4}
+        \,\big( \sum_{i=1}^{3} y_i^2 \big)
+        \,\big( \sum_{i=1}^{3} x_i \big)^2
      ~.
@@
     Indeed, the formulae \eqref{eq:school} through \eqref{eq:Heron} either (implicitly) involve square roots of polynomials, or else absolute value signs, such that the corresponding pure polynomial expression would be for the squared area.}
-is the primary contribution of this \documentname, along with the generalization to higher-dimensional vectors $\vec{r}_i$ below in \eqref{eq:quadrature}.
+is the primary contribution of this \documentname, along with the generalizations below.
 
@@
 \begin{gather}
-    \phi(\vec{r}_i) = (
+    \label{eq:resultDeepSets}
+    \phi(\vec{r}_i) = \big[
         x_i, \,
@@
         y_i^2
-    ), \quad
-    (
+    \big], \quad
+    \big[
         F_{10}, \,
@@
         F_{02}
-    ) = \sum_{i=1}^{3} \phi(\vec{r}_i),
+    \big] = \sum_{i=1}^{3} \phi(\vec{r}_i),
     \\
@@
     + \frac{2}{4} \, F_{11} \, F_{10} \, F_{01}
-    - \frac{1}{4} \, F_{02} \, F_{10}^2
-    - \frac{1}{4} \, F_{20} \, F_{01}^2~.
+    - \frac{1}{4} \, F_{20} \, F_{01}^2
+    - \frac{1}{4} \, F_{02} \, F_{10}^2~.
     \nonumber
 \end{gather}
+Notice that the runtime is formally linear in the number of points, since $\phi$ and $\rho$ can be considered to have constant time complexity.
+However, because the dimensionality of $\phi$ is larger than the number of points, evaluating the triangle area in this way will not necessarily be faster than a loop over pairs of points.
 
-\CH{moved this to here from farther down}
 The area of a triangle is not only invariant to permutation, but also to translation, rotation, and reflection.
@@
 
-Note that a triangle defined by three points has one-half the area of the parallelogram spanned by the vectors $\vec{a} = \vec{r}_2 - \vec{r}_1$ and $\vec{b} = \vec{r}_3 - \vec{r}_1$.
-Multiplying both sides by 4 reveals that \eqref{eq:result} is an integer polynomial of permutation-invariant factors for the squared area of the parallelogram.
-Using this fact, we have written code to perform a symbolic regression that recovers our result.\footnote{
-    The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
-}
+Recall that a triangle with vertices $\{ \vec{r}_1, \vec{r}_2, \vec{r}_3 \}$ has one-half the area of the parallelogram spanned by the displacement vectors $\{ \vec{r}_2 - \vec{r}_1, \vec{r}_3 - \vec{r}_1 \}$.
+So our result \eqref{eq:result} is actually an integer polynomial for the squared area of this parallelogram.
+Using this fact, we have written code that performs a symbolic regression to recover our result.\footnote{
+    The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.} 
+The code performs least-squares regression to find a fourth-order homogeneous polynomial of the multi-power sums $\sum_i \, x_i^m \, y_i^n$, and it requires that the result gives the squared areas to machine precision after rounding the coefficients to the nearest integers.
+Using and extending the code, we also find several generalizations to our formula.
 
-In this symbolic regression, we consider a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
-All the possible monomials are made of permutation-invariant factors of the form 
-\begin{equation}
-F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n~;
-\end{equation}
-each such a factor has degree $m + n$.
-The invariant monomials of degree $d$ are then products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
-We generate all factors $F_{mn}$ with degree $1 \leq m + n \leq 4$, then form all products of these factors with total degree four.
-However, these monomials are not all linearly-independent;
-the five monomials made of a single degree-four factor ($F_{40}$, $F_{31}$, $F_{22}$, $F_{13}$, $F_{04}$) can be written as a linear combination of the other monomials and are thus removed.
-This leaves 28 linearly-independent monomials.
+% Note that a triangle defined by three points has one-half the area of the parallelogram spanned by the vectors $\vec{a} = \vec{r}_2 - \vec{r}_1$ and $\vec{b} = \vec{r}_3 - \vec{r}_1$.
+% Multiplying both sides by 4 reveals that \eqref{eq:result} is an integer polynomial of permutation-invariant factors for the squared area of the parallelogram.
+% Using this fact, we have written code to perform a symbolic regression that recovers our result.\footnote{
+%     The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
+% }
 
-We fit a linear combination of the monomials to a data set of generated triples of points, drawn from a two-dimensional Gaussian whose mean was also drawn from a Gaussian in a two-dimensional space.
-Each triple defines a parallelogram whose squared area we use as a label for regression.
-We then perform least-squares regression to fit a linear combination of the invariant monomials.
-We require that the labels are reproduced to machine precision after rounding the coefficients to the nearest integers.
+% In this symbolic regression, we consider a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
+% All the possible monomials are made of permutation-invariant factors of the form 
+% \begin{equation}
+% F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n~;
+% \end{equation}
+% each such a factor has degree $m + n$.
+% The invariant monomials of degree $d$ are then products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
+% We generate all factors $F_{mn}$ with degree $1 \leq m + n \leq 4$, then form all products of these factors with total degree four.
+% However, these monomials are not all linearly-independent;
+% the five monomials made of a single degree-four factor ($F_{40}$, $F_{31}$, $F_{22}$, $F_{13}$, $F_{04}$) can be written as a linear combination of the other monomials and are thus removed.
+% This leaves 28 linearly-independent monomials.
 
-If the triangle is embedded in a higher-dimensional space with the positions of the three vertices given by $\vec{r}_i \in \mathbb{R}^{d}$ for $d \geq 2$, the area of the triangle can be determined by summing in quadrature the areas of the triangles made by all mutually orthogonal two-dimensional projections of the points:
+% We fit a linear combination of the monomials to a data set of generated triples of points, drawn from a two-dimensional Gaussian whose mean was also drawn from a Gaussian in a two-dimensional space.
+% Each triple defines a parallelogram whose squared area we use as a label for regression.
+% We then perform least-squares regression to fit a linear combination of the invariant monomials.
+% We require that the labels are reproduced to machine precision after rounding the coefficients to the nearest integers.
+% The result exactly matches equation \eqref{eq:result}.
+
+If the triangle is embedded in a higher-dimensional space with vertex positions given by $\vec{r}_i \in \mathbb{R}^{d}$ for $d \geq 2$, the area of the triangle can be determined by summing in quadrature the areas of the triangles made by all pairs of mutually orthogonal projections of the points:
 \begin{equation}
@@
 
-Our method also extends to the volume of a simplex. In three dimensions, four vectors $\{ \vec{r}_1, \vec{r}_2, \vec{r}_3, \vec{r}_4 \}$ determine the vertices of a simplex. 
-Like the area of a triangle is one-half the area of a related parallelogram, the volume of a simplex is one-sixth the volume of the parallelepiped spanned by the vectors $\{ \vec{r}_2 - \vec{r}_1, \vec{r}_3 - \vec{r}_1, \vec{r}_4 - \vec{r}_1 \}$.
-Expecting an integer solution once again, we modify the symbolic regression to generate 4-tuplets of points, using the squared volumes of the parallelepipeds they produce as labels.
-The invariant factors then have three indices, $F_{\ell m n} = \sum_i x_i^\ell \, y_i^m \, z_i^n$, and we form all invariant monomials of degree six.
-The result, verifiable by computational algebra, is that the volume $V$ of the simplex is given by
-\begin{align}
-(6 V)^2 = &
-+8 \, F_{110} \, F_{101} \, F_{011}
-+4 \, F_{200} \, F_{020} \, F_{002}
-\nonumber \\ &
--4 \, F_{200} \, F_{011}^2
--4 \, F_{020} \, F_{101}^2
--4 \, F_{002} \, F_{110}^2
-\nonumber \\ &
-+2 \, F_{200} \, F_{011} \, F_{010} \, F_{001}
-+2 \, F_{020} \, F_{101} \, F_{100} \, F_{001}
-+2 \, F_{002} \, F_{110} \, F_{100} \, F_{010}
-\nonumber \\ &
--2 \, F_{110} \, F_{101} \, F_{010} \, F_{001}
--2 \, F_{110} \, F_{011} \, F_{100} \, F_{001}
--2 \, F_{101} \, F_{011} \, F_{100} \, F_{010}
-\nonumber \\ &
-+1 \, F_{110}^2 \, F_{001}^2
-+1 \, F_{101}^2 \, F_{010}^2
-+1 \, F_{011}^2 \, F_{100}^2
-\nonumber \\ &
--1 \, F_{200} \, F_{020} \, F_{001}^2
--1 \, F_{200} \, F_{002} \, F_{010}^2
--1 \, F_{020} \, F_{002} \, F_{100}^2~.
-\label{eq:simplex}
-\end{align}
-There are many similarities between this and our result for the triangle area \eqref{eq:result}, which we believe may indicate a way to generalize this result to the hypervolume of a $d$-simplex.
+The result can also be extended to the volume of a $d$-dimensional simplex.
+A $d$-simplex is defined by $d+1$ vectors $\{ \vec{r}_1, \dots, \vec{r}_{d+1} \}$ in $\mathbb{R}^d$.
+Then, the volume $V$ of the $d$-simplex is given by
+\begin{equation}
+    V^2 = \frac{1}{(d!)^2} \, \det \mathsf{M}~,
+    \quad
+    \mathsf{M} \equiv \left[\begin{array}{cccc}
+        \sum_i \, r_{i1}^2
+            & \cdots 
+            & \sum_i \, r_{i1} \, r_{id}
+            & \sum_i \, r_{i1} \\
+        \vdots
+            & \ddots 
+            & \vdots
+            & \vdots \\
+        \sum_i \, r_{id} \, r_{i1}
+            & \cdots 
+            & \sum_i \, r_{id}^2
+            & \sum_i \, r_{id} \\
+        \sum_i \, r_{i1}
+            & \cdots 
+            & \sum_i \, r_{id}
+            & d + 1 \\
+    \end{array}\right]~,
+    \label{eq:dsimplex}
+\end{equation}
+where $r_{ij}$ is the $j^{\rm th}$ component of $\vec{r}_i$.\footnote{%
+    This generalization was pointed out by Kendrick Smith (Perimeter Institute).}
+A proof is deferred to the \nameref{sec:appendix}.
+This is in Deep Sets form, as the components of $\mathsf{M}$ are all of the form $\sum_i \phi(\vec{r}_i)$.
+
+A few notes we'd make about this formula are as follows.
+This formula with $d=2$ gives back exactly our result \eqref{eq:result}, since the 2-simplex is a triangle and the 2-volume is its area.
+The components of $\mathsf{M}$ are in fact the polarizations of the first and second power sums, so the determinant of $\mathsf{M}$ gives the polynomial in these power sums that we were guaranteed by invariant theory.
+The runtime complexity can again be considered to be linear, with $\rho$ and $\phi$ having constant complexity.
+However, the dimensionality of $\phi$ is $(d+1)(d+2)/2$, the number of unique components in $\mathsf{M}$, so there are no real performance improvements.
+As with the relationship between triangles and parallelograms, the volume of the $d$-parallelotope spanned by the displacement vectors $\{ \vec{r}_2 - \vec{r}_1, \dots, \vec{r}_{d+1} - \vec{r}_1 \}$ is $d!$ the volume of a $d$-simplex; thus \eqref{eq:dsimplex} is an integer polynomial for the squared volume of this $d$-parallelotope.
+
+% Our method also extends to the volume of a simplex. In three dimensions, four vectors $\{ \vec{r}_1, \vec{r}_2, \vec{r}_3, \vec{r}_4 \}$ determine the vertices of a simplex. 
+% Like the area of a triangle is one-half the area of a related parallelogram, the volume of a simplex is one-sixth the volume of the parallelepiped spanned by the displacement vectors $\{ \vec{r}_2 - \vec{r}_1, \vec{r}_3 - \vec{r}_1, \vec{r}_4 - \vec{r}_1 \}$.
+% Expecting an integer solution once again, we modify the symbolic regression to generate 4-tuplets of points, using the squared volumes of the parallelepipeds they produce as labels.
+% The invariant factors then have three indices, $F_{\ell m n} = \sum_i x_i^\ell \, y_i^m \, z_i^n$, and we form all invariant monomials of degree six.
+% The result, verifiable by computational algebra, is that the volume $V$ of the simplex is given by
+% \begin{align}
+% (6 V)^2 = &
+% +8 \, F_{110} \, F_{101} \, F_{011}
+% +4 \, F_{200} \, F_{020} \, F_{002}
+% \nonumber \\ &
+% -4 \, F_{200} \, F_{011}^2
+% -4 \, F_{020} \, F_{101}^2
+% -4 \, F_{002} \, F_{110}^2
+% \nonumber \\ &
+% +2 \, F_{200} \, F_{011} \, F_{010} \, F_{001}
+% +2 \, F_{020} \, F_{101} \, F_{100} \, F_{001}
+% +2 \, F_{002} \, F_{110} \, F_{100} \, F_{010}
+% \nonumber \\ &
+% -2 \, F_{110} \, F_{101} \, F_{010} \, F_{001}
+% -2 \, F_{110} \, F_{011} \, F_{100} \, F_{001}
+% -2 \, F_{101} \, F_{011} \, F_{100} \, F_{010}
+% \nonumber \\ &
+% +1 \, F_{110}^2 \, F_{001}^2
+% +1 \, F_{101}^2 \, F_{010}^2
+% +1 \, F_{011}^2 \, F_{100}^2
+% \nonumber \\ &
+% -1 \, F_{200} \, F_{020} \, F_{001}^2
+% -1 \, F_{200} \, F_{002} \, F_{010}^2
+% -1 \, F_{020} \, F_{002} \, F_{100}^2~.
+% \label{eq:simplex}
+% \end{align}
+% There are many similarities between this and our result for the triangle area \eqref{eq:result}, which we believe may indicate a way to generalize this result to the hypervolume of a $d$-simplex.
 
 \secbreak
-% We now consider two more applications of the Deep Sets form to Heron's formula and to the potential energy of a system of point masses.
+While the triangle area, by virtue of being a sum over pairs of points, is in a form similar to some physics problems, the polynomial methods we employ are not so relevant for those problems.
+However, in physics it is common to encode a set of points by defining the density field to be a sum of delta functions at the points.
+Notice that the Fourier transform of this field at a wavevector $\vec{k}$ is
+\begin{equation}
+    \mathcal{D}_{\vec{k}} = \mathcal{F} \,\bigg\{
+        \sum_{j=1}^{N} \delta(\vec{r} - \vec{r}_j)
+    \bigg\}
+    = \sum_{j=1}^{N} e^{-i \, \vec{k} \cdot \vec{r}_j}~,
+\end{equation}
+which is exactly in the form of $\sum_i \phi(\vec{r}_i)$.
+Hence, any function of the Fourier transform $\mathcal{D}$ of the density field is technically in Deep Sets form, and has permutation symmetry guaranteed.
+This is physically useful also because the Fourier basis $\exp (i\, \vec{k}\cdot\vec{x})$ is the eigenbasis of the translation operator, so this choice of the map $\phi$ naturally encodes translation symmetry.\footnote{%
+%
+Interestingly, a transformation like $\phi(r) = \exp(i\,k\,r)$ is used widely in natural language processing.
+Most large language models, such as GPT-3 \citep{Brown2020GPT3}, are transformer models \citep{Vaswani2017Attention}, which are fundamentally permutation invariant.
+In order to use the positional information of tokens (often words), the tokens are augmented by a ``positional encoding.''
+One of the most popular choices for this encoding is the Rotary Positional Encoding \citep[RoPE;][]{Su2024Roformer}, which takes the word embedding vector $\vec{x} \in \mathbb{R}^{d}$ for the token in position $m$ and multiplies it by $e^{i \, m \, \theta}$ for $d/2$ preset constant values $\theta \in \mathbb{R}$.
+While this actually \emph{breaks} permutation invariance rather than enforcing it (since the encoding explicitly involves the index $m$), it underscores the broad applications of these transformations.
+}
 
-% \CH{I want to cut this?}
-% We stated above that Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, and yet it is not represented in Deep Sets form. Consider instead
-% \begin{equation}
-% \label{eq:HeronDeepSets}
-%     \Delta = \frac{1}{4} \, \sqrt{
-%         \left( a^2 + b^2 + c^2 \right)^2
-%         - 2 \left( a^4 + b^4 + c^4 \right)
-%     }~.
-% \end{equation}
-% It is simple to show that this is exactly equivalent to Heron's formula when expanded, but it is now in Deep Sets form.\footnote{
-%     This is not novel.
-Heron's formula is a special case of Brahmagupta's formula, which gives the area $K$ of a convex cyclic quadrilateral in terms of its four side lengths $a, b, c, d$ \citep{coolidge1939}.
-%     On Wikipedia, it has been written
-%     \begin{equation}
-%         K = \frac{1}{4} \, \sqrt{
-%             \left( a^2 + b^2 + c^2 + d^2 \right)^2
-%             + 8 \, a\,b\,c\,d
-%             - 2 \left( a^4 + b^4 + c^4 + d^4 \right)
-%         }~,
-%     \end{equation}
-%     which is exactly \eqref{eq:HeronDeepSets} if $d$ is zero.
-%     This expression is in Deep Sets form, with the product $a \, b \, c \, d$ 
-%     representable as $\exp( \sum_{x \in \{ a,b,c,d \}} \log x )$.
-% }
-% 
-While the triangle area is in a form similar to some physics problems, the polynomial methods we employ to find the Deep Sets form for the triangle area \eqref{eq:result} are not so relevant for those problems.
-More relevant instead is to consider a $\phi$ which Fourier transforms the input data.
-In doing so, translation symmetry is encoded as well as permutation symmetry, since the Fourier basis $\exp (i\, \vec{k}\cdot\vec{x})$ is the eigenbasis of the translation operator.
-
-For example, this can be used to produce a Deep Sets form for the gravitational potential energy of a system of point masses.
+Because any function of the Fourier transform $\mathcal{D}$ of the density field is automatically in Deep Sets form,
+we can find a Deep Sets form for the gravitational potential energy of a system of point masses.
 Explicitly, consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
@@
 \begin{align}
-    \phi_{\vec{k}}(m_j, \vec{r}_j)
-    &=m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}
+    &\phi_{\vec{k}}(m_j, \vec{r}_j)
+    =m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}
     ~,
@@
     \sum_{i=1}^{N} \phi_{\vec{k}}(m_i, \vec{r}_i)
-    = \mathcal{F}[\rho](\vec{k})~,
+    = \mathcal{D}_{\vec{k}}~,
     \nonumber\\
-    U \, \big( \sum_{i=1}^{N} \, \phi_{\vec{k}}(m_i, \vec{r}_i) \big)
-    &= -\frac{G}{4 \pi^2} \,
+    U = \ &\rho \, \big( \sum_{i=1}^{N} \, \phi_{\vec{k}}(m_i, \vec{r}_i) \big)
+    = -\frac{G}{4 \pi^2} \,
         \int \d^3 r \,
-        \mathcal{F}^{-1} \left[
-            \mathcal{F}[\rho] \ast \left(
-            k^{-2} \, \mathcal{F}[\rho] \right)
-        \right]~.
+        \mathcal{F}^{-1}\!\left\{
+            \mathcal{D}_{\vec{k}} \ast \left(
+            k^{-2} \, \mathcal{D}_{\vec{k}} \right)
+        \right\}~.
     \label{eq:EnergyDeepSets}
@@
 In principle this requires $\phi$ to be evaluated at every possible wavenumber $\vec{k}$.
-To avoid making this infinite-dimensional $\phi$, one can instead evaluate $\mathcal{F}[\rho]$ at a finite set of wavenumbers $\vec{k}$ and live with approximations to the integrals necessary for the convolution and inverse Fourier transform.
+To avoid making this infinite-dimensional $\phi$, one can instead evaluate $\mathcal{D}_{\vec{k}}$ at a finite set of wavenumbers $\vec{k}$ and live with approximations to the integrals necessary for the convolution and inverse Fourier transform.
 The approximation will improve as the size of $\phi$ grows.
@@
 
-We can even use these methods to find another formula for the area $\Delta$ of a triangle.
+Of course, these methods can be used to find another triangle area formula.
 \begin{align}
@@
     \quad
-    \mathcal{F}[\rho](\vec{k}) = \sum_{j=1}^{3} \phi_{\vec{k}}(\vec{r}_j)~,
+    \sum_{j=1}^{3} \phi_{\vec{k}}(\vec{r}_j) = \mathcal{D}_{\vec{k}}~,
+    \quad
+    \mathcal{P}(\vec{r}) = (2\pi)^d \, \mathcal{F}^{-1}\!\left\{
+    \abs{ \mathcal{D}_{\vec{k}} }^2 \right\}~,
     \nonumber\\
-    \mathcal{P}(\vec{r}) &= (2\pi)^d \, \mathcal{F}^{-1} \left[ 
-    \abs{ \mathcal{F}[\rho] }^2 \right]~,
+    P_m &= \int_{\mathbb{R}^d} \d^d r \, \mathcal{P}(\vec{r}) \, \frac{\abs{\vec{r}}^m}{2}
+    = a^m+b^m+c^m
+    ~,
     \quad
-    F_m = \int_{\mathbb{R}^d} \d^d r \, \mathcal{P}(\vec{r}) \, \frac{\abs{\vec{r}}^m}{2}~,
-    \nonumber\\
     \Delta
-    &= \frac{1}{4} \, \sqrt{F_2^2 - 2 \, F_4}~.
+    = \frac{1}{4} \, \sqrt{P_2^2 - 2 \, P_4}~.
+    \label{eq:Fresult}
 \end{align}
 A complete derivation is provided in the \nameref{sec:appendix}.
-Briefly, $\mathcal{P}$ is the Patterson function of X-ray crystallography \citep{Patterson1934} which has 6 delta function peaks corresponding to the three displacement vectors and their negatives. 
-Multiplying the Patterson function by a function of the magnitude $r \equiv \vec{r}$ and integrating over all space gives a sum of the values of the function evaluated at the three side lengths. 
-Choosing this function properly allows us to write the power sums of the side lengths $F_m = a^m+b^m+c^m$ in this form.
+Briefly, $\mathcal{P}$ is the Patterson function used in x-ray crystallography \citep{Patterson1934} which has six delta function peaks corresponding to the three displacement vectors and their negatives.
+Multiplying the Patterson function by a function of the magnitude $r \equiv \abs{\vec{r}}$ and integrating over all space gives a sum of the values of the function evaluated at the three side lengths.
+Choosing this function properly allows us to write the $m^{\rm th}$ power sum $P_m$ of the side lengths in this form.
 
-To verify this formula, note that we can write it in terms of the side lengths as
+To verify this formula \eqref{eq:Fresult}, note that we can write it in terms of the side lengths as
 \begin{equation}
@@
     \Delta = \frac{1}{4} \, \sqrt{
-        F_2^2
-        - 2 \, F_4
+        P_2^2
+        - 2 \, P_4
     }
@@
         - 2 \, \left( a^4 + b^4 + c^4 \right)
-    }~.
+    }~,
 \end{equation}
-This formula is equivalent to Heron's formula \eqref{eq:Heron} by expansion.
-Note that it is also in Deep Sets form as a function of the side lengths.
+which is equivalent to Heron's ancient formula \eqref{eq:Heron}.
+It is even in Deep Sets form as a function of the side lengths.
 
-\CH{I don't like this ending ... thinking about other ways to conclude}
+\secbreak
+In this \documentname, we were motivated by the question of how to convert $k$-point functions in cosmology, which involve loops over $k$-tuplets of points and na\"ively scale as $N^k$, into Deep Sets forms \eqref{eq:DeepSets}, which na\"ively scale as $N$, though perhaps with very large constant-runtime computational costs.
+These conversions must be possible, since all permutation-invariant functions can be written in Deep Sets form.
+The simplest permutation-invariant problem involving a loop over pairs of points is the computation of the area of a triangle, for which no traditional formula is in Deep Sets form.
+We deliver two new formulae, a polynomial \eqref{eq:result} and a Fourier expression \eqref{eq:Fresult}.
+These triangle formulae have no practical value, and we also do not solve any question of practical importance for cosmology.
 
-We leave the reader with the following questions. 
-It is trivial to take the form of \eqref{eq:result} and evaluate the sums for more than three vectors; does the result have any meaning?
-Is there a generalization of the triangle area \eqref{eq:result} and simplex volume \eqref{eq:simplex} results to a $d$-simplex?
-Our formulae \eqref{eq:result} and \eqref{eq:simplex} only used invariant factors of degree $2$, with exponents that are larger than $2$.
-Does this generalize to higher dimension simplexes, and can it be shown that higher order factors aren't needed?
-One can replace the sum in \eqref{eq:DeepSets} with any other permutation-invariant aggregator (e.g., max or product), though we do not know if there are any claims to universality in these cases.
-Does another area formula exist for these other aggregators?
-% We have discussed the general value of a map function $\phi$ which Fourier transforms the points;\footnote{%
-% \Hogg{The Deep Sets version of the Fourier transform is reminiscent of a positional encoding (CITE), similar to those used to preserve ordering information in (otherwise permutation invariant) transformers (CITE).}}
-% does a formula like \eqref{eq:EnergyDeepSets} exist for the area of a triangle?
+% We leave the reader with the following question. 
+% It is trivial to take the form of \eqref{eq:result} and evaluate the sums for more than three vectors; does the result have any meaning?
+% One can replace the sum in \eqref{eq:DeepSets} with any other permutation-invariant aggregator (e.g., max or product), though we do not know if there are any claims to universality in these cases.
+Do area formulae exist for these other aggregators?
 
@@
 This project would not have been possible without originating inspiration and discussions with Soledad Villar (JHU).
+The authors give special thanks to Ben Blum-Smith (JHU) and Kendrick Smith (Perimeter) for insightful and important input.
 The authors thank
-  Ben Blum-Smith (JHU),
+  % Ben Blum-Smith (JHU),
   Jonathan Goodman (NYU),
@@
   Mike O'Neil (NYU),
+  Simon Park (Princeton),
   Oliver Philcox (Columbia),
+  % Kendrick Smith (Perimeter Institute),
   Ben Wandelt (JHU),
@@
 CH is supported by the National Science Foundation Graduate Research Fellowship under Grant Number DGE-2234660.
-The authors also note that, while we believe the formula is (barely) novel, it is difficult to be sure, given the challenge of conducting a complete literature review for this ancient topic.\par}
+The authors also note that, while we believe the formulae are (barely) novel, it is difficult to be sure, given the challenge of conducting a complete literature review for this ancient topic.\par}
 
@@
 
-\paragraph{Proof that area sums in quadrature} We prove equation \eqref{eq:quadrature}, that the area of a triangle in $d$ dimensions is the sum in quadrature of the areas of the triangles made by all orthogonal projections to two axes.
+\paragraph{Proof that area sums in quadrature} We prove equation \eqref{eq:quadrature}, that the area of a triangle in $d$ dimensions is the sum in quadrature of the areas of the triangles made by all orthogonal projections onto pairs of axes.
 
@@
 
-\paragraph{Derivation of Deep Sets energy} Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
-The energy density $\mathcal{E} = d U / d^3 x$ is given by
+\paragraph{Volume of $d$-simplex} Consider a $d$-simplex in $d$-dimensional space. Let its vertices be labeled $\{ \vec{r}_1, \dots, \vec{r}_{d+1} \}$. The volume of the $d$-simplex is known to be
 \begin{equation}
-    \mathcal{E} = \frac{\rho \, \Phi,}{2}~,
+    V = \frac{1}{d!} \, \abs{\det \mathsf{A}},
+    \quad \mathsf{A} \equiv \left[ \begin{array}{cc}
+        \vec{r}_1 & 1 \\
+        \vdots & \vdots \\
+        \vec{r}_{d+1} & 1 \\
+    \end{array} \right]~,
 \end{equation}
-where $\rho$ is the mass density and $\Phi$ is the gravitational potential \citep{LandauLifshitzFields}.
+where $\mathsf{A}$ is a $(d+1) \times (d+1)$ matrix whose rows are the vertices, appended by a single 1.
+
+Squaring both sides, we find
+\begin{equation}
+    V^2
+    = \frac{1}{(d!)^2} \, (\det \mathsf{A})^2
+    = \frac{1}{(d!)^2} \, \det \mathsf{A}^\top \mathsf{A},
+\end{equation}
+and $\mathsf{A}^\top \mathsf{A}$ matches $\mathsf{M}$ as defined in equation \eqref{eq:dsimplex}.
+
+\paragraph{Deep Sets energy} Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
+The energy density $\mathcal{E} = \d U / \d^3 x$ is given by
+\begin{equation}
+    \mathcal{E} = \frac{\varrho \, \Phi,}{2}~,
+\end{equation}
+where $\varrho$ is the mass density and $\Phi$ is the gravitational potential \citep{LandauLifshitzFields}.
 These are related by the Poisson equation
 \begin{equation}
-    \nabla^2 \Phi = 4 \pi G \, \rho~,
+    \nabla^2 \Phi = 4 \pi G \, \varrho~,
 \end{equation}
@@
 \begin{equation}
-    -k^2 \, \mathcal{F}[\Phi] = 4 \pi G \, \mathcal{F}[\rho]~,
+    -k^2 \, \mathcal{F}\{\Phi\} = 4 \pi G \, \mathcal{F}\{\varrho\}~,
 \end{equation}
@@
 \begin{align}
-    \mathcal{F}[\mathcal{E}] 
-    &= \frac{1}{2} \mathcal{F}[\rho \, \Phi] \nonumber\\
-    &= \frac{1}{2} \, \frac{\mathcal{F}[\rho] \ast \mathcal{F}[\Phi]}{(2\pi)^3} \nonumber\\
-    &= \frac{1}{2 \, (2\pi)^3} \, \mathcal{F}[\rho] \ast \left(
-        -4 \pi G \, k^{-2} \, \mathcal{F}[\rho]
-    \right)~.
+    \mathcal{F}\{\mathcal{E}\}
+    &= \frac{1}{2} \, \mathcal{F}\{\varrho \, \Phi\} \nonumber\\
+    &= \frac{1}{2 \, (2\pi)^3} \, \mathcal{F}\{\varrho\} \ast \mathcal{F}\{\Phi\} \nonumber\\
+    &= -\frac{G}{4\pi^2} \, \mathcal{F}\{\varrho\} \ast \left( k^{-2} \, \mathcal{F}\{\varrho\} \right)~.
     \label{eq:ConvolutionThm}
@@
 \begin{equation}
-    \mathcal{F}[\rho](\vec{k})
+    \mathcal{F}\{\varrho\}(\vec{k})
     = \int \d^3 r
@@
 where we sum over $j$ and use $i=\sqrt{-1}$.
-Note that the value of the Fourier transform at a wavenumber $\vec{k}$ is of the required form $\sum_i \phi(m_i, \vec{r}_i)$ for Deep Sets features.
-Hence, the total energy is given in Deep Sets form by
-\begin{align}
-    \phi_{\vec{k}}(m_j, \vec{r}_j)
-    &=m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}
-    ~,
-    \quad
-    \sum_{i=1}^{N} \phi_{\vec{k}}(m_i, \vec{r}_i)
-    = \mathcal{F}[\rho](\vec{k})~,
-    \nonumber\\
-    U \, \big( \sum_{i=1}^{N} \, \phi_{\vec{k}}(m_i, \vec{r}_i) \big)
-    &= -\frac{G}{4 \pi^2} \,
+Thus, the total energy is given in terms of the Fourier transform of the density field by
+\begin{equation}
+    U 
+    = -\frac{G}{4 \pi^2} \,
         \int \d^3 r \,
-        \mathcal{F}^{-1} \left[
-            \mathcal{F}[\rho] \ast \left(
-            k^{-2} \, \mathcal{F}[\rho] \right)
-        \right]~.
-\end{align}
+        \mathcal{F}^{-1} \! \left\{
+            \mathcal{F}\{\varrho\} \ast \left(
+            k^{-2} \, \mathcal{F}\{\varrho\} \right)
+        \right\}~.
+\end{equation}
 
@@
 \begin{equation}
-    \mathcal{F}[\rho \, \Phi] 
-    = (2\pi)^{-3} \, \mathcal{F}[\rho] \ast \mathcal{F}[\Phi] ~.
+    \mathcal{F}\{\varrho \, \Phi\}
+    = (2\pi)^{-3} \, \mathcal{F}\{\varrho\} \ast \mathcal{F}\{\Phi\} ~.
 \end{equation}
@@
     \tilde{f}(\vec{k})
-    = \mathcal{F}[f(\vec{x})]
+    = \mathcal{F}\{f(\vec{x})\}
     &= \int \d^3 x
@@
     f(\vec{x})
-    = \mathcal{F}^{-1}[\tilde{f}(\vec{k})]
+    = \mathcal{F}^{-1}\{\tilde{f}(\vec{k})\}
     &= \int \frac{\d^3 k}{(2\pi)^3}
         \, e^{+i \, \vec{k} \cdot \vec{x}}
-        \, f(\vec{x})~.
+        \, \tilde{f}(\vec{k})~.
 \end{align}
-Consider $\mathcal{F}^{-1}[\tilde{\rho} \ast \tilde{\Phi}](\vec{x})$. 
+Consider $\mathcal{F}^{-1}\{\tilde{\varrho} \ast \tilde{\Phi}\}$. 
 \begin{align}
-    \mathcal{F}^{-1}[\tilde{\rho} \ast \tilde{\Phi}](\vec{x})
+    \mathcal{F}^{-1}\{\tilde{\varrho} \ast \tilde{\Phi}\}(\vec{x})
     &= \int_{\mathbb{R}^3} \frac{\d^3 k}{(2\pi)^3}
         \, e^{i \, \vec{k} \cdot \vec{x}} 
-        \, \big( \tilde{\rho} \ast \tilde{\Phi} \big) (\vec{k}) 
+        \, \big( \tilde{\varrho} \ast \tilde{\Phi} \big) (\vec{k}) 
         \nonumber\\
@@
         \, \int_{\mathbb{R}^3} \d^3 \vec{k}' 
-        \, \tilde{\rho}(\vec{k}') 
+        \, \tilde{\varrho}(\vec{k}') 
         \, \tilde{\Phi}(\vec{k} - \vec{k}')
@@
     &= \int_{\mathbb{R}^3} \d^3 \vec{k}'
-        \, \tilde{\rho}(\vec{k}') 
+        \, \tilde{\varrho}(\vec{k}') 
         \, \int_{\mathbb{R}^3} \frac{\d^3 k}{(2\pi)^3}
@@
     &= \int_{\mathbb{R}^3} \d^3 \vec{k}'
-        \, \tilde{\rho}(\vec{k}') 
+        \, \tilde{\varrho}(\vec{k}') 
         \, e^{i \, \vec{k}' \cdot \vec{x}}
@@
     &= (2\pi)^3
-        \, \rho(\vec{x}) 
+        \, \varrho(\vec{x}) 
         \, \Phi(\vec{x})~.
@@
 \begin{equation}
-    \rho(\vec{r})
+    \varrho(\vec{r})
     = \sum_i \, \delta(\vec{r} - \vec{r}_i)~.
@@
 \begin{equation}
-    \tilde{\rho}(\vec{k})
+    \tilde{\varrho}(\vec{k})
     =
@@
 \begin{equation}
-    \abs{\tilde{\rho}(\vec{k})}^2
+    \abs{\tilde{\varrho}(\vec{k})}^2
     =
@@
     \mathcal{P}(\vec{r})
-    &= (2\pi)^d \, \mathcal{F}^{-1} \left[ \abs{\tilde{\rho}}^2 \right]
+    &= (2\pi)^d \, \mathcal{F}^{-1} \! \left\{ \abs{\tilde{\varrho}}^2 \right\}
     \nonumber\\ 
@@
 \end{align}
-In particular, we can define the $k^{\rm th}$ power sum of the side lengths
+In particular, we can define the $m^{\rm th}$ power sum of the side lengths
 \begin{equation}
-    F_k
-    = \int_{\mathbb{R}^d} \d^d r \, \mathcal{P}(\vec{r}) \, \frac{\abs{\vec{r}}^k}{2}
-    = a^k + b^k + c^k~,
+    P_m
+    = \int_{\mathbb{R}^d} \d^d r \, \mathcal{P}(\vec{r}) \, \frac{\abs{\vec{r}}^m}{2}
+    = a^m + b^m + c^m~,
 \end{equation}
@@
     = \frac{1}{4} \, \sqrt{\left(a^2+b^2+c^2\right)^2 - 2\, \left(a^4+b^4+c^4\right)}
-    = \frac{1}{4} \, \sqrt{F_2^2 - 2\, F_4}~.
+    = \frac{1}{4} \, \sqrt{P_2^2 - 2\, P_4}~.
 \end{equation}

### 2025-03-27 paper/main.tex
getting very close now

@@
 \renewcommand{\paragraph}[1]{\par\addvspace{1em}\noindent\textsl{#1}~---}
-\newcommand{\secbreak}{\bigskip{\centering\footnotesize $\triangle~~~\triangle~~~\triangle$\par}\bigskip\noindent}
+\newcommand{\secbreak}{\bigskip{\centering\footnotesize%
+\rotatebox[origin=c]{55}{$\triangle$}~~~%
+\rotatebox[origin=c]{35}{$\triangle$}~~~%
+\rotatebox[origin=c]{15}{$\triangle$}\par}\bigskip\noindent}
 \titleformat*{\subsection}{\normalfont\normalsize\slshape}
@@
 Formula~\eqref{eq:school} is the usual introduction to the subject, traceable back to \citet{Euclid300BC}.
-Formula~\eqref{eq:sine} makes use of the fact that the height can be found from $\sin(C) = (\text{height}) / a$, commonly seen in introductory geometry courses.
+Formula~\eqref{eq:sine} makes use of the fact that if the base is $b$ then the height is $a\,\sin(C)$.
 Formula~\eqref{eq:cross} generalizes to triangles in higher dimensional spaces, using the fact that the cross or wedge product has a magnitude equal to the area of the parallelogram spanned by $\vec{a}$ and $\vec{b}$.
@@
 It has long been known that any permutation-invariant (or symmetric) polynomial of scalar data $f(\{r_i\})$ can be written as a polynomial of the power sums $\sum_i r_i^m$ \citep{Waring1782}.
-Further, symmetric polynomials of vector data $\{ \vec{r}_i \}$ are decomposable into polarizations of the scalar symmetric polynomials \citep{Weyl1939}, which look like sums of the form $\sum_i \, (\vec{r}_{i})_{j}^{m_j} \cdots (\vec{r}_{i})_{k}^{m_k}$, where $(\vec{r}_i)_j$ is the $j^{\rm th}$ component of $\vec{r}_i$.
-Together, these imply that a permutation-invariant polynomial of vector data is a polynomial in the polarizations of the power sums.
-This result is \emph{universal}, in the sense that it is true for all symmetric polynomials.
+Further, symmetric polynomials of vector data $\{ \vec{r}_i \}$ are decomposable into polarizations of the scalar symmetric polynomials \citep{Weyl1939}.
+Together, these imply that any permutation-invariant polynomial of vector data can be written as a polynomial in the polarizations of the power sums, which look like sums of the form $\sum_i \, (\vec{r}_{i})_{j}^{m_j} \cdots (\vec{r}_{i})_{k}^{m_k}$, where $(\vec{r}_i)_j$ is the $j^{\rm th}$ component of $\vec{r}_i$.
+% This result is \emph{universal}, in the sense that it is true for all symmetric polynomials.
 It also matches onto the Deep Sets form, where the polarizations of the power sums are the components of the vector $\sum_i \phi(\vec{r}_i)$, which are combined in a polynomial $\rho$.
@@
 which is exactly in the form of $\sum_i \phi(\vec{r}_i)$.
-Hence, any function of the Fourier transform $\mathcal{D}$ of the density field is technically in Deep Sets form, and has permutation symmetry guaranteed.
+Hence, any function of the Fourier transform $\mathcal{D}$ of the density field is in Deep Sets form and has permutation symmetry guaranteed.
 This is physically useful also because the Fourier basis $\exp (i\, \vec{k}\cdot\vec{x})$ is the eigenbasis of the translation operator, so this choice of the map $\phi$ naturally encodes translation symmetry.\footnote{%
-%
-Interestingly, a transformation like $\phi(r) = \exp(i\,k\,r)$ is used widely in natural language processing.
+    Interestingly, a transformation like $\phi(r) = \exp(i\,k\,r)$ is often used in natural language processing.
     Most large language models, such as GPT-3 \citep{Brown2020GPT3}, are transformer models \citep{Vaswani2017Attention}, which are fundamentally permutation invariant.
@@
     One of the most popular choices for this encoding is the Rotary Positional Encoding \citep[RoPE;][]{Su2024Roformer}, which takes the word embedding vector $\vec{x} \in \mathbb{R}^{d}$ for the token in position $m$ and multiplies it by $e^{i \, m \, \theta}$ for $d/2$ preset constant values $\theta \in \mathbb{R}$.
-While this actually \emph{breaks} permutation invariance rather than enforcing it (since the encoding explicitly involves the index $m$), it underscores the broad applications of these transformations.
-}
+    While this actually \emph{breaks} permutation invariance rather than enforcing it (since the encoding explicitly involves the index $m$), it underscores the broad applicability of these kinds of transformations.}
 
@@
 The authors give special thanks to Ben Blum-Smith (JHU) and Kendrick Smith (Perimeter) for insightful and important input.
-The authors thank
+The authors also thank
   % Ben Blum-Smith (JHU),
@@
   Oliver Philcox (Columbia),
-  % Kendrick Smith (Perimeter Institute),
+  % Kendrick Smith (Perimeter),
   Ben Wandelt (JHU),
   Kaze Wong (JHU),
-and the Blanton--Hogg group meeting at NYU for valuable discussions.
+and the Blanton--Hogg group meeting at NYU for valuable comments.
 The Flatiron Institute is a division of the Simons Foundation.
@@
 
-\paragraph{Proof that area sums in quadrature} We prove equation \eqref{eq:quadrature}, that the area of a triangle in $d$ dimensions is the sum in quadrature of the areas of the triangles made by all orthogonal projections onto pairs of axes.
+\paragraph{Proof that the area sums in quadrature} We prove equation \eqref{eq:quadrature}, that the area of a triangle in $d$ dimensions is the sum in quadrature of the areas of the triangles made by all pairs of mutually orthogonal projections of the points.
 
 From the three position vectors $\vec{r}_1, \vec{r}_2, \vec{r}_3 \in \mathbb{R}^d$ (for $d \geq 2$), we consider the two displacement vectors $\vec{a} \equiv \vec{r}_2 - \vec{r}_1$ and $\vec{b} \equiv \vec{r}_3 - \vec{r}_1$.
-Given these, we can write the two-dimensional triangle sub-area $\Delta_{jk}$ by first defining the projected two-dimensional displacement vectors $\vec{a}_{jk} = ( a_j, a_k )$, $\vec{b}_{jk} = ( b_j, b_k )$.
+Given these, we can write the two-dimensional triangle sub-area $\Delta_{jk}$ by first defining the projected two-dimensional displacement vectors $\vec{a}_{jk} = \big[ a_j, a_k \big]$, $\vec{b}_{jk} = \big[ b_j, b_k \big]$.
 Then,
@@
 
-\paragraph{Volume of $d$-simplex} Consider a $d$-simplex in $d$-dimensional space. Let its vertices be labeled $\{ \vec{r}_1, \dots, \vec{r}_{d+1} \}$. The volume of the $d$-simplex is known to be
+\paragraph{Volume of the $d$-simplex} Consider a $d$-simplex in $d$-dimensional space. Let its vertices be labeled $\{ \vec{r}_1, \dots, \vec{r}_{d+1} \}$. The volume of the $d$-simplex is known to be
 \begin{equation}
@@
 
-\paragraph{Deep Sets energy} Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
+\paragraph{Deep Sets form for the energy} Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
 The energy density $\mathcal{E} = \d U / \d^3 x$ is given by
@@
 
-\paragraph{Fourier triangle area}
+\paragraph{Fourier-space triangle area formula}
 Suppose our triangle is specified by three vertices $\vec{r}_1,\vec{r}_2,\vec{r}_3 \in \mathbb{R}^d$. We represent this distribution by the field

### 2025-03-27 paper/main.tex
Tiny hogg adjustments

@@
 This result---often known by the name of the deep learning architecture derived from it, \emph{Deep Sets} (\citealt{Zaheer+17deepsets}; see Theorem~2)---is not at all obvious.
-For example, even though Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, it is not actually in the form of \eqref{eq:DeepSets}. 
+For example, even though Heron's formula \eqref{eq:Heron} is explicitly permutation-invariant, it is not actually in the form of \eqref{eq:DeepSets}.\footnote{But see equation \eqref{eq:HeronDeepSets} below.}
 Further, despite the triangle area being permutation-invariant, there is no known (prior to this work) formula for the area of a triangle that is explicitly in Deep Sets form.
@@
 While this result was recently popularized as Deep Sets, it follows from the classical theory of invariant polynomials.
-It has long been known that any permutation-invariant (or symmetric) polynomial of scalar data $f(\{r_i\})$ can be written as a polynomial of the power sums $\sum_i r_i^m$ \citep{Waring1782}.
+Any permutation-invariant (or symmetric) polynomial of scalar data $f(\{r_i\})$ can be written as a polynomial of the power sums $\sum_i r_i^m$ \citep{Waring1782}.
 Further, symmetric polynomials of vector data $\{ \vec{r}_i \}$ are decomposable into polarizations of the scalar symmetric polynomials \citep{Weyl1939}.
@@
 What resolves this discrepancy?
-The resolution must be either that $\phi$ has a dimensionality that grows with $N$, or else that the complexity of $\rho$ grows with $N$.
+The resolution must be either that $\phi$ has a dimensionality that effectively grows with $N$, or else that the complexity of $\rho$ effectively grows with $N$.
 Still, these time complexities may be better than the $N^k$ runtime of na\"ive implementations.
@@
 It can be executed independently and asynchronously on each data point.
-The sums are then the ``reduce'', which can be performed hierarchically across a data center such that the runtime of the sums scales only logarithmically with the number of points.
+The sums are then the ``reduce,'' which can be performed hierarchically across a data center such that the runtime of the sums scales only logarithmically with the number of points.
 (Granted, a triangle will never have more than three points!)
@@
 Notice that the runtime is formally linear in the number of points, since $\phi$ and $\rho$ can be considered to have constant time complexity.
-However, because the dimensionality of $\phi$ is larger than the number of points, evaluating the triangle area in this way will not necessarily be faster than a loop over pairs of points.
+However, because the dimensionality of $\phi$ is larger than the number of points, evaluating the triangle area in this way will probably not ever be faster than a loop over pairs of points.
 
@@
 While the triangle area, by virtue of being a sum over pairs of points, is in a form similar to some physics problems, the polynomial methods we employ are not so relevant for those problems.
-However, in physics it is common to encode a set of points by defining the density field to be a sum of delta functions at the points.
+In physics contexts it is common to encode a set of points by defining a density field that is a sum of delta functions located at the points.
 Notice that the Fourier transform of this field at a wavevector $\vec{k}$ is
@@
 we can find a Deep Sets form for the gravitational potential energy of a system of point masses.
-Explicitly, consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
+To be explicit: Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
 The total energy is given in Deep Sets form by
@@
 
-Of course, these methods can be used to find another triangle area formula.
+Of course, these Fourier methods can be used to find another triangle area formula.
 \begin{align}

### 2025-03-29 paper/main.tex
as submitted to arXiv for April 1

@@
 % Copyright 2025 the authors.
-
-% to-do items
-% -----------
-% - [x] CH will mention positional encodings.
-% - [x] CH will consider changing the outer function rho() to sigma().
-% - [x] CH should define \d as \mathrm{d} and use it for differentials.
 
@@
 \usepackage[T1]{fontenc}
-\usepackage{amsmath,amsfonts,lmodern,mathtools,nameref,setspace,tikz,titlesec}
+\usepackage{amsmath,amsfonts,lmodern,mathtools,nameref,natbib,setspace,tikz,titlesec}
 \usetikzlibrary{angles,calc,quotes}
@@
 % bibliography issues
-\usepackage{aas_macros,natbib}
 \bibliographystyle{aasjournal}
@@
 It is invariant to translation, rotation, and reflection, as well as to the labeling of sides or corners.
-However, these symmetries are not always obvious from a given formula; for example, of those listed above, only the polynomial \eqref{eq:polynomial} and Heron's formula \eqref{eq:Heron} are explicitly permutation-invariant.
+However, these symmetries are not always obvious from a given formula; for example, of those listed above, only the absolute-value polynomial \eqref{eq:polynomial} and Heron's formula \eqref{eq:Heron} are explicitly permutation-invariant.
 
@@
 Together, these imply that any permutation-invariant polynomial of vector data can be written as a polynomial in the polarizations of the power sums, which look like sums of the form $\sum_i \, (\vec{r}_{i})_{j}^{m_j} \cdots (\vec{r}_{i})_{k}^{m_k}$, where $(\vec{r}_i)_j$ is the $j^{\rm th}$ component of $\vec{r}_i$.
-% This result is \emph{universal}, in the sense that it is true for all symmetric polynomials.
 It also matches onto the Deep Sets form, where the polarizations of the power sums are the components of the vector $\sum_i \phi(\vec{r}_i)$, which are combined in a polynomial $\rho$.
@@
 \end{equation}
-This is explicitly permutation invariant since it involves a sum without repeats over all pairs of particles, but it is not in Deep Sets form \eqref{eq:DeepSets} because it involves a double sum (with $N^2$ computational scaling) of quantities computed from pairs of particles.
+This is explicitly permutation invariant since it involves a sum without repeats over all pairs of particles, but it is not in Deep Sets form \eqref{eq:DeepSets} because it involves a double sum---with $\mathcal{O}(N^2)$ computational scaling---of quantities computed from pairs of particles.
 
@@
 are all permutation-invariant.
-Calculating these statistics na\"ively involves loops over all $k$-tuplets of points, which is not only inconsistent with Deep Sets form, but also has a runtime that scales as $N^k$, where $N$ the number of points in the set.
+Calculating these statistics na\"ively involves loops over all $k$-tuplets of points, which is not only inconsistent with Deep Sets form, but also has a runtime complexity of $\mathcal{O}(N^k)$, where $N$ is the number of points in the set.
 
 Because all these (computationally expensive) statistics are permutation-invariant, it must be possible to write them in Deep Sets form.
-At the same time, Deep Sets form \eqref{eq:DeepSets} looks like it is computable in linear time.
+At the same time, Deep Sets form \eqref{eq:DeepSets} looks like it is computable in linear $\mathcal{O}(N)$ time.
 What resolves this discrepancy?
 The resolution must be either that $\phi$ has a dimensionality that effectively grows with $N$, or else that the complexity of $\rho$ effectively grows with $N$.
-Still, these time complexities may be better than the $N^k$ runtime of na\"ive implementations.
+Still, these time complexities may be better than the $\mathcal{O}(N^k)$ runtime of na\"ive implementations.
 
@@
 It can be executed independently and asynchronously on each data point.
-The sums are then the ``reduce,'' which can be performed hierarchically across a data center such that the runtime of the sums scales only logarithmically with the number of points.
+The sums are then the ``reduce,'' which can be performed hierarchically across a data center such that the wall-clock runtime of the sums scales only logarithmically $\mathcal{O}(\log N)$ with the number of points.
 (Granted, a triangle will never have more than three points!)
@@
 angles, as in \eqref{eq:sine};
-or point displacements, as in \eqref{eq:cross};
+point displacements, as in \eqref{eq:cross};
 or as a sum over $x_i \, y_j - x_j \, y_i$, as in \eqref{eq:polynomial}.
-% which can be written as a sum over pairs of points as
-% \begin{equation}
-%     \label{eq:pairwiseSums}
-%     \Delta = \frac{1}{2} \, \bigg|\,
-%     \sum_{i=1}^{3} \sum_{j=i+1}^{3}
-%     x_i \, y_j - x_j \, y_i
-%     \,\bigg|~,
-% \end{equation}
-% very similar in form to \eqref{eq:EnergyNaive}.
 
@@
 \end{gather}
-Notice that the runtime is formally linear in the number of points, since $\phi$ and $\rho$ can be considered to have constant time complexity.
+Notice that the runtime is formally linear in the number of points, since $\phi$ and $\rho$ can be considered to have constant $\mathcal{O}(1)$ time complexity.
 However, because the dimensionality of $\phi$ is larger than the number of points, evaluating the triangle area in this way will probably not ever be faster than a loop over pairs of points.
@@
 Recall that a triangle with vertices $\{ \vec{r}_1, \vec{r}_2, \vec{r}_3 \}$ has one-half the area of the parallelogram spanned by the displacement vectors $\{ \vec{r}_2 - \vec{r}_1, \vec{r}_3 - \vec{r}_1 \}$.
-So our result \eqref{eq:result} is actually an integer polynomial for the squared area of this parallelogram.
+So our result \eqref{eq:result} gives an integer polynomial for the squared area of this parallelogram.
 Using this fact, we have written code that performs a symbolic regression to recover our result.\footnote{
@@
 The code performs least-squares regression to find a fourth-order homogeneous polynomial of the multi-power sums $\sum_i \, x_i^m \, y_i^n$, and it requires that the result gives the squared areas to machine precision after rounding the coefficients to the nearest integers.
-Using and extending the code, we also find several generalizations to our formula.
-
-% Note that a triangle defined by three points has one-half the area of the parallelogram spanned by the vectors $\vec{a} = \vec{r}_2 - \vec{r}_1$ and $\vec{b} = \vec{r}_3 - \vec{r}_1$.
-% Multiplying both sides by 4 reveals that \eqref{eq:result} is an integer polynomial of permutation-invariant factors for the squared area of the parallelogram.
-% Using this fact, we have written code to perform a symbolic regression that recovers our result.\footnote{
-%     The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
-% }
-
-% In this symbolic regression, we consider a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
-% All the possible monomials are made of permutation-invariant factors of the form 
-% \begin{equation}
-% F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n~;
-% \end{equation}
-% each such a factor has degree $m + n$.
-% The invariant monomials of degree $d$ are then products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
-% We generate all factors $F_{mn}$ with degree $1 \leq m + n \leq 4$, then form all products of these factors with total degree four.
-% However, these monomials are not all linearly-independent;
-% the five monomials made of a single degree-four factor ($F_{40}$, $F_{31}$, $F_{22}$, $F_{13}$, $F_{04}$) can be written as a linear combination of the other monomials and are thus removed.
-% This leaves 28 linearly-independent monomials.
-
-% We fit a linear combination of the monomials to a data set of generated triples of points, drawn from a two-dimensional Gaussian whose mean was also drawn from a Gaussian in a two-dimensional space.
-% Each triple defines a parallelogram whose squared area we use as a label for regression.
-% We then perform least-squares regression to fit a linear combination of the invariant monomials.
-% We require that the labels are reproduced to machine precision after rounding the coefficients to the nearest integers.
-% The result exactly matches equation \eqref{eq:result}.
+Using and extending the code, we also find several equivalent but more complicated formulae, which are given in the \nameref{sec:appendix}.
+All the additional formulae require a higher-dimensional $\phi$ with higher-order terms like $x_i^2 \, y_i^2$ which are combined in ways that reduce to our result \eqref{eq:result}, indicating that \eqref{eq:result} may be the ``simplest'' such formula.
 
@@
 where $r_{ij}$ is the $j^{\rm th}$ component of $\vec{r}_i$.\footnote{%
-    This generalization was pointed out by Kendrick Smith (Perimeter Institute).}
+    This generalization was pointed out by Kendrick Smith (Perimeter).}
 A proof is deferred to the \nameref{sec:appendix}.
@@
 However, the dimensionality of $\phi$ is $(d+1)(d+2)/2$, the number of unique components in $\mathsf{M}$, so there are no real performance improvements.
-As with the relationship between triangles and parallelograms, the volume of the $d$-parallelotope spanned by the displacement vectors $\{ \vec{r}_2 - \vec{r}_1, \dots, \vec{r}_{d+1} - \vec{r}_1 \}$ is $d!$ the volume of a $d$-simplex; thus \eqref{eq:dsimplex} is an integer polynomial for the squared volume of this $d$-parallelotope.
-
-% Our method also extends to the volume of a simplex. In three dimensions, four vectors $\{ \vec{r}_1, \vec{r}_2, \vec{r}_3, \vec{r}_4 \}$ determine the vertices of a simplex. 
-% Like the area of a triangle is one-half the area of a related parallelogram, the volume of a simplex is one-sixth the volume of the parallelepiped spanned by the displacement vectors $\{ \vec{r}_2 - \vec{r}_1, \vec{r}_3 - \vec{r}_1, \vec{r}_4 - \vec{r}_1 \}$.
-% Expecting an integer solution once again, we modify the symbolic regression to generate 4-tuplets of points, using the squared volumes of the parallelepipeds they produce as labels.
-% The invariant factors then have three indices, $F_{\ell m n} = \sum_i x_i^\ell \, y_i^m \, z_i^n$, and we form all invariant monomials of degree six.
-% The result, verifiable by computational algebra, is that the volume $V$ of the simplex is given by
-% \begin{align}
-% (6 V)^2 = &
-% +8 \, F_{110} \, F_{101} \, F_{011}
-% +4 \, F_{200} \, F_{020} \, F_{002}
-% \nonumber \\ &
-% -4 \, F_{200} \, F_{011}^2
-% -4 \, F_{020} \, F_{101}^2
-% -4 \, F_{002} \, F_{110}^2
-% \nonumber \\ &
-% +2 \, F_{200} \, F_{011} \, F_{010} \, F_{001}
-% +2 \, F_{020} \, F_{101} \, F_{100} \, F_{001}
-% +2 \, F_{002} \, F_{110} \, F_{100} \, F_{010}
-% \nonumber \\ &
-% -2 \, F_{110} \, F_{101} \, F_{010} \, F_{001}
-% -2 \, F_{110} \, F_{011} \, F_{100} \, F_{001}
-% -2 \, F_{101} \, F_{011} \, F_{100} \, F_{010}
-% \nonumber \\ &
-% +1 \, F_{110}^2 \, F_{001}^2
-% +1 \, F_{101}^2 \, F_{010}^2
-% +1 \, F_{011}^2 \, F_{100}^2
-% \nonumber \\ &
-% -1 \, F_{200} \, F_{020} \, F_{001}^2
-% -1 \, F_{200} \, F_{002} \, F_{010}^2
-% -1 \, F_{020} \, F_{002} \, F_{100}^2~.
-% \label{eq:simplex}
-% \end{align}
-% There are many similarities between this and our result for the triangle area \eqref{eq:result}, which we believe may indicate a way to generalize this result to the hypervolume of a $d$-simplex.
+As with the relationship between triangles and parallelograms, the volume of the $d$-parallelotope spanned by the displacement vectors $\{ \vec{r}_2 - \vec{r}_1, \dots, \vec{r}_{d+1} - \vec{r}_1 \}$ is $d!$ times the volume of a $d$-simplex; thus \eqref{eq:dsimplex} is an integer polynomial for the squared volume of this $d$-parallelotope.
 
@@
     Most large language models, such as GPT-3 \citep{Brown2020GPT3}, are transformer models \citep{Vaswani2017Attention}, which are fundamentally permutation invariant.
-    In order to use the positional information of tokens (often words), the tokens are augmented by a ``positional encoding.''
-    One of the most popular choices for this encoding is the Rotary Positional Encoding \citep[RoPE;][]{Su2024Roformer}, which takes the word embedding vector $\vec{x} \in \mathbb{R}^{d}$ for the token in position $m$ and multiplies it by $e^{i \, m \, \theta}$ for $d/2$ preset constant values $\theta \in \mathbb{R}$.
+    In order to use the positional information of tokens (which are like words), the tokens are augmented by a ``positional encoding.''
+    One of the most popular choices for this encoding is the Rotary Positional Encoding \citep[RoPE;][]{Su2024Roformer}, which takes the word embedding vector for the token in position $m$ and multiplies its components by $e^{i \, m \, \theta}$ for preset constant values $\theta \in \mathbb{R}$.
     While this actually \emph{breaks} permutation invariance rather than enforcing it (since the encoding explicitly involves the index $m$), it underscores the broad applicability of these kinds of transformations.}
@@
 To be explicit: Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
-The total energy is given in Deep Sets form by
+The total energy is given as a function of $\mathcal{D}$, and thus in Deep Sets form, by
 \begin{align}
@@
 A derivation is given in the \nameref{sec:appendix}.
-In principle this requires $\phi$ to be evaluated at every possible wavenumber $\vec{k}$.
-To avoid making this infinite-dimensional $\phi$, one can instead evaluate $\mathcal{D}_{\vec{k}}$ at a finite set of wavenumbers $\vec{k}$ and live with approximations to the integrals necessary for the convolution and inverse Fourier transform.
+In principle this requires $\phi$ to be evaluated at every possible wavevector $\vec{k}$.
+To avoid making this infinite-dimensional $\phi$, one can instead evaluate $\mathcal{D}_{\vec{k}}$ at a finite set of wavevectors $\vec{k}$ and live with approximations to the integrals necessary for the convolution and inverse Fourier transform.
 The approximation will improve as the size of $\phi$ grows.
-Approaches akin to this Fourier approach can be used \citep[and are used; e.g.,][]{Portillo+2018,Philcox+2022ENCORE} to estimate or approximate $k$-point statistics in cosmology, which was our motivating question.
+
+Approaches akin to this Fourier approach can be used \citep[and are used; e.g.,][]{SlepianEisenstein2016,Portillo+2018,Philcox+2022ENCORE} to estimate or approximate $k$-point statistics in cosmology, which was our motivating question.
+However, these approaches are not in Deep Sets form, as they generally put the points on a grid and use the Fast Fourier Transform to compute a multipole expansion in the spherical harmonics.
+This has linearithmic $\mathcal{O}(N_g \log N_g)$ runtime complexity in the number of grid cells $N_g$, since the Fast Fourier Transform is not in map--reduce form.
+Perhaps there are gains to be made by considering the problem explicitly in terms of $\mathcal{D}$ for a small set of wavevectors, which could be computed in map--reduce form.
 
@@
 A complete derivation is provided in the \nameref{sec:appendix}.
-Briefly, $\mathcal{P}$ is the Patterson function used in x-ray crystallography \citep{Patterson1934} which has six delta function peaks corresponding to the three displacement vectors and their negatives.
-Multiplying the Patterson function by a function of the magnitude $r \equiv \abs{\vec{r}}$ and integrating over all space gives a sum of the values of the function evaluated at the three side lengths.
+Briefly, $\mathcal{P}$ is the Patterson function\footnote{%
+    It is also the two-point correlation function $\xi(\vec{r})$ without isotropy, but our usage here is more directly inspired by the crystallography method.}
+used in x-ray crystallography \citep{Patterson1934} which has six delta function peaks corresponding to the three displacement vectors and their negatives.
+Multiplying the Patterson function by a function of the magnitude $\abs{\vec{r}}$ and integrating over all space gives a sum of the values of the function evaluated at the three side lengths. 
 Choosing this function properly allows us to write the $m^{\rm th}$ power sum $P_m$ of the side lengths in this form.
@@
 
-% We leave the reader with the following question. 
-% It is trivial to take the form of \eqref{eq:result} and evaluate the sums for more than three vectors; does the result have any meaning?
-% One can replace the sum in \eqref{eq:DeepSets} with any other permutation-invariant aggregator (e.g., max or product), though we do not know if there are any claims to universality in these cases.
-Do area formulae exist for these other aggregators?
-
 {\footnotesize\par\bigskip\noindent
@@
   % Kendrick Smith (Perimeter),
+  Kate Storey-Fisher (Stanford),
   Ben Wandelt (JHU),
@@
 \label{sec:appendix}
+
+\paragraph{Additional formulae} We noted that symbolic regression revealed several equivalent but more complicated triangle area formulae which are polynomials in the multi-power sums.
+Below is a list of these.
+
+For brevity, we define
+\begin{equation}
+    F_{mn} \equiv \sum_{i=1}^{3} \, x_i^m \, y_i^n.
+\end{equation}
+For reference, our result \eqref{eq:result} is
+\begin{equation}
+4\,\Delta^2 =
++3\,F_{20}\,F_{02}
+-3\,F_{11}^2
++2\,F_{11}\,F_{10}\,F_{01}
+-1\,F_{20}\,F_{01}^2
+-1\,F_{02}\,F_{10}^2~.
+\end{equation}
+The additional formulae are
+\begin{align}
+4\,\Delta^2 =
+&
++6\,F_{22}
+-4\,F_{21}\,F_{01}
+-4\,F_{12}\,F_{10}
++2\,F_{20}\,F_{02}
+\nonumber\\ &
++6\,F_{11}\,F_{10}\,F_{01}
+-5\,F_{11}^2
+-1\,F_{10}^2\,F_{01}^2~,
+\\
+4\,\Delta^2 =
+&
++18\,F_{22}
+-12\,F_{21}\,F_{01}
+-12\,F_{12}\,F_{10}
+\nonumber\\ &
++2\,F_{20}\,F_{01}^2
++2\,F_{02}\,F_{10}^2
++14\,F_{11}\,F_{10}\,F_{01}
+-9\,F_{11}^2
+-3\,F_{10}^2\,F_{01}^2~,
+\\
+8\,\Delta^2 =
+&
+-6\,F_{22}
++4\,F_{21}\,F_{01}
++4\,F_{12}\,F_{10}
++7\,F_{20}\,F_{02}
+\nonumber\\ &
+-3\,F_{20}\,F_{01}^2
+-3\,F_{02}\,F_{10}^2
+-4\,F_{11}^2
++1\,F_{10}^2\,F_{01}^2~,
+\\
+8\,\Delta^2 =
+&
+-18\,F_{22}
++12\,F_{21}\,F_{01}
++12\,F_{12}\,F_{10}
++9\,F_{20}\,F_{02}
+\nonumber\\ &
+-5\,F_{20}\,F_{01}^2
+-5\,F_{02}\,F_{10}^2
+-8\,F_{11}\,F_{10}\,F_{01}
++3\,F_{10}^2\,F_{01}^2~,
+\end{align}
+and, our personal favorite,
+\begin{align}
+368\,\Delta^2 =
+&
+-54\,F_{22}
++36\,F_{21}\,F_{01}
++36\,F_{12}\,F_{10}
+\nonumber\\ &
++285\,F_{20}\,F_{02}
+-101\,F_{20}\,F_{01}^2
+-101\,F_{02}\,F_{10}^2
+\nonumber\\ &
+-258\,F_{11}^2
++148\,F_{11}\,F_{10}\,F_{01}
++9\,F_{10}^2\,F_{01}^2~.
+\end{align}
+We note that 368 is divisible by 23.
 
@@
 This form of the convolution theorem is proven below.
-The Fourier transform of the mass density is
-\begin{equation}
-    \mathcal{F}\{\varrho\}(\vec{k})
-    = \int \d^3 r
-        \, e^{-i \, \vec{k} \cdot \vec{r}}
-        \, \sum_{j=1}^{N}
-        \, m_j
-        \, \delta(\vec{r} - \vec{r}_j)
-    = \sum_{j=1}^{N} \, m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}~,
-\end{equation}
-where we sum over $j$ and use $i=\sqrt{-1}$.
 Thus, the total energy is given in terms of the Fourier transform of the density field by
@@
 \paragraph{Fourier-space triangle area formula}
-Suppose our triangle is specified by three vertices $\vec{r}_1,\vec{r}_2,\vec{r}_3 \in \mathbb{R}^d$. We represent this distribution by the field
+Suppose our triangle is specified by three vertices $\vec{r}_1,\vec{r}_2,\vec{r}_3 \in \mathbb{R}^d$. We represent this distribution by the density field
 \begin{equation}
@@
 \end{equation}
-Fourier transforming this field gives
+The Fourier transform of this field is
 \begin{equation}
@@
 \end{align}
-This is known as the Patterson function.
+This is known as the Patterson function in x-ray crystallography.
 
-By multiplying the Patterson function $\mathcal{P}$ by a function of $r \equiv \abs{\vec{r}}$ and integrating over all space, we can extract sums of functions of the side lengths. Explicitly,
+By multiplying the Patterson function $\mathcal{P}$ by a function of $r \equiv \abs{\vec{r}}$ and integrating over all space, we can compute sums of functions of the side lengths.
 \begin{align}
@@
 \end{align}
-In particular, we can define the $m^{\rm th}$ power sum of the side lengths
+In particular, we can compute the $m^{\rm th}$ power sum of the side lengths
 \begin{equation}
